### 2023-02-22 manuscript.tex
Initial Overleaf Import

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% This is a (brief) model paper using the achemso class
%% The document class accepts keyval options, which should include
%% the target journal and optionally the manuscript type. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[journal=jacsat,manuscript=article]{achemso}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Place any additional packages needed here.  Only include packages
%% which are essential, to avoid problems later. Do NOT use any
%% packages which require e-TeX (for example etoolbox): the e-TeX
%% extensions are not currently available on the ACS conversion
%% servers.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[version=3]{mhchem} % Formula subscripts using \ce{}
\usepackage{algorithm}
\usepackage{algpseudocode}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% If issues arise when submitting your manuscript, you may want to
%% un-comment the next line.  This provides information on the
%% version of every file you have used.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%\listfiles

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Place any additional macros here.  Please use \newcommand* where
%% possible, and avoid layout-changing macros (which are not used
%% when typesetting).
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand*\mycommand[1]{\texttt{\emph{#1}}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Meta-data block
%% ---------------
%% Each author should be given as a separate \author command.
%%
%% Corresponding authors should have an e-mail given after the author
%% name as an \email command. Phone and fax numbers can be given
%% using \phone and \fax, respectively; this information is optional.
%%
%% The affiliation of authors is given after the authors; each
%% \affiliation command applies to all preceding authors not already
%% assigned an affiliation.
%%
%% The affiliation takes an option argument for the short name.  This
%% will typically be something like "University of Somewhere".
%%
%% The \altaffiliation macro should be used for new address, etc.
%% On the other hand, \alsoaffiliation is used on a per author basis
%% when authors are associated with multiple institutions.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\author{Nicholas Runchie}
\altaffiliation{EaSTCHEM School of Chemistry, University of Edinburgh, EH9 3FJ}
\author{Antonia S.J.S. Mey}
\altaffiliation{EaSTCHEM School of Chemistry, University of Edinburgh, EH9 3FJ}
\email{antonia.mey@ed.ac.uk}
\phone{xxx}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The document title should be given as usual. Some journals require
%% a running title from the author: this should be supplied as an
%% optional argument to \title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title[SILVR: Molecular Generation for binding modes]
  {SILVR: Conditional diffusion model for molecule generation without additional training}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Some journals require a list of abbreviations or keywords to be
%% supplied. These should be set up here, and will be printed after
%% the title and author information, if needed.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\abbreviations{}
\keywords{ML, generative models, docking}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The manuscript does not need to include \maketitle, which is
%% executed automatically.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The "tocentry" environment can be used to create an entry for the
%% graphical table of contents. It is given here as some journals
%% require that it is printed as part of the abstract page. It will
%% be automatically moved as appropriate.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{tocentry}

Some journals require a graphical entry for the Table of Contents.
This should be laid out ``print ready'' so that the sizing of the
text is correct.

Inside the \texttt{tocentry} environment, the font used is Helvetica
8\,pt, as required by \emph{Journal of the American Chemical
Society}.

The surrounding frame is 9\,cm by 3.5\,cm, which is the maximum
permitted for  \emph{Journal of the American Chemical Society}
graphical table of content entries. The box will not resize if the
content is too big: instead it will overflow the edge of the box.

This box and the associated title will always be printed on a
separate page at the end of the document.

\end{tocentry}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The abstract environment will automatically gobble the contents
%% if an abstract is not used by the target journal.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
I'll be an abstract when I grow up. 
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Start the main part of the manuscript here.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
- Equivariant diffusion models have been very promising in predicting 3D structures of drug-like molecules. 
- This does not make them suitable candidates for a target binding site. 
- We combine equivariant diffusion models~\cite{huang2022mdm} with Iterative latent variable refinement proposed in Denoising Diffusion probabilistic models~\cite{choi2021ilvr}.
- This allows the generation of new molecules in the shape of the binding site using information from existing fragments.

\section{Theory}
\subsectin{Unconditional Equivariant Denoising Diffusion probabilistic models}
The main idea behind diffusion models is for a neural netowrk to learn the reverse of a diffusion process, often referred to as denoising. If you start out with a set of data $\mathbf{x}$ and add noise to this data over a time interval t = [0,\ldots,T] you can define the following distribution:
\begin{equation}
  p(\mathbf{z}_t)|\mathbf{x} = \mathcal{N}(\mathbf{z}_t|\alpha_t,\amthbf{x}_t,\sigma^2_t\mathbf{I} ) 
\end{equation}

We do not need to know $\matbf{x}$ and want to approximate this using a neural network $\phi$, according to $\mathbf{x} = \phi(\mathbf{z}_t,t)$
\subsection{Iterative Latent Variable Refinement as a conditioning tool}
Choi et al.~\cite{choi2021ilvr} introduced a way to condition DDPMs by sampling their images from the conditional distribution $p(x_0|c)$, with the condition c:
\begin{equation}
    P_\theta(x_0|c) = \int p_\theta(x_{0:T}|c)Dx_{1:T}
\end{equation}
\subsection{Selective Iterative Variable Refinement}

\begin{algorithm}
\caption{SILVR}\label{alg:cap}
\begin{algorithmic}
\State Sample $x_T \sim \mathcal{N}(0,I)$
\For {$t$ = $T,...,1$}
    \State $z \sim \mathcal{N}(0,I)$
    \State $x_{t-1}^\prime \sim p_\theta(x_{t-1}^\prime|x_t)$
    \State $y_{t-1} \sim q(y_{t-1}^\prime|y)$
    \State $x_{t-1} \gets x_{t-1}^\prime - k x_{t-1}^\prime +k y_{t-1}$ \Comment{SILVR equation}

\EndFor
\State Sample $x,h \sim p(x,h|z_0)$ 
\end{algorithmic}
\end{algorithm}

\subsecsection{Algorithm}
Both of these are wrong. I have just quickly tried to transcribe them and try and modify them. I think the ILVR style is easier to read and convey the point clearly. 
\begin{algorithm}
\caption{Sampling from EDM (coppied from EDM)}\label{alg:cap}
\begin{algorithmic}
\State Sample $z_T \sim \mathcal{N}(0,I)$
\For {$t$ in $T, T-1,...,1$ where $s=t-1$}
    \State Sample $\epsilon \sim \mathcal{N}(0,I)$
    \State Subtract center of gravity from $\epsilon^{(x)}$ in $\epsilon = [\epsilon^{(x)}, \epsilon^{(h)}]$
    \State $z_s^\prime = \frac{1}{\alpha_{t|s}}z_t - \frac{\sigma^2_{t|s}}{\alpha_{t|s}\sigma_t} \cdot \phi(z_t,t) + \sigma_{t \to s} \cdot \epsilon$
    \State $z_s \gets z_s^\prime - \lambda z_s^\prime +\lambda z_{ref}$ \Comment{SILVR equation}
\EndFor
\State Sample $x,h \sim p(x,h|z_0)$ 
\end{algorithmic}
\end{algorithm}

\section{Methods}

\section{Results}

\section{Discussion and Conclusions}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The "Acknowledgement" section can be given in all manuscript
%% classes.  This should be given within the "acknowledgement"
%% environment, which will make the correct section or running title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{acknowledgement}
The authors thank \ldots.

\end{acknowledgement}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The appropriate \bibliography command should be placed here.
%% Notice that the class file automatically sets \bibliographystyle
%% and also names the section correctly.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliography{achemso-demo}

\end{document}

### 2023-03-08 manuscript.tex
Updates from Overleaf
committing theory

@@
 \usepackage{algpseudocode}
+\usepackage{svg}
+\usepackage{amsmath}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\author{Nicholas Runchie}
+\author{Nicholas Runcie}
 \altaffiliation{EaSTCHEM School of Chemistry, University of Edinburgh, EH9 3FJ}
@@
 \section{Theory}
-\subsectin{Unconditional Equivariant Denoising Diffusion probabilistic models}
-The main idea behind diffusion models is for a neural netowrk to learn the reverse of a diffusion process, often referred to as denoising.
-If you start out with a set of data $\mathbf{x}$ and add noise to this data over a time interval t = [0,\ldots,T] you can define the following distribution:
+\subsection{Unconditional Equivariant Denoising Diffusion probabilistic models}
+\subsubsection*{Denoising Diffusion probabilistic models (DDPM) as generative models}
+DDPMs are often used as generative models that were developed for the generation of new images~\cite{}.
+More recently the same idea has also been applied to molecular generation~\cite{}.
+The main idea behind diffusion models is for a neural network to learn the reverse of a diffusion process, often referred to as denoising to \textit{sample} a new image or new molecule.
+In practice this is done by training a neural network $\phi$ and generating samples $\hat{\mathbf{x}}=\psi(X_t,t)$, with $X_t$ with data $X$, noised up to time $t$.
+DDPMs consist of two main parts the diffusion process and the denoising process, see the schematic in figure~\ref{fig:diffusion_schematic}.
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{Figures/model_no_stickman.png}
+    \caption{This should show the basic idea of a diffusion model}
+    \label{fig:diffusion_schematic}
+\end{figure}
+The \textit{forward} part of the diffusion model is a Markov process, where you start out with a set of data $\mathbf{x}$ and add noise to this data over a time interval t = [0,\ldots,T] according to the following distribution:
 \begin{equation}
-  p(\mathbf{z}_t)|\mathbf{x} = \mathcal{N}(\mathbf{z}_t|\alpha_t,\amthbf{x}_t,\sigma^2_t\mathbf{I} ) 
-\end{equation}
-
-We do not need to know $\matbf{x}$ and want to approximate this using a neural network $\phi$, according to $\mathbf{x} = \phi(\mathbf{z}_t,t)$
+  q(\mathbf{x_{1:T}})|\mathbf{x_0} = q(x_0)\prod_{t=1}^T q(\mathbf{x}_t|\mathbf{x}_{t-1})
+\end{equation}.
+The product of conditional probabilities $q(..)$ can be modelled as a Gaussian Kernel given by:
+\begin{equation}
+    q(x_t, x_s)=\mathcal{N}(x_t|\alpha_t/\alpha_s, \sigma_t^2-\frac{\alpha_t}{\alpha_s}\sigma^2_s
+\end{equation},
+for any $t> s$. The parameters $\alpha_t \in \mathbb{R}^+$ contains the amount of retained signal and $\sigma_t \in \mathbb{R}^+$ represents the variance and thus the amount of white noise added. 
+Different people have looked at different noise schedules~\cite{sohl-dickstein2015deep, ho2020denoising}.
+% Next talk about the denoising step
+$\hat{p}$ is the parametric probability distribution we want to learn. 
+\subsubsection*{Euquivariant diffusion model}
+Unlike in image problems the orientation of molecules matters under translations and rotations.
+For this purpose, we chose to equivariant diffusion model (EDM) by Hoogeboom et al.~\cite{} as our baseline generative model.
+The basic concept behind equivarience is that the model is invariant to rotations and translation in this case the E(3) group.
+This means that scalar (features such as atom types) and vector node properties (such as the positions) will be invariant to group transformations.
+This means that the order in which a rotation is applied does not matter.
+You can rotate the input to the model then apply the diffusion/denoising process and get a structure out, or you can apply the diffusion/denoising process first and then do the same rotation to get the same output.
+Mathematically this means that if we have a set of point $\mathbf{x} = (\mathbf{x}_1,\ldots,\mathbf{x}_N) \in \mathbb{^{N\times 3}}$ and each of these points has a set of scalar feature vectors $\mathbf{h}\in \mathbb{R}^{N\times k}$ associated with it these features are invariant to group transformations.
+The positions translations and rotations are defined according to the orthogonal matrix: $\mathbf{Rx + t} = (\mathbf{R_{x_1}+t},\ldots, \mathbf{R_{x_N}+t}$.
+%%%%
+% Subsection: ILVR
+%%%%
 \subsection{Iterative Latent Variable Refinement as a conditioning tool}
@@
 \end{equation}
-\subsection{Selective Iterative Variable Refinement}
+
+%%%%
+% Subsection: SILVR
+%%%%
+\subsection{Selective Iterative Latent Variable Refinement}
 
@@
 
-\subsecsection{Algorithm}
-Both of these are wrong. I have just quickly tried to transcribe them and try and modify them. I think the ILVR style is easier to read and convey the point clearly. 
-\begin{algorithm}
-\caption{Sampling from EDM (coppied from EDM)}\label{alg:cap}
-\begin{algorithmic}
-\State Sample $z_T \sim \mathcal{N}(0,I)$
-\For {$t$ in $T, T-1,...,1$ where $s=t-1$}
-    \State Sample $\epsilon \sim \mathcal{N}(0,I)$
-    \State Subtract center of gravity from $\epsilon^{(x)}$ in $\epsilon = [\epsilon^{(x)}, \epsilon^{(h)}]$
-    \State $z_s^\prime = \frac{1}{\alpha_{t|s}}z_t - \frac{\sigma^2_{t|s}}{\alpha_{t|s}\sigma_t} \cdot \phi(z_t,t) + \sigma_{t \to s} \cdot \epsilon$
-    \State $z_s \gets z_s^\prime - \lambda z_s^\prime +\lambda z_{ref}$ \Comment{SILVR equation}
-\EndFor
-\State Sample $x,h \sim p(x,h|z_0)$ 
-\end{algorithmic}
-\end{algorithm}
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{Figures/minimal_model.png}
+    \caption{Minimal architecture figure - I prefer this style as it fits better with the theme}
+    \label{fig:fig_1}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{Figures/molecules_silvr_rates_horizontal.png}
+    \caption{This looks much better than I thought}
+    \label{fig:fig_2}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{Figures/silvr_plots.png}
+    \caption{Preliminary SILVR results (labels missing).
+A) Ratio of stable atoms by considering bond distances (note this includes protons but strictly maybe they should be removed.
+I would have to make new code for that I think).
+B) RMSD between samples and reference coordinated - pairwise mapping (samples rotated and translated to minimise RMSD).
+C)Openeye Shapegauss - still concerned I have run this wrong - lower value indicates better shape agreement with receptor D) Geometry stability - samples converted to smiles, geometry predicted using Auto3D, then RMSD with sample}
+    \label{fig:fig_3}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{Figures/fragment_sample_bound.png}
+    \caption{Examples of other reference sets and samples fitting binding site}
+    \label{fig:fig_5}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includesvg{Figures/silvr_model_svg.svg}
+    \caption{SVG of SILVR model - text not correctly aligned. Opinion vs previous figure.}
+    \label{fig:fig_6}
+\end{figure}
 
 \section{Methods}
+The selective iterative latent variable refinement (SILVR) method makes use of the pre-trained equivariant diffusion model (EDM) as introduced by Hegeboom et al.~\cite{}.
+The core of the method is the introduction of an additional refinement within the denoising process during run time.
+The resulting SILVR model can then produce conditional samples without any conditional training.
+In this work, we show how SILVR can be applied to fragment-based drug design to produce molecules resembling a reference set of atomic coordinates.
+
+The EDM by Hoodgeboom et al was trained on the 30 lowest energy conformations of 430,000 molecules from the GEOM dataset with explicit protons.
+A simplified schematic describing EDM architecture is shared in scheme X for more details on the model theory and training refer to Hegeboom et al.~\cite{}.
+This model strictly only considers atomic coordinates; all bond information is ignored.
+During training, the atomic coordinates (and their element) are passed through a forward diffusion process with iterative addition of Gaussian noise, where the extent of noise added at each step is defined by parameter $\beta$ (N.B.
+scheme X shows the diffusion process as a Markov chain, however in practice the state at time $t$ is determined as a direct transformation of the initial state).
+The diffusion process is eventually terminated when $t=T$, by which point all structure is lost and all coordinates follow a gaussian distribution.
+An equivariant graph neural network (EGNN) is then trained to predict the reverse process, predicting the previous state in the sequence given any state.
+At run time, the generative model is instantiated with a sample from a gaussian distribution and the series of denoising steps are applied resulting in a generated sample consisting of atomic coordinates resembling a drug-like molecule.
+The XYZ coordinates can then be interpreted using cheminformatics software to determine the molecular graph.
+As described, the EDM model is only trained to generate random molecules; hoodgeboom’s model does enable training given additional context, however this is outside the scope of this paper.
 

### 2023-03-08 paper/SI.tex
restructuring

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The same is true for Supporting Information, which should use the
%% suppinfo environment.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{suppinfo}

This will usually read something like: ``Experimental procedures and
characterization data for all new compounds. The class will
automatically add a sentence pointing to the information on-line:

\end{suppinfo}

### 2023-03-08 silvr_projects/silvr_demo.py
add demo notebook and script

@@ -1,0 +1,410 @@
+"""
+The maintainability of this code is poor and shoud be improved.
+This file was not created with the entent to be read by anyone else
+"""
+
+
+import sys
+sys.path
+_ = (sys.path.append("/usr/local/lib/python3.8/site-packages"))
+_ = (sys.path.append("e3_diffusion_for_molecules"))
+
+import eval_sample
+
+# Rdkit import should be first, do not move it
+try:
+    from rdkit import Chem
+except ModuleNotFoundError:
+    pass
+
+import utils
+import argparse
+from configs.datasets_config import qm9_with_h, qm9_without_h
+from qm9 import dataset
+from qm9.models import get_model
+
+from equivariant_diffusion.utils import assert_correctly_masked
+import torch
+import pickle
+import qm9.visualizer as vis
+from qm9.analyze import check_stability
+from os.path import join
+from qm9.sampling import sample_chain, sample
+from configs.datasets_config import get_dataset_info
+
+
+################
+
+
+import torch
+from torch.distributions.categorical import Categorical
+
+import numpy as np
+from egnn.models import EGNN_dynamics_QM9
+
+from equivariant_diffusion.en_diffusion import EnVariationalDiffusion
+
+
+
+####################
+
+
+import numpy as np
+import torch
+import torch.nn.functional as F
+from equivariant_diffusion.utils import assert_mean_zero_with_mask, remove_mean_with_mask,\
+    assert_correctly_masked
+from qm9.analyze import check_stability
+
+
+####################
+from qm9.analyze import check_stability
+from argparse import Namespace
+import datetime
+import os
+
+
+####################
+from multiprocessing import Process
+import time
+import math
+import datetime
+
+
+
+
+def read_sxyz(file):
+    """
+    Read reference.sxyz file and parse into correct format for silvr
+    
+    ---explanation of parsing---
+    Line[0] contains total atoms in the reference. This is the same as for an xyz file
+    line[1] contains experimental set up params
+        dummy: Number of additional atoms to add that are not contained within the reference
+        sample: how many molecules should be generated. Molecules are generated serially, so samples
+                defines iterations in a for loop
+                nb: future work should look at parallelisation
+        comment: This does not affect model running
+        
+    lines[2:] contain all xyz and silvr rate information: element X Y Z rate
[…]

### 2023-03-09 silvr_projects/docking_oe/00-prep-all-receptors.py
dump all openeye scripts

@@ -1,0 +1,399 @@
+"""
+Prepare all SARS-CoV-2 Mpro structures for docking and simulation in monomer and dimer forms
+
+This should be run from the covid-moonshot/scripts directory
+
+"""
+import rich
+import openeye
+
+structures_path = '../structures'
+output_basepath = '../receptors'
+
+# REMARK 350 (biological multimer symmetry operations) from 5RGG
+# TODO: Remove this when REMARK 350 is restored
+BIOLOGICAL_SYMMETRY_HEADER = """\
+REMARK 350
+REMARK 350 COORDINATES FOR A COMPLETE MULTIMER REPRESENTING THE KNOWN
+REMARK 350 BIOLOGICALLY SIGNIFICANT OLIGOMERIZATION STATE OF THE
+REMARK 350 MOLECULE CAN BE GENERATED BY APPLYING BIOMT TRANSFORMATIONS
+REMARK 350 GIVEN BELOW.  BOTH NON-CRYSTALLOGRAPHIC AND
+REMARK 350 CRYSTALLOGRAPHIC OPERATIONS ARE GIVEN.
+REMARK 350
+REMARK 350 BIOMOLECULE: 1
+REMARK 350 AUTHOR DETERMINED BIOLOGICAL UNIT: DIMERIC
+REMARK 350 SOFTWARE DETERMINED QUATERNARY STRUCTURE: DIMERIC
+REMARK 350 SOFTWARE USED: PISA
+REMARK 350 TOTAL BURIED SURFACE AREA: 4170 ANGSTROM**2
+REMARK 350 SURFACE AREA OF THE COMPLEX: 25430 ANGSTROM**2
+REMARK 350 CHANGE IN SOLVENT FREE ENERGY: -3.0 KCAL/MOL
+REMARK 350 APPLY THE FOLLOWING TO CHAINS: A
+REMARK 350   BIOMT1   1  1.000000  0.000000  0.000000        0.00000
+REMARK 350   BIOMT2   1  0.000000  1.000000  0.000000        0.00000
+REMARK 350   BIOMT3   1  0.000000  0.000000  1.000000        0.00000
+REMARK 350   BIOMT1   2 -1.000000  0.000000  0.000000        0.00000
+REMARK 350   BIOMT2   2  0.000000  1.000000  0.000000        0.00000
+REMARK 350   BIOMT3   2  0.000000  0.000000 -1.000000        0.00000
+"""
+
+def download_url(url, save_path, chunk_size=128):
+    """
[…]

### 2023-03-09 silvr_projects/docking_oe/02-dock-and-prep.py

@@ -1,0 +1,737 @@
+"""
+Dock specified ligand to all DiamondMX Mpro structures and prepare for alchemical free energy calculations
+
+"""
+
+covalent_warhead_precedence = [
+    'acrylamide', 'acrylamide_adduct', 'chloroacetamide', 'chloroacetamide_adduct', 'vinylsulfonamide', 'vinylsulfonamide_adduct', 'nitrile', 'nitrile_adduct',
+]
+
+covalent_warhead_smarts = {
+    'acrylamide' : '[C;H2:1]=[C;H1]C(N)=O',
+    'acrylamide_adduct' : 'NC(C[C:1]S)=O',
+    'chloroacetamide' : 'Cl[C;H2:1]C(N)=O',
+    'chloroacetamide_adduct' : 'S[C:1]C(N)=O',
+    'vinylsulfonamide' : 'NS(=O)([C;H1]=[C;H2:1])=O',
+    'vinylsulfonamide_adduct' : 'NS(=O)(C[C:1]S)=O',
+    'nitrile' : 'N#[C:1]-[*]',
+    'nitrile_adduct' : 'C-S-[C:1](=N)',
+    }
+
+# Read ligands
+def read_csv_molecules(filename):
+    """Read molecules from the specified path
+
+    Parameters
+    ----------
+    filename : str
+        File from which molecules are to be read
+
+    Returns
+    -------
+    molecules : list of openeye.oechem.OEMol
+        The read molecules
+    """
+
+    from openeye import oechem
+    mol = oechem.OEMol()
+    molecules = list()
+    with oechem.oemolistream(filename) as ifs:
+        while oechem.OEReadCSVFile(ifs, mol):
+            molecules.append(oechem.OEMol(mol))
+    # TODO: Also store original SMILES as "original SMILES"
+    return molecules
+
+def score(molecule, field='Hybrid2'):
+    """Return the docking score"""
+    from openeye import oechem
+    value = oechem.OEGetSDData(molecule, field)
+    return float(value)
+
+def find_warheads(molecule):
+    """Find covalent warheads
+
+    Parameters
+    ----------
+    molecule : OEMol
+      The molecule
+
+    Returns
+    -------
+    warheads_found : dict
+       warheads_found[warhead_type] corresponds to atom index of heavy atom that forms covalent adduct
+
+    """
+    # Identify highest priority warhead found in compound
+    print('Finding warhead atoms...')
+    warheads_found = dict()
[…]

### 2023-03-09 silvr_projects/docking_oe/CleanThenDockMolecules.py

@@ -1,0 +1,91 @@
+#!/usr/bin/env python
+# (C) 2017 OpenEye Scientific Software Inc. All rights reserved.
+#
+# TERMS FOR USE OF SAMPLE CODE The software below ("Sample Code") is
+# provided to current licensees or subscribers of OpenEye products or
+# SaaS offerings (each a "Customer").
+# Customer is hereby permitted to use, copy, and modify the Sample Code,
+# subject to these terms. OpenEye claims no rights to Customer's
+# modifications. Modification of Sample Code is at Customer's sole and
+# exclusive risk. Sample Code may require Customer to have a then
+# current license or subscription to the applicable OpenEye offering.
+# THE SAMPLE CODE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND,
+# EXPRESS OR IMPLIED.  OPENEYE DISCLAIMS ALL WARRANTIES, INCLUDING, BUT
+# NOT LIMITED TO, WARRANTIES OF MERCHANTABILITY, FITNESS FOR A
+# PARTICULAR PURPOSE AND NONINFRINGEMENT. In no event shall OpenEye be
+# liable for any damages or liability in connection with the Sample Code
+# or its use.
+
+import sys
+
+from openeye import oechem
+from openeye import oedocking
+from openeye import oequacpac
+
+
+def main(argv=[__name__]):
+    dockOpts = oedocking.OEDockOptions()
+    opts = oechem.OERefInputAppOptions(dockOpts, "DockMolecules", oechem.OEFileStringType_Mol3D,
+                                       oechem.OEFileStringType_Mol3D, oechem.OEFileStringType_DU, "-receptor")
+    if oechem.OEConfigureOpts(opts, argv, False) == oechem.OEOptsConfigureStatus_Help:
+        return 0
+    dockOpts.UpdateValues(opts)
+
+    ifs = oechem.oemolistream()
+    if not ifs.open(opts.GetInFile()):
+        oechem.OEThrow.Fatal("Unable to open %s for reading" % opts.GetInFile())
+
+    rfs = oechem.oeifstream()
+    if not rfs.open(opts.GetRefFile()):
+        oechem.OEThrow.Fatal("Unable to open %s for reading" % opts.GetRefFile())
+
+    ofs = oechem.oemolostream()
[…]

### 2023-03-09 silvr_projects/docking_oe/DockMolecules.py

@@ -1,0 +1,78 @@
+#!/usr/bin/env python
+# (C) 2017 OpenEye Scientific Software Inc. All rights reserved.
+#
+# TERMS FOR USE OF SAMPLE CODE The software below ("Sample Code") is
+# provided to current licensees or subscribers of OpenEye products or
+# SaaS offerings (each a "Customer").
+# Customer is hereby permitted to use, copy, and modify the Sample Code,
+# subject to these terms. OpenEye claims no rights to Customer's
+# modifications. Modification of Sample Code is at Customer's sole and
+# exclusive risk. Sample Code may require Customer to have a then
+# current license or subscription to the applicable OpenEye offering.
+# THE SAMPLE CODE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND,
+# EXPRESS OR IMPLIED.  OPENEYE DISCLAIMS ALL WARRANTIES, INCLUDING, BUT
+# NOT LIMITED TO, WARRANTIES OF MERCHANTABILITY, FITNESS FOR A
+# PARTICULAR PURPOSE AND NONINFRINGEMENT. In no event shall OpenEye be
+# liable for any damages or liability in connection with the Sample Code
+# or its use.
+
+import sys
+
+from openeye import oechem
+from openeye import oedocking
+
+
+def main(argv=[__name__]):
+    dockOpts = oedocking.OEDockOptions()
+    opts = oechem.OERefInputAppOptions(dockOpts, "DockMolecules", oechem.OEFileStringType_Mol3D,
+                                       oechem.OEFileStringType_Mol3D, oechem.OEFileStringType_DU, "-receptor")
+    if oechem.OEConfigureOpts(opts, argv, False) == oechem.OEOptsConfigureStatus_Help:
+        return 0
+    dockOpts.UpdateValues(opts)
+
+    ifs = oechem.oemolistream()
+    if not ifs.open(opts.GetInFile()):
+        oechem.OEThrow.Fatal("Unable to open %s for reading" % opts.GetInFile())
+
+    rfs = oechem.oeifstream()
+    if not rfs.open(opts.GetRefFile()):
+        oechem.OEThrow.Fatal("Unable to open %s for reading" % opts.GetRefFile())
+
+    ofs = oechem.oemolostream()
[…]

### 2023-03-09 silvr_projects/docking_oe/MakeReceptor.py

@@ -1,0 +1,50 @@
+#!/usr/bin/env python
+# (C) 2017 OpenEye Scientific Software Inc. All rights reserved.
+#
+# TERMS FOR USE OF SAMPLE CODE The software below ("Sample Code") is
+# provided to current licensees or subscribers of OpenEye products or
+# SaaS offerings (each a "Customer").
+# Customer is hereby permitted to use, copy, and modify the Sample Code,
+# subject to these terms. OpenEye claims no rights to Customer's
+# modifications. Modification of Sample Code is at Customer's sole and
+# exclusive risk. Sample Code may require Customer to have a then
+# current license or subscription to the applicable OpenEye offering.
+# THE SAMPLE CODE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND,
+# EXPRESS OR IMPLIED.  OPENEYE DISCLAIMS ALL WARRANTIES, INCLUDING, BUT
+# NOT LIMITED TO, WARRANTIES OF MERCHANTABILITY, FITNESS FOR A
+# PARTICULAR PURPOSE AND NONINFRINGEMENT. In no event shall OpenEye be
+# liable for any damages or liability in connection with the Sample Code
+# or its use.
+
+import sys
+
+from openeye import oechem
+from openeye import oedocking
+
+
+def main(argv=[__name__]):
+    recOpts = oedocking.OEMakeReceptorOptions()
+    opts = oechem.OESimpleAppOptions(recOpts, "MakeReceptor", oechem.OEFileStringType_DU, oechem.OEFileStringType_DU)
+    if oechem.OEConfigureOpts(opts, argv, False) == oechem.OEOptsConfigureStatus_Help:
+        return 0
+    recOpts.UpdateValues(opts)
+
+    ifs = oechem.oeifstream()
+    if not ifs.open(opts.GetInFile()):
+        oechem.OEThrow.Fatal("Unable to open %s for reading" % opts.GetInFile())
+
+    ofs = oechem.oeofstream()
+    if not ofs.open(opts.GetOutFile()):
+        oechem.OEThrow.Fatal("Unable to open %s for writing" % opts.GetOutFile())
+
+    du = oechem.OEDesignUnit()
+    while oechem.OEReadDesignUnit(ifs, du):
+        if oedocking.OEMakeReceptor(du, recOpts):
+            oechem.OEWriteDesignUnit(ofs, du)
+        else:
[…]

### 2023-03-09 silvr_projects/docking_oe/RescorePoses.py

@@ -1,0 +1,104 @@
+#!/usr/bin/env python
+# (C) 2017 OpenEye Scientific Software Inc. All rights reserved.
+#
+# TERMS FOR USE OF SAMPLE CODE The software below ("Sample Code") is
+# provided to current licensees or subscribers of OpenEye products or
+# SaaS offerings (each a "Customer").
+# Customer is hereby permitted to use, copy, and modify the Sample Code,
+# subject to these terms. OpenEye claims no rights to Customer's
+# modifications. Modification of Sample Code is at Customer's sole and
+# exclusive risk. Sample Code may require Customer to have a then
+# current license or subscription to the applicable OpenEye offering.
+# THE SAMPLE CODE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND,
+# EXPRESS OR IMPLIED.  OPENEYE DISCLAIMS ALL WARRANTIES, INCLUDING, BUT
+# NOT LIMITED TO, WARRANTIES OF MERCHANTABILITY, FITNESS FOR A
+# PARTICULAR PURPOSE AND NONINFRINGEMENT. In no event shall OpenEye be
+# liable for any damages or liability in connection with the Sample Code
+# or its use.
+
+import sys
+
+from openeye import oechem
+from openeye import oedocking
+
+
+def main(argv=[__name__]):
+    itf = oechem.OEInterface(InterfaceData)
+    # @ <SNIPPET-RESCORE-POSES-CONFIGURE>
+    oedocking.OEScoreTypeConfigure(itf, "-score")
+    # @ </SNIPPET-RESCORE-POSES-CONFIGURE>
+    if not oechem.OEParseCommandLine(itf, argv):
+        return 1
+
+    receptor = oechem.OEDesignUnit()
+    if not oechem.OEReadDesignUnit(itf.GetString("-receptor"), receptor):
+        oechem.OEThrow.Fatal("Unable to read receptor")
+    imstr = oechem.oemolistream()
+    if not imstr.open(itf.GetString("-in")):
+        oechem.OEThrow.Fatal("Unable to open input file of ligands")
+    omstr = oechem.oemolostream()
+    if not omstr.open(itf.GetString("-out")):
+        oechem.OEThrow.Fatal("Unable to open out file for rescored ligands")
+        
[…]

### 2023-03-09 silvr_projects/docking_oe/docking_full.py

@@ -1,0 +1,159 @@
+#!/usr/bin/env python
+# coding: utf-8
+
+# # Run below for all molecule 
+
+# In[1]:
+
+
+"""
+#-------old code--------------
+
+import os
+import glob
+
+exp_list = glob.glob("../experiments/exp_*")
+
+
+if not os.path.exists(f'chemgauss_docking/'):
+    os.makedirs(f'chemgauss_docking/')
+        
+        
+for exp in exp_list:
+    exp_id = exp.split("/")[-1]
+    
+    if not os.path.exists(f'chemgauss_docking/{exp_id}/'):
+        os.makedirs(f'chemgauss_docking/{exp_id}/')
+        
+        
+    
+    ligands = glob.glob(f"../experiments/{exp_id}/*.txt")
+
+    for ligand in ligands:
+        mol_id = ligand.split("/")[-1].split(".")[0]
+
+        with open(ligand, "r") as readfile:
+            file = readfile.read()
+
+        with open("tmp.xyz","w") as writefile:
+            writefile.write(file)
+
+        #os.system(f'python DockMolecules.py -in tmp.xyz -out chemgauss_docking/{exp_id}/{mol_id}.sdf -receptor mpro2.oedu')
+        
+"""
+
+
+# In[2]:
+
+
+import os
+import glob
+
+#------New code-------
+#This could probably be set to run in parallel
+#Only molecules that are not fragmented get docked
+
+exp_list = glob.glob("../experiments/exp_*")
+
+if not os.path.exists(f'chemgauss_docking_cleaned/'):
+    os.makedirs(f'chemgauss_docking_cleaned/')
+        
+for exp in exp_list:
+    exp_id = exp.split("/")[-1]
+    
+    if not os.path.exists(f'chemgauss_docking_cleaned/{exp_id}/'):
+        os.makedirs(f'chemgauss_docking_cleaned/{exp_id}/')
+        
+        
+    
+    ligands = glob.glob(f"../experiments/{exp_id}/*.txt")
+
+    for ligand in ligands:
+        mol_id = ligand.split("/")[-1].split(".")[0]
+
+        with open(ligand, "r") as readfile:
+            file = readfile.read()
+
+        with open("tmp.xyz","w") as writefile:
+            writefile.write(file)
+
[…]

### 2023-03-09 silvr_projects/docking_oe/make_apo_design_units.py

@@ -1,0 +1,94 @@
+#!/usr/bin/env python
+# (C) 2018 OpenEye Scientific Software Inc. All rights reserved.
+#
+# TERMS FOR USE OF SAMPLE CODE The software below ("Sample Code") is
+# provided to current licensees or subscribers of OpenEye products or
+# SaaS offerings (each a "Customer").
+# Customer is hereby permitted to use, copy, and modify the Sample Code,
+# subject to these terms. OpenEye claims no rights to Customer's
+# modifications. Modification of Sample Code is at Customer's sole and
+# exclusive risk. Sample Code may require Customer to have a then
+# current license or subscription to the applicable OpenEye offering.
+# THE SAMPLE CODE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND,
+# EXPRESS OR IMPLIED.  OPENEYE DISCLAIMS ALL WARRANTIES, INCLUDING, BUT
+# NOT LIMITED TO, WARRANTIES OF MERCHANTABILITY, FITNESS FOR A
+# PARTICULAR PURPOSE AND NONINFRINGEMENT. In no event shall OpenEye be
+# liable for any damages or liability in connection with the Sample Code
+# or its use.
+
+#############################################################################
+# Script to prepare proteins into design units
+#############################################################################
+import sys
+import os
+from openeye import oechem
+from openeye import oegrid
+from openeye import oespruce
+
+
+def main(argv=sys.argv):
+
+    if len(argv) < 3 or len(argv) > 5:
+        oechem.OEThrow.Usage(
+            "%s <infile> <site_residue> [<mtzfile>] [<loopdbfile>]" % argv[0]
+        )
+
+    ifs = oechem.oemolistream()
+    ifile = argv[1]
+    if not ifs.open(ifile):
+        oechem.OEThrow.Fatal("Unable to open %s for reading" % ifile)
+
+    site_residue = argv[2]
+
+    include_loop = False
+    include_ed = False
+    if len(argv) > 3:
+        if len(argv) == 5 or (len(argv) == 4 and "mtz" in argv[3]):
+            edfile = argv[3]
+            ed = oegrid.OESkewGrid()
+            if not oegrid.OEReadMTZ(edfile, ed, oegrid.OEMTZMapType_Fwt):
+                oechem.OEThrow.Fatal(
[…]

### 2023-03-09 silvr_projects/docking_oe/make_design_units.py

@@ -1,0 +1,90 @@
+#!/usr/bin/env python
+# (C) 2018 OpenEye Scientific Software Inc. All rights reserved.
+#
+# TERMS FOR USE OF SAMPLE CODE The software below ("Sample Code") is
+# provided to current licensees or subscribers of OpenEye products or
+# SaaS offerings (each a "Customer").
+# Customer is hereby permitted to use, copy, and modify the Sample Code,
+# subject to these terms. OpenEye claims no rights to Customer's
+# modifications. Modification of Sample Code is at Customer's sole and
+# exclusive risk. Sample Code may require Customer to have a then
+# current license or subscription to the applicable OpenEye offering.
+# THE SAMPLE CODE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND,
+# EXPRESS OR IMPLIED.  OPENEYE DISCLAIMS ALL WARRANTIES, INCLUDING, BUT
+# NOT LIMITED TO, WARRANTIES OF MERCHANTABILITY, FITNESS FOR A
+# PARTICULAR PURPOSE AND NONINFRINGEMENT. In no event shall OpenEye be
+# liable for any damages or liability in connection with the Sample Code
+# or its use.
+
+#############################################################################
+# Script to prepare proteins into design units
+#############################################################################
+import sys
+import os
+from openeye import oechem
+from openeye import oegrid
+from openeye import oespruce
+
+
+def main(argv=sys.argv):
+
+    if len(argv) < 2 or len(argv) > 4:
+        oechem.OEThrow.Usage("%s <infile> [<mtzfile>] [<loopdbfile>]" % argv[0])
+
+    ifs = oechem.oemolistream()
+    ifile = argv[1]
+    if not ifs.open(ifile):
+        oechem.OEThrow.Fatal("Unable to open %s for reading" % ifile)
+
+    include_loop = False
+    include_ed = False
+    if len(argv) > 2:
+        if len(argv) == 4 or (len(argv) == 3 and "mtz" in argv[2]):
+            edfile = argv[2]
+            ed = oegrid.OESkewGrid()
+            if not oegrid.OEReadMTZ(edfile, ed, oegrid.OEMTZMapType_Fwt):
+                oechem.OEThrow.Fatal(
+                    "Unable to read electron density file %s" % edfile
+                )  # noqa
[…]

### 2023-03-09 silvr_projects/docking_oe/proteinprep.py

@@ -1,0 +1,342 @@
+#!/usr/bin/env python
+# (C) 2017 OpenEye Scientific Software Inc. All rights reserved.
+#
+# TERMS FOR USE OF SAMPLE CODE The software below ("Sample Code") is
+# provided to current licensees or subscribers of OpenEye products or
+# SaaS offerings (each a "Customer").
+# Customer is hereby permitted to use, copy, and modify the Sample Code,
+# subject to these terms. OpenEye claims no rights to Customer's
+# modifications. Modification of Sample Code is at Customer's sole and
+# exclusive risk. Sample Code may require Customer to have a then
+# current license or subscription to the applicable OpenEye offering.
+# THE SAMPLE CODE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND,
+# EXPRESS OR IMPLIED.  OPENEYE DISCLAIMS ALL WARRANTIES, INCLUDING, BUT
+# NOT LIMITED TO, WARRANTIES OF MERCHANTABILITY, FITNESS FOR A
+# PARTICULAR PURPOSE AND NONINFRINGEMENT. In no event shall OpenEye be
+# liable for any damages or liability in connection with the Sample Code
+# or its use.
+
+#############################################################################
+# prepare a molecule: alts, hydrogens, split ligand
+#############################################################################
+import sys
+from openeye import oechem
+
+
+def WaterProcess(processName):
+    if processName == "fullsearch":
+        return oechem.OEPlaceHydrogensWaterProcessing_FullSearch
+    elif processName == "focused":
+        return oechem.OEPlaceHydrogensWaterProcessing_Focused
+    return oechem.OEPlaceHydrogensWaterProcessing_Ignore
+
+
+def main(argv=[__name__]):
+
+    itf = oechem.OEInterface(InterfaceData)
+    oechem.OEConfigureSplitMolComplexOptions(itf,
+                                             oechem.OESplitMolComplexSetup_All &
+                                             ~ (oechem.OESplitMolComplexSetup_CovBondTreatment |
+                                                oechem.OESplitMolComplexSetup_CovCofactor))
+
+    if not oechem.OEParseCommandLine(itf, argv):
+        oechem.OEThrow.Fatal("Unable to interpret command line!")
+
[…]

### 2023-03-14 paper/manuscript.tex
Updates from Overleaf

--- 2023-03-08 manuscript.tex
@@
 \section{Methods}
-The selective iterative latent variable refinement (SILVR) method makes use of the pre-trained equivariant diffusion model (EDM) as introduced by Hegeboom et al.~\cite{}.
+The selective iterative latent variable refinement (SILVR) method makes use of the pre-trained equivariant diffusion model (EDM) as introduced by Hoogeboom \textit{et al.}~\cite{hoogeboom2022equivariant}.
 The core of the method is the introduction of an additional refinement within the denoising process during run time.
@@
 
-The EDM by Hoodgeboom et al was trained on the 30 lowest energy conformations of 430,000 molecules from the GEOM dataset with explicit protons.
-A simplified schematic describing EDM architecture is shared in scheme X for more details on the model theory and training refer to Hegeboom et al.~\cite{}.
+The EDM by Hoodgeboom \textit{et al.} was trained on the 30 lowest energy conformations of 430,000 molecules from the GEOM dataset with explicit protons.
+A simplified schematic describing EDM architecture is shared in scheme X for more details on the model theory and training refer to Hoogeboom et al.~\cite{hoogeboom2022equivariant}.
 This model strictly only considers atomic coordinates; all bond information is ignored.
@@
 The XYZ coordinates can then be interpreted using cheminformatics software to determine the molecular graph.
-As described, the EDM model is only trained to generate random molecules; hoodgeboom’s model does enable training given additional context, however this is outside the scope of this paper.
+As described, the EDM model is only trained to generate random molecules; Hoodgeboom’s model does enable training given additional context, however this is outside the scope of this paper.
 

### 2023-03-15 paper/manuscript.tex

@@
 \end{equation},
-for any $t> s$. The parameters $\alpha_t \in \mathbb{R}^+$ contains the amount of retained signal and $\sigma_t \in \mathbb{R}^+$ represents the variance and thus the amount of white noise added. 
+for any $t>s$. The parameters $\alpha_t \in \mathbb{R}^{+}$ contains the amount of retained signal and $\sigma_t \in \mathbb{R}^+$ represents the variance and thus the amount of white noise added. 
 Different people have looked at different noise schedules~\cite{sohl-dickstein2015deep, ho2020denoising}.
@@
 \subsubsection*{Euquivariant diffusion model}
-Unlike in image problems the orientation of molecules matters under translations and rotations.
+Unlike in image problems, the orientation of molecules matters under translations and rotations.
 For this purpose, we chose to equivariant diffusion model (EDM) by Hoogeboom et al.~\cite{} as our baseline generative model.
@@
 The selective iterative latent variable refinement (SILVR) method makes use of the pre-trained equivariant diffusion model (EDM) as introduced by Hoogeboom \textit{et al.}~\cite{hoogeboom2022equivariant}.
-The core of the method is the introduction of an additional refinement within the denoising process during run time.
-The resulting SILVR model can then produce conditional samples without any conditional training.
-In this work, we show how SILVR can be applied to fragment-based drug design to produce molecules resembling a reference set of atomic coordinates.
+The core of the new method is the addition of a refinement process within the denoising part during run time.
+The resulting SILVR model produces conditional samples without any conditional training when generating new molecules.
+To illustrate the usefulness of this run-time modification, we show how SILVR can be used in the context of fragment-based drug design.
+The goal is to produce molecules that resemble a binding site based on exiting fragments and their set of atomic coordinates.
 
+\subsection{Conditioning the pre-trained EDM with SILVR}
 The EDM by Hoodgeboom \textit{et al.} was trained on the 30 lowest energy conformations of 430,000 molecules from the GEOM dataset with explicit protons.
-A simplified schematic describing EDM architecture is shared in scheme X for more details on the model theory and training refer to Hoogeboom et al.~\cite{hoogeboom2022equivariant}.
+A simplified schematic describing the EDM architecture is shared in scheme~\ref{fig:_architecture} for more details on the model theory and training refer to Hoogeboom et al.~\cite{hoogeboom2022equivariant}.
 This model strictly only considers atomic coordinates; all bond information is ignored.
+A more recent version of the EDM also accounts for bond order and SILVR could also be incorporated into this new model.~\cite{hvignac2023midi}.
 During training, the atomic coordinates (and their element) are passed through a forward diffusion process with iterative addition of Gaussian noise, where the extent of noise added at each step is defined by parameter $\beta$ (N.B.
 scheme X shows the diffusion process as a Markov chain, however in practice the state at time $t$ is determined as a direct transformation of the initial state).
-The diffusion process is eventually terminated when $t=T$, by which point all structure is lost and all coordinates follow a gaussian distribution.
+The diffusion process is eventually terminated when $t=T$, by which point all structure is lost and all coordinates follow a Gaussian distribution.
 An equivariant graph neural network (EGNN) is then trained to predict the reverse process, predicting the previous state in the sequence given any state.
-At run time, the generative model is instantiated with a sample from a gaussian distribution and the series of denoising steps are applied resulting in a generated sample consisting of atomic coordinates resembling a drug-like molecule.
+At run time, the generative model is instantiated with a sample from a Gaussian distribution and the series of denoising steps are applied resulting in a generated sample consisting of atomic coordinates resembling a drug-like molecule.
 The XYZ coordinates can then be interpreted using cheminformatics software to determine the molecular graph.
-As described, the EDM model is only trained to generate random molecules; Hoodgeboom’s model does enable training given additional context, however this is outside the scope of this paper.
+
+The EDM was then adapted by introducing SILVR within the denoising process algorithm~\ref{alg:silvr}.
+At run time, each atom of the reference set of coordinates is mapped to an atom in the EDM.
+The reference coordinates are then translated such that their center of mass is at the origin.
+The reference coordinates are then diffused to the same timestep as that of the denoising process.
+That is, the amount of structure remaining from the reference should match the amount of structure formed by the generative process.
+A small refinement is applied to add information from the reference coordinates to the latent variable of the denoising process~\ref{eq:denoising}.
+This equation has the effect of shrinking the coordinates towards the origin, and then expanding the coordinates out in the direction of the reference.
+Importantly, the extent of this refinement is defined by the SILVR rate $r_{\mathrm{\{S}}$, with $r_{\mathrm{\{S}}=0$ providing no additional refinement and $r_{\mathrm{\{S}}=1$ will result in a total replacement of atoms.
+The diffusion of the reference to $t=\tau$ requires the sampling of a Gaussian; while in principle this sampling could be performed only once prior to the denoising loop, in practice better results were obtained when the Gaussian was repeatedly sampled during the denoising process.
+Once denoising is complete, sampled molecules can be translated back to the same coordinates as the reference by reverting the initial centre of gravity transformation; in the case of fragment data, this has the effect of returning samples to the binding site of the protein.
+
+By introducing iterative refinement steps, the unconditional EDM can be guided to sample from a smaller region of chemical space that resembles the reference set of coordinates.
+Figure~\ref{fig:} demonstrates this architecture with the example of three disconnected fragments.
+Here, the model generates a single 5-membered ring molecule with each atom maintaining the same element, however, notice each atom has drifted slightly from the reference.
+This is due to the competing effects of SILVR and EDM: the EDM tries to push atoms into a chemically reasonable position, while SILVR pulls atoms towards the reference.
+The resulting samples, therefore, therefore resemble both valid-looking molecules and the reference set of coordinates.
+The ability for reference atoms of fragments to move during generation distinguishes SILVR from standard linker design~\cite{linkerdesign1, hegeboom_linker}.
+
+\subsection{Reference Dataset: Covid-moonshot}
+Reference molecules were selected from the original 23 non-covalent hits of the SARS-CoV-2 main protease (Mpro) identified by the XChem fragment screen as part of the Covid Moonshot Project~\cite{moonshot}.
+A more detailed picture of all fragements is presented in Figure~\ref{si:figure}.
+Fragments were visualised using XXX v XXX and combinations were arbitrarily selected as test cases.
+Fragments \textit{x0072} and \textit{x0354} were selected for benchmarking the effect of $r_{\mathrm{S}}$ on sampling; \textit{x1093}, \textit{x0072}, \textit{x2193} were selected to represent three significantly overlapping fragments; \textit{x0434}, \textit{x0305} and \textit{x0072}, \textit{x0354} were selected as partially overlapping fragments; and \textit{x0874}, \textit{x0397} were selected as two disconnected fragments, resembling a linker design type problem.
+Bonding information of selected fragments was deleted and XYZ coordinates were combined into a single XYZ file.
+Values of $r_{\mathrm{S}}$ were selected and added to the XYZ file to create a reference file containing all experiment setup information.
+Each experiment was sampled 1000 times.
+
+\subsection{Diferent observables were monitor to gauge the success of computational experiment for the generation of new models}
+
+Different observables were used to monitor how realistic and reliable new molecules were generated and how well they can fit into the existing binding site of the SARS-CoV-2 main protease.
+
+\subsubsection{Atom stability}
+The accuracy of placement of atoms was determined using the stability metric proposed by Hoogeboom \textit{et al.}.
+This metric infers bonds and bond orders between atoms by considering their interatomic distances.
+Once all bonds are defined, the valance of each atom is compared to its expected valance and if these values match, the atom is determined as stable.
+It should be noted that this metric requires the presence of all hydrogens (both explicit and implicit) for an atom to be decided as stable.
+For comparability with other similar published models, this metric was used unmodified.
+The additional metric of “molecular stability” is often reported together with atom stability (if every atom is stable then the whole molecule is stable), however as has been previously identified, large molecules sampled from this GEOM trained EDM tend to be unstable.
+
+\subsubsection{SILVR to reference}
+The SILVR algorithm creates a one-to-one mapping between reference atom coordinates and sample atom coordinates.
+The RMSD for this pairwise mapping, ignoring atom identities, was calculated to determine the spacial similarity of samples to the reference.
+
+\subsubsection{Shape similarity of sample to protein}
+The agreement in the shape of samples and the main protease binding site was determined using OpenEye Tools Gaussian scoring function \textit{Shapegauss}.
+This scoring function measures the shape similarity between the ligand and receptor by considering each heavy atom as a Gaussian function~\cite{mcgann2003gaussian}.
+The most favourable score occurs “when two atoms touch but do not overlap”.
+This metric does not consider any intermolecular interactions.
+The protein receptor file was prepared from the Mpro-x0072 crystal structure with removal of the ligand.
+The XYZ coordinates of samples were directly read into OpenEye Tools and the pose was re-scored with \textit{shapegauss}.
+
+\subsection{Maybe: Docking of molecules}
 

### 2023-03-15 paper/manuscript.tex
Updates from Overleaf

@@
 - This allows the generation of new molecules in the shape of the binding site using information from existing fragments.
-
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{paper/Figures/Fig1/Schematic.png}
+    \caption{Schematic of the equivariant diffusion model with  SILVR indicated for every denoising step.}
+    \label{fig:schematic}
+\end{figure}
 \section{Theory}
@@
 \end{algorithm}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{Figures/minimal_model.png}
-    \caption{Minimal architecture figure - I prefer this style as it fits better with the theme}
-    \label{fig:fig_1}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{Figures/molecules_silvr_rates_horizontal.png}
-    \caption{This looks much better than I thought}
-    \label{fig:fig_2}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{Figures/silvr_plots.png}
-    \caption{Preliminary SILVR results (labels missing).
-A) Ratio of stable atoms by considering bond distances (note this includes protons but strictly maybe they should be removed.
-I would have to make new code for that I think).
-B) RMSD between samples and reference coordinated - pairwise mapping (samples rotated and translated to minimise RMSD).
-C)Openeye Shapegauss - still concerned I have run this wrong - lower value indicates better shape agreement with receptor D) Geometry stability - samples converted to smiles, geometry predicted using Auto3D, then RMSD with sample}
-    \label{fig:fig_3}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{Figures/fragment_sample_bound.png}
-    \caption{Examples of other reference sets and samples fitting binding site}
-    \label{fig:fig_5}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includesvg{Figures/silvr_model_svg.svg}
-    \caption{SVG of SILVR model - text not correctly aligned. Opinion vs previous figure.}
-    \label{fig:fig_6}
-\end{figure}
 
@@
 
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{paper/Figures/Fig3/fig_3_metrics.png}
+    \caption{Preliminary SILVR results (labels missing).
+A) Ratio of stable atoms by considering bond distances (note this includes protons but strictly maybe they should be removed.
+I would have to make new code for that I think).
+B) RMSD between samples and reference coordinated - pairwise mapping (samples rotated and translated to minimise RMSD).
+C)Openeye Shapegauss - still concerned I have run this wrong - lower value indicates better shape agreement with receptor D) Geometry stability - samples converted to smiles, geometry predicted using Auto3D, then RMSD with sample}
+    \label{fig:metrics}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{Figures/fragment_sample_bound.png}
+    \caption{Examples of other reference sets and samples fitting binding site}
+    \label{fig:samples}
+\end{figure}
+
 \section{Discussion and Conclusions}

### 2023-03-15 paper/SI.tex
Updates from Overleaf
substantial edits to results section

@@
 automatically add a sentence pointing to the information on-line:
+\section{Dataset}
+Here have a summary of all the fragment data
+
+\section{Samples from different SILVR rates}
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{paper/Figures/FigS2/fig_2_sample_vs_rate.png}
+    \caption{Two random samples from the same reference using different SILVR rates.
+All bonds were inferred from XYZ coordinates with openbabel.
+All bonds were visualised as single bonds and hydrogen atoms were deleted for clarity.
+Increasing SILVR rate results in sampled atom coordinates coming closer in space, and element type, to the reference while still resembling a true molecular structure.}
+    \label{fig:fig_2}
+\end{figure}
 

### 2023-03-15 paper/manuscript.tex
substantial edits to results section

@@
 \subsection{Reference Dataset: Covid-moonshot}
-Reference molecules were selected from the original 23 non-covalent hits of the SARS-CoV-2 main protease (Mpro) identified by the XChem fragment screen as part of the Covid Moonshot Project~\cite{moonshot}.
+Reference molecules were selected from the original 23 non-covalent hits of the SARS-CoV-2 Main Protease (Mpro) identified by the XChem fragment screen as part of the Covid Moonshot Project~\cite{moonshot}.
 A more detailed picture of all fragements is presented in Figure~\ref{si:figure}.
@@
 
-\subsection{Diferent observables were monitor to gauge the success of computational experiment for the generation of new models}
+\subsection{Diferent observables were monitored to gauge the success of computational experiment for the generation of new models}
 
@@
 
+\subsubsection{Geometric stability - Auto3D}
+To determine whether samples represent a true molecular geometry, independent minimisation of molecular geometries was performed using Atoms In Molecules Neural Network Potential (AIMNet) with Auto3D.
+All samples were read by RDKit and samples containing more than one molecule were removed from the test set.
+The SMILES string of each molecule was written to a new file, read by Auto3D, and geometry predicted by AIMNet.
+The RMSD between SILVR-generated coordinates and the Auto3D minimised coordinates were calculated with RDKit.
+
 \subsubsection{Shape similarity of sample to protein}
@@
 
-\subsection{Maybe: Docking of molecules}
+\subsubsection{Maybe: Docking of molecules}
 
 \section{Results}
+In the following, we will demonstrate how SILVR can be used to generate conditioned samples to a reference using a pre-trained EDM without additional training.
+Some of the flaws in the presented results stem from the underlying EDM, but retraining an improved EDM is out of the scope of this work.
+The main questions we set out to answer with SILVR were:
+\begin{enumerate}
+    \item Can we generate samples from the EDM that are similar to the reference structures?
+    \item Is there a SILVR rate $r_{\mathrm{S}}$ that will provide enough diversity while still retaining reference features?
+    \item Do the generated samples of new molecules still fit into the Mpro binding site?
+    \item Can we link molecule fragments without additional training?
+\end{enumerate}
+
+\subsection{The SILVR rate $r_{\mathrm{S}}$ effectively modulates similarity to the reference structures}
+Qualitatively, the generated molecular samples from the recondition EDM  show a clear resemblance to their reference structures, with similarity increasing with $r_{\mathrm{S}}$.
+Figure S2, in the SI, shows two example samples started from fragments xxx and xxx over a range of SILVR rates between $r_{\mathrm{S}}=0$ to $r_{\mathrm{S}}=0.02$.
+As expected at no conditioning random samples are generated that do not resemble the reference fragments.
+At low values of rs (< 0.0025) the sampled molecules only show an approximate agreement in orientation.
+At medium values of $r_{\mathrm{S}}$ ($0.0025 \le r_{\mathrm{S}} < 0.01$) the resulting samples begin to produce key structural features like ring systems and heteroatoms at positions seen in the reference.
+At high values of $r_{\mathrm{S}}$ ( $\ge 0.01$) the resulting samples have a very high resemblance to the reference with most structural features in correct positions, however, the diversity of samples is significantly reduced and structures start to become chemically less reasonable.
+At very high values of $r_{\mathrm{S}}$ ($> 0.02$) there is a very high similarity between samples and the reference, however, most structures no longer resemble valid molecules.
+It seems best molecules are formed at intermediate values of $r_{\mathrm{S}}$ ($0.05 < r_{\mathrm{S}} < 0.01$) offering a trade-off between similarity to the reference, sampling diversity, and molecular likeness.
+This is further validated by looking at stability measures.
+
+\subsection{Intermediate SILVR rates produce stable and varied molecules}
+To assess the stability and variability of generated molecules we looked at three different metrics, as discussed in the methods section~\ref{sec:methods}.
+We generated 1000 samples at different $r_{\mathrm{S}}$ from the two fragements \textit{X0072} and \textit{X0354} as a reference.
+Figure~\ref{fig:measures} summarises the findings from these experiments with box plots generated across the 1000 samples.
+The zeroth test we made with the generated samples was looking at how many generated molecules were fragemented i.e.
+are not a single connected molecular graph or not, with respect to $r_{\mathrm{S}}$.
+This is presented in Figure S4 in the SI.
+At an intermediate $r_{\mathrm{S}} = 0.025$ just over 50\% of the generated samples are not fragmented meaning that in the worst-case scenario, only one in two generated molecules can be analysed further.
+The subsequent analysis is carried out only on whole molecular graphs.
+Figure~\ref{fig:measures} a looks at the \textit{atom stability} measure as introduced by Hoogeboom et al.~\cite{hoogeboom2022equivariant}.
+Samples generated at low $r_{\mathrm{S}}$ tend to have similar stabilities to each other, samples start becoming less stable around $r_{\mathrm{S}}=0.005$, and become totally unstable at $r_{\mathrm{S}}=0.02$.
+This trend can largely be explained due to issues around protons.
+The atom stability measure calculates whether the valance of each atom matches what is expected for that atom, however, the measure requires the presence of explicit protons.
+A carbon skeleton with appropriate C-C bond lengths would be determined as \textit{unstable} unless each carbon was populated with explicit protons.
+In the case of high $r_{\mathrm{S}}$ values, the SILVR method pulls atom types strongly towards the reference.
+Since there are no protons in the reference, all atoms are mapped to heavy atoms, and therefore most atoms are unable to satisfy a full valence.
+Adding protons explicitly to the molecules, through openBable or RDkit, is a way of improving this measure.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=\textwidth]{paper/Figures/Fig3/fig_3_metrics.png}
+    \caption{Validation measures of SILVR model using fragments Mpro-x0072 and Mpro-x0354 as reference coordinates.
+a) Ratio of stable atoms - an atom is determine as stable if the valence matches the expected valence for the element b) RMSD from reference - the calculated RMSD between the reference and sample, using an absolute one-to-one mapping ignore atom identity c) OpenEye measure “shapegauss” - a gaussian scoring function describing the shape fit between Mpro and samples, ignoring chemical interactions d) Geometry stability - AIMNet geometry optimisation was completed with Auto3D using the SMILES string of each sample.
+RMSD was calculated between the predicted geometry and the sampled geometry using RDKit.}
+    \label{fig:measures}
+\end{figure}
+The similarity of samples to their reference set of coordinates was determined by RMSD, with a clear inverse correlation observed between rs and RMSD, as seen in Figure~\ref{fig:measures} b.
+This indicates that the extent of guidance of atoms towards the reference set of coordinates can be fine-tuned by varying $r_{\mathrm{S}}$.
+
+A successful 3D generative model should ideally produce ground state conformations.
+To determine whether the sampled molecular geometries were reasonable, a separate geometry optimisation protocol was applied to the SMILES strings of sampled molecules and RMSD calculated.
+At low to medium values of $r_{\mathrm{S}}$ ($< 0.01$), the average RMSD values all fall in the $1.6 - 1.7 \AA$ range, which is arguably good, however still leaves sufficient room for improvement.
+Importantly, no difference is seen between the control set ($r_{\mathrm{S}}=0$) and the SILVR samples ($0 < r_{\mathrm{S}} < 0.01$) indicating the quality of generated molecules was not hindered by the SILVR protocol.
+
+\subsection{Generated samples with SILVR fit the binding site of Mpro}
+
+The shape complementarity between sampled molecules and Mpro was calculated using the openeye shapegauss scoring function; a scoring function which only considers the shape of molecules and ignores any intermolecular interactions.
+As expected, the shape complementarity of samples improved with increasing rs, demonstrating that SILVR can produce ligands of binding site geometry when guided by the coordinates of fragment molecules.
+The lack of plot data for rs=2.0 and rs=3.0 was because all shapegauss calculations failed.
+We believe this was due to the atom coordinates representing highly stained and internally clashing molecules, and the scoring algorithm either failed to read the molecules or identified them as bad conformations.
+Scoring functions using intermolecular interactions, and molecular docking, were not attempted.
 
@@
     \centering
-    \includegraphics[width=\textwidth]{paper/Figures/Fig3/fig_3_metrics.png}
-    \caption{Preliminary SILVR results (labels missing).
-A) Ratio of stable atoms by considering bond distances (note this includes protons but strictly maybe they should be removed.
-I would have to make new code for that I think).
-B) RMSD between samples and reference coordinated - pairwise mapping (samples rotated and translated to minimise RMSD).
-C)Openeye Shapegauss - still concerned I have run this wrong - lower value indicates better shape agreement with receptor D) Geometry stability - samples converted to smiles, geometry predicted using Auto3D, then RMSD with sample}
-    \label{fig:metrics}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{Figures/fragment_sample_bound.png}
-    \caption{Examples of other reference sets and samples fitting binding site}
+    \includegraphics[width=\textwidth]{paper/Figures/Fig4/fig_4_ref_sample_bound.png}
+    \caption{Examples of generated molecules from different reference sets.
+Molecules were hand-curated.
+Figure depicts reference fragments used as input to SILVR (left), the sampled molecule (middle), and the sampled molecule translated to the protein binding site (right).
+Left set of samples (A,C,E) rs=0.005, right set of samples (B, D, F) rs=0.01.
+Row 1: Three significantly overlapping fragments A and B (x1093, x0072, x2193).
+Row 2: Two fragments partially overlapping C (x0434, x0305), D (x0072, x0354).
+Row 3: Two disconnected fragments E and F (x0874, x0397) samples generated including 10 dummy atoms (method described in SI)}
     \label{fig:samples}
@@
 
+\subsection{Fragements can be linked with SILVR and through the addition of dummy atoms}
+
 \section{Discussion and Conclusions}
+
+When applied to a drug design context, the conservation of key functional groups in exact spacial positions is crucial for maintenance of protein-ligand interactions.
+The series of molecules in figure X show a loss of the sulfonyl chloride group present in the reference, which may be undesirable.
+This issue can be solved by changing rs from a scalar to a vector, and by assigning particularly high rs values to selected atoms of the reference.
+Optimisation of rs vectors for drug design is scope for future work.
 

### 2023-03-15 paper/manuscript.tex
Finished first draft of Results

@@
 \usepackage{amsmath}
+\usepackage{hyperref}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 DDPMs consist of two main parts the diffusion process and the denoising process, see the schematic in figure~\ref{fig:diffusion_schematic}.
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{Figures/model_no_stickman.png}
-    \caption{This should show the basic idea of a diffusion model}
-    \label{fig:diffusion_schematic}
-\end{figure}
+
 The \textit{forward} part of the diffusion model is a Markov process, where you start out with a set of data $\mathbf{x}$ and add noise to this data over a time interval t = [0,\ldots,T] according to the following distribution:
@@
 This equation has the effect of shrinking the coordinates towards the origin, and then expanding the coordinates out in the direction of the reference.
-Importantly, the extent of this refinement is defined by the SILVR rate $r_{\mathrm{\{S}}$, with $r_{\mathrm{\{S}}=0$ providing no additional refinement and $r_{\mathrm{\{S}}=1$ will result in a total replacement of atoms.
+Importantly, the extent of this refinement is defined by the SILVR rate $r_{\mathrm{S}}$, with $r_{\mathrm{S}}=0$ providing no additional refinement and $r_{\mathrm{S}}=1$ will result in a total replacement of atoms.
 The diffusion of the reference to $t=\tau$ requires the sampling of a Gaussian; while in principle this sampling could be performed only once prior to the denoising loop, in practice better results were obtained when the Gaussian was repeatedly sampled during the denoising process.
@@
 Fragments were visualised using XXX v XXX and combinations were arbitrarily selected as test cases.
-Fragments \textit{x0072} and \textit{x0354} were selected for benchmarking the effect of $r_{\mathrm{S}}$ on sampling; \textit{x1093}, \textit{x0072}, \textit{x2193} were selected to represent three significantly overlapping fragments; \textit{x0434}, \textit{x0305} and \textit{x0072}, \textit{x0354} were selected as partially overlapping fragments; and \textit{x0874}, \textit{x0397} were selected as two disconnected fragments, resembling a linker design type problem.
-Bonding information of selected fragments was deleted and XYZ coordinates were combined into a single XYZ file.
+Fragments \texttt{x0072} and \texttt{x0354} were selected for benchmarking the effect of $r_{\mathrm{S}}$ on sampling; \texttt{x1093}, \texttt{x0072}, \texttt{x2193} were selected to represent three significantly overlapping fragments; \texttt{x0434}, \texttt{x0305} and \texttt{x0072}, \textit{x0354} were selected as partially overlapping fragments; and \texttt{x0874}, \texttt{x0397} were selected as two disconnected fragments, resembling a linker design type problem.
+The bonding information of selected fragments was deleted and XYZ coordinates were combined into a single XYZ file.
 Values of $r_{\mathrm{S}}$ were selected and added to the XYZ file to create a reference file containing all experiment setup information.
@@
 
-\subsection{Diferent observables were monitored to gauge the success of computational experiment for the generation of new models}
+\subsection{Diferent observables were monitored to gauge the success of computational experiments for the generation of new models}
 
@@
 For comparability with other similar published models, this metric was used unmodified.
-The additional metric of “molecular stability” is often reported together with atom stability (if every atom is stable then the whole molecule is stable), however as has been previously identified, large molecules sampled from this GEOM trained EDM tend to be unstable.
+The additional metric of “molecular stability” is often reported together with atom stability (if every atom is stable then the whole molecule is stable), however as has been previously identified, large molecules sampled from this GEOM-trained EDM tend to be unstable.
 
@@
 
+All data for the experiments carried and instruction on how to reproduce this work can be found at~\url{https://github.com/meyresearch/SILVR}.
+An updated version of the Hoogeboom et al.
+EDM that includes SILVR can be found at~\url{https://github.com/nichrun/e3_diffusion_for_molecules}.
+
 \section{Results}
@@
 As expected at no conditioning random samples are generated that do not resemble the reference fragments.
-At low values of rs (< 0.0025) the sampled molecules only show an approximate agreement in orientation.
+At low values of $r_{\mathrm{S}}$ (< 0.0025) the sampled molecules only show an approximate agreement in orientation.
 At medium values of $r_{\mathrm{S}}$ ($0.0025 \le r_{\mathrm{S}} < 0.01$) the resulting samples begin to produce key structural features like ring systems and heteroatoms at positions seen in the reference.
@@
 To assess the stability and variability of generated molecules we looked at three different metrics, as discussed in the methods section~\ref{sec:methods}.
-We generated 1000 samples at different $r_{\mathrm{S}}$ from the two fragements \textit{X0072} and \textit{X0354} as a reference.
+We generated 1000 samples at different $r_{\mathrm{S}}$ from the two fragments \textit{X0072} and \textit{X0354} as a reference.
 Figure~\ref{fig:measures} summarises the findings from these experiments with box plots generated across the 1000 samples.
-The zeroth test we made with the generated samples was looking at how many generated molecules were fragemented i.e.
+The zeroth test we made with the generated samples was looking at how many generated molecules were fragmented i.e.
 are not a single connected molecular graph or not, with respect to $r_{\mathrm{S}}$.
@@
 Since there are no protons in the reference, all atoms are mapped to heavy atoms, and therefore most atoms are unable to satisfy a full valence.
-Adding protons explicitly to the molecules, through openBable or RDkit, is a way of improving this measure.
+Adding protons explicitly to the molecules, through OpenBable or RDkit, is a way of improving this measure.
 
@@
     \caption{Validation measures of SILVR model using fragments Mpro-x0072 and Mpro-x0354 as reference coordinates.
-a) Ratio of stable atoms - an atom is determine as stable if the valence matches the expected valence for the element b) RMSD from reference - the calculated RMSD between the reference and sample, using an absolute one-to-one mapping ignore atom identity c) OpenEye measure “shapegauss” - a gaussian scoring function describing the shape fit between Mpro and samples, ignoring chemical interactions d) Geometry stability - AIMNet geometry optimisation was completed with Auto3D using the SMILES string of each sample.
+a) Ratio of stable atoms - an atom is determined as stable if the valence matches the expected valence for the element b) RMSD from reference - the calculated RMSD between the reference and sample, using an absolute one-to-one mapping ignore atom identity c) OpenEye measure “shapegauss” - a gaussian scoring function describing the shape fit between Mpro and samples, ignoring chemical interactions d) Geometry stability - AIMNet geometry optimisation was completed with Auto3D using the SMILES string of each sample.
 RMSD was calculated between the predicted geometry and the sampled geometry using RDKit.}
@@
 
-A successful 3D generative model should ideally produce ground state conformations.
-To determine whether the sampled molecular geometries were reasonable, a separate geometry optimisation protocol was applied to the SMILES strings of sampled molecules and RMSD calculated.
-At low to medium values of $r_{\mathrm{S}}$ ($< 0.01$), the average RMSD values all fall in the $1.6 - 1.7 \AA$ range, which is arguably good, however still leaves sufficient room for improvement.
-Importantly, no difference is seen between the control set ($r_{\mathrm{S}}=0$) and the SILVR samples ($0 < r_{\mathrm{S}} < 0.01$) indicating the quality of generated molecules was not hindered by the SILVR protocol.
+The next test we carried out was to determine whether the sampled molecular geometries are reasonable.
+For this purpose a separate geometry optimisation protocol was devised using the SMILES strings of the generated molecules and the RMSD between the generated molecule and the geometrically optimised molecule calculated.
+The results of this are found in Figure~\ref{fig:measures} d.
+At low to medium values of $r_{\mathrm{S}}$ ($< 0.01$), the average RMSD values all fall in the $1.6 - 1.7$ \AA~  range, which is arguably good, however still leaves sufficient room for improvement.
+Importantly, no difference is seen between the control set ($r_{\mathrm{S}}=0$) and the SILVR samples ($0 < r_{\mathrm{S}} < 0.01$) indicating the quality of generated molecules was not impeded by the SILVR protocol.
 
 \subsection{Generated samples with SILVR fit the binding site of Mpro}
-
-The shape complementarity between sampled molecules and Mpro was calculated using the openeye shapegauss scoring function; a scoring function which only considers the shape of molecules and ignores any intermolecular interactions.
-As expected, the shape complementarity of samples improved with increasing rs, demonstrating that SILVR can produce ligands of binding site geometry when guided by the coordinates of fragment molecules.
-The lack of plot data for rs=2.0 and rs=3.0 was because all shapegauss calculations failed.
-We believe this was due to the atom coordinates representing highly stained and internally clashing molecules, and the scoring algorithm either failed to read the molecules or identified them as bad conformations.
-Scoring functions using intermolecular interactions, and molecular docking, were not attempted.
+As one of the main motivators for SILVR is to be able to generate new molecules that fit directly into a binding site based on input fragements, we measured shape complementarity between newly generated samples and the Mpro binding site.
+We used the OpenEye shapegauss scoring function for this purpose.
+From Figure~\ref{fig:measures} c it can be seen that the shape complementarity of samples improves with increasing $r_{\mathrm{S}}$, demonstrating that SILVR can produce ligands of binding site geometry when guided by the coordinates of fragment molecules.
+The lack of plot data for $r_{\mathrm{S}}=0.02$ and $r_{\mathrm{S}}=0.03$ was because all shapegauss calculations failed.
+We believe this was due to the atom coordinates representing highly strained and internally clashing molecules, and the scoring algorithm either failed to read the molecules or identified them as bad conformations.
 
@@
 
+In addition to the experiments using two fragments and generating 1000 samples, we also looked at different combinations of fragments and resulting molecules.
+In general, the trends of Figure~\ref{fig:measures} were preserved for all experiments with slight overall variations.
+Here we present three cases we investigated in detail.
+\begin{enumerate}
+    \item Using three fragements with substantial overlap as a reference.
+    \item Using two fragements with a slight overlap. 
+    \item Using two fragements that are disconnected to investigate linker generation. This scenario will be presented in detail in section~\ref{sec:linker}.
+\end{enumerate}
+
+Figure~\ref{fig:samples} A-F shows hand-curated samples for the 3 different scenarios for $r_{\mathrm{S}}=0.005$ (A,C,E) and $r_{\mathrm{S}}=0.01$ (B,D,F).
+
+Figures~\ref{fig:samples}A and~\ref{fig:samples}B demonstrate a superposition of three significantly overlapping fragments that result in generated molecules that fit the Mpro binding site well.
+Scrutinising sample A with $r_{\mathrm{S}}=0.005$, we can see the azaindole fused ring system has been interpreted as a pyrrole ring, the ketone transformed into an amide (maintaining the same carbonyl position), the sulfonyl group vanished, and the overlapping atoms have transformed into a fused ring system.
+As a whole, the general geometry of the sample reflects the reference, however, functional groups are only weakly preserved.
+In contrast, sample B presents the same reference set but with $r_{\mathrm{S}}=0.01$.
+This new sample maintains the same geometry but better preserves key functional groups: the fused ring system is the same size, and satisfyingly the carbonyl oxygen has merged with the sulfonyl group to form a cyclic sulfamate ester.
+
+Figure~\ref{fig:samples}C shows a merger of two partially overlapping fragments with $r_{\mathrm{S}}=0.005$.
+While the urea group was successfully preserved, the 6-membered ring shrank to a 5-membered heterocycle.
+Of particular interest is the formation of the fused ring system.
+At first glance, it might be assumed that reference atoms map to the sample atom closest in space, however in actuality they travel up to 1.7 \AA~to arrive at their final position (SI figure 4).
+In this case, the nitrogen atoms observed in the fused ring are directly obtained from the nitrogen atoms in the reference, however, their final position is one bond’s length from their reference.
+This shows the flexibility of each sample atom to explore within a radius (defined by $r_{\mathrm{S}}$) of the reference atom.
+The fact that the sample molecule populates a similar region of space to the reference is the result of the aggregate effects of each mapping, as opposed to the strict fixation of each atom.
+
+In contrast, Figure~\ref{fig:samples}D shows a stricter merging of two fragments, with $r_{\mathrm{S}}=0.01$.
+Visibly, the scaffold of the lower fragment has been maintained while the top fragment has contributed to a fused ring.
+Interestingly, the sulfonamide and carbonyl (from opposing fragments) have merged to form N-oxazinane sulfonyl chloride, demonstrating a particularly creative result from SILVR.
+
 \subsection{Fragements can be linked with SILVR and through the addition of dummy atoms}
+\label{sec:linker}
+Being able to reliably link fragments that sit in a binding site of a protein is crucial for the design of potential new drugs.
+Being able to do so without retraining a model as was done by~\cite{huang20223dlinker}.
+Using conditioning through SILVR allows the generation of linkers between fragments without the need to retrain the EDM.
+This is illustrated in the example of Figure~\ref{fig:samples}E and F.
+While it was possible to use SILVR as described in the Theory section~\ref{sec:tehory}, better results were obtained with the addition of \textit{dummy atoms}.
+These are atoms which are present in the EDM without a mapping to a reference atom, and so are free to explore the whole coordinate space without guidance from SILVR.
+The successful implementation of dummy atoms requires a slight modification of the SILVR algirhtm, and is outlined in the SI.
+The results of these experiments continue the same trends previously observed where the $r_{\mathrm{S}}=0.005$ produces samples of approximate geometric similarity, whereas $r_{\mathrm{S}}=0.01$ produces a more strict mapping, with a clearly preserved urea group, a slightly modified ring system, and an amide interpreted as a carboxylate.
 
-\section{Discussion and Conclusions}
+\section{Discussion and Outlook}
+\label{sec:discussion}
+Quick summary of what we have done
+
+Considering all results with respect to the control EDM (rs=0.0), it can be seen that at intermediate values of rs the SILVR protocol produces molecules of equal quality to that of the unmodified EDM, while also being conditionally generated.
+We therefore claim that if a diffusion model can be successfully trained to produces random high quality drug like structures, SILVR should be applicable to guide generation towards desired regions of chemical space without harming the quality of molecules.
+While our method poses a direct interface between crystallographic fragment data and de novo molecular generation, we believe future work on SILVR could be expanded further.
+To answer structure based drug design, a scoring function of protein surface potentials could be used to guide atoms towards energetically favoured pocket regions; issues of synthetic accessibility might be solved by clever design of refinement methods; and interfacing with alternative data sources (such as NMR) might be achievable.
+
+Issues:
+- Poor quality of samples and fragmentation
+- Synthetic accessibility of the generated molecules
+- No guaranteed conservation of functional groups in the current scheme. 
+- No information incorporated from the binding site in terms of electrostatics
+
+The combined set of case studies demonstrates the ability of SILVR to complete both fragment merging and linking type problems, without any a priori knowledge of these design challenges.
+It should be noted however that the samples generated by SILVR are often of poorer quality than the samples selected in Figure X.
+Across all samples around half were determined by RDKit to be fragmented, meaning the sample contained two or more distinct molecular graphs.
+It was observed qualitatively that fragmented samples typically contained corrupted structures (multiple fragmentations, linear carbon chains, flattened rings, etc).
+We believe this fragmentation is triggered during intermediate steps of denoising, resulting in an unstable latent representation and subsequently poor EDM inference.
+Fragmentation becomes of particular issue for linker design type SILVR experiments (figure XE and F), where the reference coordinates direct the latent variables away from eachother, triggering fragmentation.
+For these experiments 65\% of all samples were fragmented.
+Further work is needed both with EDM and with SILVR to reduce these rates of fragmentation.
+Additionally, as is common for generative models, the synthetic accessibility of most sampled molecules is very poor.
+For our experiments the synthetic accessibility of SILVR generated molecules resembles the performance of the unmodified EDM, so in order to achieve synthetically accessible samples with SILVR an improved EDM will need to be designed.
 
 When applied to a drug design context, the conservation of key functional groups in exact spacial positions is crucial for maintenance of protein-ligand interactions.
-The series of molecules in figure X show a loss of the sulfonyl chloride group present in the reference, which may be undesirable.
-This issue can be solved by changing rs from a scalar to a vector, and by assigning particularly high rs values to selected atoms of the reference.
-Optimisation of rs vectors for drug design is scope for future work.
+The series of molecules in figure S2 shows a loss of the sulfonyl chloride group present in the reference, which may be undesirable.
+This issue can be solved by changing $r_{\mathrm{S}}$ from a scalar to a vector ($\mathbf{r}_{\mathrm{S}}$), and by assigning particularly high $\mathbf{r}_{\mathrm{S}}$ values to selected atoms of the reference.
+Optimisation of $\mathbf{r}_{\mathrm{S}}$ vectors for actual drug design applications may become viable with a more suitably trained EDM.
+
+\section{Conclusions}
 
@@
 \begin{acknowledgement}
-The authors thank \ldots.
-
+The authors thank Matteo T. Degiacomi and John D. Chodera for useful discussions and feedback on the manuscript. 
 \end{acknowledgement}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\bibliography{achemso-demo}
+\bibliography{SILVR}
 

### 2023-03-15 paper/manuscript.tex

@@
 Considering all results with respect to the control EDM (rs=0.0), it can be seen that at intermediate values of rs the SILVR protocol produces molecules of equal quality to that of the unmodified EDM, while also being conditionally generated.
-We therefore claim that if a diffusion model can be successfully trained to produces random high quality drug like structures, SILVR should be applicable to guide generation towards desired regions of chemical space without harming the quality of molecules.
-While our method poses a direct interface between crystallographic fragment data and de novo molecular generation, we believe future work on SILVR could be expanded further.
-To answer structure based drug design, a scoring function of protein surface potentials could be used to guide atoms towards energetically favoured pocket regions; issues of synthetic accessibility might be solved by clever design of refinement methods; and interfacing with alternative data sources (such as NMR) might be achievable.
+We, therefore, claim that if a diffusion model can be successfully trained to produce random high-quality drug like structures, SILVR should be applicable to guide generation towards desired regions of chemical space without harming the quality of molecules.
+While our method poses a direct interface between crystallographic fragment data and \textit{de novo} molecular generation, we believe future work on SILVR could be expanded further.
+To answer structure-based drug design, a scoring function of protein surface potentials could be used to guide atoms towards energetically favoured pocket regions; issues of synthetic accessibility might be solved by clever design of refinement methods; and interfacing with alternative data sources (such as NMR) might be achievable.
 
@@
 
-The combined set of case studies demonstrates the ability of SILVR to complete both fragment merging and linking type problems, without any a priori knowledge of these design challenges.
+The combined set of case studies demonstrates the ability of SILVR to complete both fragments merging and linking type problems, without any a priori knowledge of these design challenges.
 It should be noted however that the samples generated by SILVR are often of poorer quality than the samples selected in Figure X.
-Across all samples around half were determined by RDKit to be fragmented, meaning the sample contained two or more distinct molecular graphs.
+Across all samples around half was determined by RDKit to be fragmented, meaning the sample contained two or more distinct molecular graphs.
 It was observed qualitatively that fragmented samples typically contained corrupted structures (multiple fragmentations, linear carbon chains, flattened rings, etc).
 We believe this fragmentation is triggered during intermediate steps of denoising, resulting in an unstable latent representation and subsequently poor EDM inference.
-Fragmentation becomes of particular issue for linker design type SILVR experiments (figure XE and F), where the reference coordinates direct the latent variables away from eachother, triggering fragmentation.
-For these experiments 65\% of all samples were fragmented.
+Fragmentation becomes of particular issue for linker design type SILVR experiments (figure XE and F), where the reference coordinates direct the latent variables away from each other, triggering fragmentation.
+For these experiments, 65\% of all samples were fragmented.
 Further work is needed both with EDM and with SILVR to reduce these rates of fragmentation.
 Additionally, as is common for generative models, the synthetic accessibility of most sampled molecules is very poor.
-For our experiments the synthetic accessibility of SILVR generated molecules resembles the performance of the unmodified EDM, so in order to achieve synthetically accessible samples with SILVR an improved EDM will need to be designed.
+For our experiments, the synthetic accessibility of SILVR-generated molecules resembles the performance of the unmodified EDM, so in order to achieve synthetically accessible samples with SILVR an improved EDM will need to be designed.
 

### 2023-03-15 paper/manuscript.tex
First draft discussion

@@
   {SILVR: Conditional diffusion model for molecule generation without additional training}
-
+\title[SILVR: Molecular Generation for binding modes]
+  {SILVR: Guided molecular generation from diffusion models without additional training}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \section{Introduction}
+\label{sec:intro}
 - Equivariant diffusion models have been very promising in predicting 3D structures of drug-like molecules. 
@@
 \section{Theory}
+\label{sec:theory}
 \subsection{Unconditional Equivariant Denoising Diffusion probabilistic models}
@@
 \begin{enumerate}
-    \item Using three fragements with substantial overlap as a reference.
-    \item Using two fragements with a slight overlap. 
-    \item Using two fragements that are disconnected to investigate linker generation. This scenario will be presented in detail in section~\ref{sec:linker}.
+    \item Using three fragments with substantial overlap as a reference.
+    \item Using two fragments with a slight overlap. 
+    \item Using two fragments that are disconnected to investigate linker generation. This scenario will be presented in detail in section~\ref{sec:linker}.
 \end{enumerate}
@@
 
-\subsection{Fragements can be linked with SILVR and through the addition of dummy atoms}
+\subsection{Fragements can be linked using SILVR and additional dummy atoms}
 \label{sec:linker}
@@
 These are atoms which are present in the EDM without a mapping to a reference atom, and so are free to explore the whole coordinate space without guidance from SILVR.
-The successful implementation of dummy atoms requires a slight modification of the SILVR algirhtm, and is outlined in the SI.
+The successful implementation of dummy atoms requires a slight modification of the SILVR algorithm, and is outlined in the SI.
 The results of these experiments continue the same trends previously observed where the $r_{\mathrm{S}}=0.005$ produces samples of approximate geometric similarity, whereas $r_{\mathrm{S}}=0.01$ produces a more strict mapping, with a clearly preserved urea group, a slightly modified ring system, and an amide interpreted as a carboxylate.
@@
 \label{sec:discussion}
-Quick summary of what we have done
+SILVR, as presented, represents a promising proof of concept method in which a general EDM can be conditioned to generate samples that resemble a reference structure without any additional training needed.
+We could show that SILVR can complete both fragment merging and linking type tasks, without any \textit{a priori} knowledge of these design challenges.
+Considering all results with respect to the control EDM ($r_{\mathrm{S}}=0.0$), we show that at intermediate values of $r_{\mathrm{S}}$ the SILVR protocol produces molecules of equal quality to that of the unmodified EDM, while also guiding molecules towards reference structures.
+We, therefore, claim that if a diffusion model can be successfully trained to produce random high-quality drug-like structures, SILVR will provide molecular designs towards desired regions of chemical space without harming the quality of molecules.
+Our method poses a direct interface between crystallographic fragment data and \textit{de novo} molecular generation.
+However, there are a few ways in which the current method can be improved further but we deem out of scope for this work.
 
-Considering all results with respect to the control EDM (rs=0.0), it can be seen that at intermediate values of rs the SILVR protocol produces molecules of equal quality to that of the unmodified EDM, while also being conditionally generated.
-We, therefore, claim that if a diffusion model can be successfully trained to produce random high-quality drug like structures, SILVR should be applicable to guide generation towards desired regions of chemical space without harming the quality of molecules.
-While our method poses a direct interface between crystallographic fragment data and \textit{de novo} molecular generation, we believe future work on SILVR could be expanded further.
-To answer structure-based drug design, a scoring function of protein surface potentials could be used to guide atoms towards energetically favoured pocket regions; issues of synthetic accessibility might be solved by clever design of refinement methods; and interfacing with alternative data sources (such as NMR) might be achievable.
-
-Issues:
-- Poor quality of samples and fragmentation
-- Synthetic accessibility of the generated molecules
-- No guaranteed conservation of functional groups in the current scheme. 
-- No information incorporated from the binding site in terms of electrostatics
-
-The combined set of case studies demonstrates the ability of SILVR to complete both fragments merging and linking type problems, without any a priori knowledge of these design challenges.
-It should be noted however that the samples generated by SILVR are often of poorer quality than the samples selected in Figure X.
-Across all samples around half was determined by RDKit to be fragmented, meaning the sample contained two or more distinct molecular graphs.
+\subsection{The number of unfragmented molecules generated can be improved}
+The samples generated by SILVR are often of poorer quality than the samples selected in Figure~\ref{fig:samples}.
+Across all samples around half of the samples were determined by RDKit to be fragmented, meaning the sample contained two or more distinct molecular graphs.
 It was observed qualitatively that fragmented samples typically contained corrupted structures (multiple fragmentations, linear carbon chains, flattened rings, etc).
 We believe this fragmentation is triggered during intermediate steps of denoising, resulting in an unstable latent representation and subsequently poor EDM inference.
-Fragmentation becomes of particular issue for linker design type SILVR experiments (figure XE and F), where the reference coordinates direct the latent variables away from each other, triggering fragmentation.
+Fragmentation becomes a particular issue for linker design type SILVR tasks (Figure~\ref{fig:samples}E and F), where the reference coordinates direct the latent variables away from each other, triggering fragmentation.
 For these experiments, 65\% of all samples were fragmented.
 Further work is needed both with EDM and with SILVR to reduce these rates of fragmentation.
-Additionally, as is common for generative models, the synthetic accessibility of most sampled molecules is very poor.
-For our experiments, the synthetic accessibility of SILVR-generated molecules resembles the performance of the unmodified EDM, so in order to achieve synthetically accessible samples with SILVR an improved EDM will need to be designed.
 
-When applied to a drug design context, the conservation of key functional groups in exact spacial positions is crucial for maintenance of protein-ligand interactions.
-The series of molecules in figure S2 shows a loss of the sulfonyl chloride group present in the reference, which may be undesirable.
+\subsection{The synthetic accessibility of the underlying EDM has a direct impact on the generated molecules}
+ For our experiments, the synthetic accessibility of SILVR-generated molecules resembles the performance of the unmodified EDM.
+In order to achieve synthetically accessible samples with SILVR an improved EDM will need to be designed.
+An improved version of the EDM we have used has recently been proposed using more explicit information on bond order and represents the next appropriate step for testing SILVR~\cite{vignac2023midi}.
+
+\subsection{There is no control over the retention of functional groups from the reference structure}
+When applied to a drug design context, the conservation of key functional groups in exact spacial positions is crucial for the maintenance of protein-ligand interactions.
+The series of molecules in Figure S2 shows a loss of the sulfonyl chloride group present in the reference, which may be undesirable.
 This issue can be solved by changing $r_{\mathrm{S}}$ from a scalar to a vector ($\mathbf{r}_{\mathrm{S}}$), and by assigning particularly high $\mathbf{r}_{\mathrm{S}}$ values to selected atoms of the reference.
 Optimisation of $\mathbf{r}_{\mathrm{S}}$ vectors for actual drug design applications may become viable with a more suitably trained EDM.
+
+\subsection{The inclusion of electrostatic information from proteins will likely improve the newly generated models}
+To answer structure-based drug design, a scoring function of protein surface potentials could be used to guide atoms towards energetically favoured pocket regions; issues of synthetic accessibility might be solved by clever design of refinement methods; and interfacing with alternative data sources (such as NMR) might be achievable.
 

### 2023-03-15 paper/manuscript.tex
first draft of conclusion

@@
 
-All data for the experiments carried and instruction on how to reproduce this work can be found at~\url{https://github.com/meyresearch/SILVR}.
-An updated version of the Hoogeboom et al.
-EDM that includes SILVR can be found at~\url{https://github.com/nichrun/e3_diffusion_for_molecules}.
-
 \section{Results}
@@
 \section{Conclusions}
+We developed SILVR, a method that can be injected into a pre-trained equivariant diffusion model that serves as a molecular generator to explore new chemical space.
+SILVR allows the conditioning of molecules based on a reference set of molecules, e.g.
+a fragment from an X-ray fragment screen.
+The SILVR rate $r_{\mathrm{S}}$ allows the tuning of 'how much of the reference' molecule should be taken into account when generating new molecules, with medium values of $r_{\mathrm{S}}$ around 0.005 giving the best results.
+The simple conditioning against a reference set of molecules means that the model can be used for tasks of fragment linker design, as well as generating new molecules that fit into an existing binding pocket without any specific training needed towards these tasks.
+In the future, given improvements in EDMs that produce more realistic and synthetically accessible molecules this method can cheaply generate structures exploring new chemical space with desired conditioning towards existing fragment hits.
+
+\section{Data Availability}
+All data for the experiments carried out and instructions on how to reproduce this work can be found at~\url{https://github.com/meyresearch/SILVR}.
+An updated version of Hoogeboom et al.
+EDM that includes SILVR can be found at~\url{https://github.com/nichrun/e3_diffusion_for_molecules}.
 

### 2023-03-16 paper/manuscript.tex

@@
 More recently the same idea has also been applied to molecular generation~\cite{}.
-The main idea behind diffusion models is for a neural network to learn the reverse of a diffusion process, often referred to as denoising to \textit{sample} a new image or new molecule.
+The main idea behind diffusion models is for a neural network to learn the reverse of a diffusion process, often referred to as denoising to \textit{sample} a new image or in our case a new molecule.
 In practice this is done by training a neural network $\phi$ and generating samples $\hat{\mathbf{x}}=\psi(X_t,t)$, with $X_t$ with data $X$, noised up to time $t$.
-DDPMs consist of two main parts the diffusion process and the denoising process, see the schematic in figure~\ref{fig:diffusion_schematic}.
+Figure~\ref{fig:diffusion_schematic} shows a schematic of the two main parts of this model, the diffusion part where noise is added for each timestep (shown in blue) and the denoising part (shown in yellow) that allows the generation of a new image (molecule).
 
@@
 \subsubsection*{Euquivariant diffusion model}
-Unlike in image problems, the orientation of molecules matters under translations and rotations.
-For this purpose, we chose to equivariant diffusion model (EDM) by Hoogeboom et al.~\cite{} as our baseline generative model.
+Unlike in image problems, the orientation of molecules matter under translations and rotations.
+For this purpose, equivariant diffusion models (EDMs) are used to account for rotational invarance.
+For this purpose we chose to equivariant diffusion model (EDM) by Hoogeboom et al.~\cite{hoogeboom2022equivariant} as our baseline generative model, as it provides a generative model for new molecules and has equivariance already built into it.
 The basic concept behind equivarience is that the model is invariant to rotations and translation in this case the E(3) group.
 This means that scalar (features such as atom types) and vector node properties (such as the positions) will be invariant to group transformations.
-This means that the order in which a rotation is applied does not matter.
+As a result the order in which a rotation is applied does not matter.
 You can rotate the input to the model then apply the diffusion/denoising process and get a structure out, or you can apply the diffusion/denoising process first and then do the same rotation to get the same output.
 Mathematically this means that if we have a set of point $\mathbf{x} = (\mathbf{x}_1,\ldots,\mathbf{x}_N) \in \mathbb{^{N\times 3}}$ and each of these points has a set of scalar feature vectors $\mathbf{h}\in \mathbb{R}^{N\times k}$ associated with it these features are invariant to group transformations.
-The positions translations and rotations are defined according to the orthogonal matrix: $\mathbf{Rx + t} = (\mathbf{R_{x_1}+t},\ldots, \mathbf{R_{x_N}+t}$.
+The positions translations and rotations are defined according to the orthogonal matrix: $\mathbf{Rx + t} = (\mathbf{R_{x_1}+t},\ldots, \mathbf{R_{x_N}+t})$.
+Satorras et al.~\cite{} have proposed an E(n) Equivariant Graph Neural Network on which the EDM by Hoogeboom et al.
+builds.
+For full details on the architecture and their code see~\cite{hoogeboom2022equivariant}.
 %%%%
@@
 \subsection{Selective Iterative Latent Variable Refinement}
-
+We propose a new method, that combines the iterative variable refinement (ILVR) by Choi et al.~\cite{choi2021ilvr} and take it from the image generation process and incorporate it into the EDM by Hoogeboom et al.
+to produce a way to generate a selective iterative laten variable refinement (SILVR) procedure in which we use a reference molecule to introduce into the denoising step.
+The reference used can be a small fragment, multiple fragments or even include dummy atoms that do not exist in the original reference.
+As a result we propose a modification of Algorithm 2 by Hoogeboom et al.~\cite{hoogeboom2022equivariant} to result in the SILVR algorithm~\ref{alg:silvr}
 \begin{algorithm}
-\caption{SILVR}\label{alg:cap}
+\caption{SILVR}\label{alg:silvr}
+\begin{flushleft}
+\textbf{Input:} reference molecule $\mathbf{y,h}$, neural network $\phi$ \\
+\textbf{Output:} generated molecule $\mathbf{x,h}$
+\end{flushleft}
 \begin{algorithmic}
-\State Sample $x_T \sim \mathcal{N}(0,I)$
+\State Align to centre of gravity
+\State Sample $\hat{x}_T \sim \mathcal{N}(0,I)$
 \For {$t$ = $T,...,1$}
@@
 
-\section{Methods}
-The selective iterative latent variable refinement (SILVR) method makes use of the pre-trained equivariant diffusion model (EDM) as introduced by Hoogeboom \textit{et al.}~\cite{hoogeboom2022equivariant}.
+\begin{figure}
+    \centering
+    \includegraphics{}
+    \caption{Caption}
+    \label{fig:my_label}
+\end{figure}
+
 The core of the new method is the addition of a refinement process within the denoising part during run time.
@@
 The goal is to produce molecules that resemble a binding site based on exiting fragments and their set of atomic coordinates.
+
+\section{Methods}
+As stated before, the selective iterative latent variable refinement (SILVR) method makes use of the pre-trained equivariant diffusion model (EDM) as introduced by Hoogeboom \textit{et al.}~\cite{hoogeboom2022equivariant}.
 
@@
     \includegraphics[width=\textwidth]{paper/Figures/Fig3/fig_3_metrics.png}
-    \caption{Validation measures of SILVR model using fragments Mpro-x0072 and Mpro-x0354 as reference coordinates.
-a) Ratio of stable atoms - an atom is determined as stable if the valence matches the expected valence for the element b) RMSD from reference - the calculated RMSD between the reference and sample, using an absolute one-to-one mapping ignore atom identity c) OpenEye measure “shapegauss” - a gaussian scoring function describing the shape fit between Mpro and samples, ignoring chemical interactions d) Geometry stability - AIMNet geometry optimisation was completed with Auto3D using the SMILES string of each sample.
+    \caption{Validation measures of SILVR model using fragments \texttt{x0072} and \texttt{x0354} as reference coordinates.
+A) Ratio of stable atoms - an atom is determined as stable if the valence matches the expected valence for the element B) RMSD from reference - the calculated RMSD between the reference and sample, using an absolute one-to-one mapping ignore atom identity C) OpenEye measure “shapegauss” - a Gaussian scoring function describing the shape fit between Mpro and samples, ignoring chemical interactions D) Geometry stability - AIMNet geometry optimisation was completed with Auto3D using the SMILES string of each sample.
 RMSD was calculated between the predicted geometry and the sampled geometry using RDKit.}
@@
     \includegraphics[width=\textwidth]{paper/Figures/Fig4/fig_4_ref_sample_bound.png}
-    \caption{Examples of generated molecules from different reference sets.
-Molecules were hand-curated.
-Figure depicts reference fragments used as input to SILVR (left), the sampled molecule (middle), and the sampled molecule translated to the protein binding site (right).
-Left set of samples (A,C,E) rs=0.005, right set of samples (B, D, F) rs=0.01.
-Row 1: Three significantly overlapping fragments A and B (x1093, x0072, x2193).
-Row 2: Two fragments partially overlapping C (x0434, x0305), D (x0072, x0354).
-Row 3: Two disconnected fragments E and F (x0874, x0397) samples generated including 10 dummy atoms (method described in SI)}
+    \caption{Examples of generated molecules from different experiments testing different overlap models.
+The reference fragments used as input to SILVR  are shown in the left column, the sampled molecule in the middle column, and the sampled molecule translated to the protein binding site in the right column.
+The left set of samples (A,C,E) has a SILVR rate $r_{\mathrm{S}}=0.005$, and the right set of samples (B, D, F) $r_{\mathrm{S}}=0.01$.
+Row 1: Three significantly overlapping fragments A and B (\texttt{x1093, x0072, x2193}).
+Row 2: Two fragments partially overlapping C (\texttt{x0434, x0305}), D (\texttt{x0072, x0354}).
+Row 3: Two disconnected fragments E and F (\texttt{x0874, x0397}) samples generated including 10 dummy atoms (method described in SI).
+The selection of molecules was hand-curated.}
     \label{fig:samples}
@@
 
-\subsection{Fragements can be linked using SILVR and additional dummy atoms}
+\subsection{Fragments can be linked using SILVR and additional dummy atoms}
 \label{sec:linker}
@@
 These are atoms which are present in the EDM without a mapping to a reference atom, and so are free to explore the whole coordinate space without guidance from SILVR.
-The successful implementation of dummy atoms requires a slight modification of the SILVR algorithm, and is outlined in the SI.
+The successful implementation of dummy atoms requires a slight modification of the SILVR algorithm and is outlined in the SI.
 The results of these experiments continue the same trends previously observed where the $r_{\mathrm{S}}=0.005$ produces samples of approximate geometric similarity, whereas $r_{\mathrm{S}}=0.01$ produces a more strict mapping, with a clearly preserved urea group, a slightly modified ring system, and an amide interpreted as a carboxylate.

### 2023-03-16 paper/manuscript.tex

@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%https://www.overleaf.com/project/63f3627b7f8554c152639004
 %% If issues arise when submitting your manuscript, you may want to
@@
 The reference used can be a small fragment, multiple fragments or even include dummy atoms that do not exist in the original reference.
-As a result we propose a modification of Algorithm 2 by Hoogeboom et al.~\cite{hoogeboom2022equivariant} to result in the SILVR algorithm~\ref{alg:silvr}
+As a result we propose a modification of Algorithm 2 by Hoogeboom et al.~\cite{hoogeboom2022equivariant} to result in the SILVR algorithm~\ref{alg:silvr}.
+We are considering latent space variables $\mathbf{z}=[\mathbf{z}^{(x)},\mathbf{z}^{(h)}]$ for the standard denoising process and  $\mathbf{\hat{z}} = [\mathbf{\hat{z}}^{(y)},\mathbf{\hat{z}}^{(h)}]$, for the set of reference coordinates given by $\mathbf{y}$.
+The vector $\mathbf{h}$ contains all the scalar node properties of the equivariant EDM.
 \begin{algorithm}
 \caption{SILVR}\label{alg:silvr}
-\begin{flushleft}
-\textbf{Input:} reference molecule $\mathbf{y,h}$, neural network $\phi$ \\
-\textbf{Output:} generated molecule $\mathbf{x,h}$
-\end{flushleft}
-\begin{algorithmic}
-\State Align to centre of gravity
-\State Sample $\hat{x}_T \sim \mathcal{N}(0,I)$
+\begin{algorithmic}[1]
+\State \textbf{Input:} Reference molecule $\mathbf{y,h}$, neural network $\phi$ 
+\State \textbf{Output:} Generated molecule $\mathbf{x,h}$
+
+\State Subtract centre of gravity (cg) from $\mathbf{y,h}$
+\State Sample $z_T \sim \mathcal{N}(\mathbf{0,I})$
 \For {$t$ = $T,...,1$}
-    \State $z \sim \mathcal{N}(0,I)$
-    \State $x_{t-1}^\prime \sim p_\theta(x_{t-1}^\prime|x_t)$
-    \State $y_{t-1} \sim q(y_{t-1}^\prime|y)$
-    \State $x_{t-1} \gets x_{t-1}^\prime - k x_{t-1}^\prime +k y_{t-1}$ \Comment{SILVR equation}
+    \State Sample $\boldsymbol{\epsilon} \sim \mathcal{N}(\mathbf{0,I})$
+    \State Subtract centre of gravity from $\boldsymbol{\epsilon}^{(x)}$
+    \State Get reference to right noise level $\mathbf{\hat{z}}_t = \alpha_t \mathbf{\hat{z}}_{t=0} +\sigma_t\times\boldsymbol{\epsilon}$
+    \State Compute $\mathbf{z}_{t-1}$ using $\phi(\mathbf{z}_t,t)$
+    \State Update $\mathbf{z}_{t-1} \gets \mathbf{z}_{t-1} - \alpha_{t-1}r_{\mathrm{S}} \mathbf{z}_{t-1} +r_{\mathrm{S}} \mathbf{\hat{z}}_{t-1}$ \Comment{SILVR equation}
+    \State Subtract centre of gravity from updated $\mathbf{z}_{t-1}$
 
 \EndFor
-\State Sample $x,h \sim p(x,h|z_0)$ 
+\State Sample $\mathbf{x,h} \sim p(\mathbf{x,h|z_0})$ 
 \end{algorithmic}
@@
     \centering
-    \includegraphics{}
-    \caption{Caption}
-    \label{fig:my_label}
+    \includegraphics[width=0.5\textwidth]{paper/Figures/Fig2/fig2.jpg}
+    \caption{Schematic illustrating the influence of SILVR of the molecules in the latent space.
+At each denoising step if $r_{mathrm{S}}>0$ the latent space vector get scaled by the SILVR rate.
+This is done from the first denoising timestep (top) until the last one (bottom).
+Repeatedly applying SILVR will result in a molecule that resembles the references (right) and does not (left)}
+    \label{fig:silvr_explanation}
 \end{figure}

### 2023-03-16 paper/manuscript.tex
pushing progress on theory

@@
 \subsection{Selective Iterative Latent Variable Refinement}
-We propose a new method, that combines the iterative variable refinement (ILVR) by Choi et al.~\cite{choi2021ilvr} and take it from the image generation process and incorporate it into the EDM by Hoogeboom et al.
-to produce a way to generate a selective iterative laten variable refinement (SILVR) procedure in which we use a reference molecule to introduce into the denoising step.
+\label{sec:silvr}
+We propose a new method, that combines the iterative variable refinement (ILVR) by Choi et al.~\cite{choi2021ilvr} proposed in an image generation context and incorporates the ideas into the EDM by Hoogeboom et al..
+This allows us to generate a selective iterative laten variable refinement (SILVR) procedure in which we introduce a reference molecule into the denoising step.
 The reference used can be a small fragment, multiple fragments or even include dummy atoms that do not exist in the original reference.
@@
 \begin{algorithmic}[1]
-\State \textbf{Input:} Reference molecule $\mathbf{y,h}$, neural network $\phi$ 
+\State \textbf{Input:} Reference molecule $\mathbf{y,h}$, EDM $\kappa$ 
 \State \textbf{Output:} Generated molecule $\mathbf{x,h}$
 
-\State Subtract centre of gravity (cg) from $\mathbf{y,h}$
+\State Compute $\mathbf{\hat{z}}_0$ from $\mathbf{y, h}$, such that $[\mathbf{z}^{(y)} , \mathbf{z}^{(h)} ] = f (\mathbf{y, h})$ is E(3). 
+\State Subtract center of gravity (COG) from $\mathbf{\hat{z}}_0$
 \State Sample $z_T \sim \mathcal{N}(\mathbf{0,I})$
@@
     \State Sample $\boldsymbol{\epsilon} \sim \mathcal{N}(\mathbf{0,I})$
-    \State Subtract centre of gravity from $\boldsymbol{\epsilon}^{(x)}$
-    \State Get reference to right noise level $\mathbf{\hat{z}}_t = \alpha_t \mathbf{\hat{z}}_{t=0} +\sigma_t\times\boldsymbol{\epsilon}$
-    \State Compute $\mathbf{z}_{t-1}$ using $\phi(\mathbf{z}_t,t)$
+    \State Subtract COG from $\boldsymbol{\epsilon}^{(x)}$
+    \State $\mathbf{\hat{z}}_{t-1} = \alpha_{t-1} \mathbf{\hat{z}}_{0} +\sigma_{t-1}\times\boldsymbol{\epsilon}$
+    \Comment{Noise reference to t=t-1}
+    \State $\mathbf{z}_{t-1} \gets \kappa (\mathbf{z}_t,t)$
+    \Comment{Compute denosing step}
     \State Update $\mathbf{z}_{t-1} \gets \mathbf{z}_{t-1} - \alpha_{t-1}r_{\mathrm{S}} \mathbf{z}_{t-1} +r_{\mathrm{S}} \mathbf{\hat{z}}_{t-1}$ \Comment{SILVR equation}
-    \State Subtract centre of gravity from updated $\mathbf{z}_{t-1}$
+    \State Subtract COG from updated $\mathbf{z}_{t-1}$
 
 \EndFor
+\State Add COG (from line 4) to $\mathbf{z}_0$
+\Comment{Move sample back to pocket}
 \State Sample $\mathbf{x,h} \sim p(\mathbf{x,h|z_0})$ 
@@
 The resulting SILVR model produces conditional samples without any conditional training when generating new molecules.
+Figure~\ref{fig:silvr_explanation} shows an illustrative example of how the SILVR rate $r_{\mathrm{S}}$ is used to shift the latent space vector $\mathbf{z}_{t-1}$ at any point in the denoising process from $T\ldots t=1$.
+Using the silver equation as defined in the SILVR algorithm~\ref{alg:silvr} line 11, an existing denoising step (green) is brought closer to the reference (purple) in latent space.
+
+\section{Methods}
 To illustrate the usefulness of this run-time modification, we show how SILVR can be used in the context of fragment-based drug design.
 The goal is to produce molecules that resemble a binding site based on exiting fragments and their set of atomic coordinates.
-
-\section{Methods}
-As stated before, the selective iterative latent variable refinement (SILVR) method makes use of the pre-trained equivariant diffusion model (EDM) as introduced by Hoogeboom \textit{et al.}~\cite{hoogeboom2022equivariant}.
+%As stated before, the selective iterative latent variable refinement (SILVR) method makes use of the pre-trained equivariant diffusion model (EDM) as introduced by Hoogeboom \textit{et al.}~\cite{hoogeboom2022equivariant}.
 
@@
 This model strictly only considers atomic coordinates; all bond information is ignored.
-A more recent version of the EDM also accounts for bond order and SILVR could also be incorporated into this new model.~\cite{hvignac2023midi}.
+A more recent version of the EDM also accounts for bond order and SILVR could also be incorporated into this new model~\cite{hvignac2023midi}.
 During training, the atomic coordinates (and their element) are passed through a forward diffusion process with iterative addition of Gaussian noise, where the extent of noise added at each step is defined by parameter $\beta$ (N.B.
-scheme X shows the diffusion process as a Markov chain, however in practice the state at time $t$ is determined as a direct transformation of the initial state).
+scheme~\ref{fig:schematic} shows the diffusion process as a Markov chain, however in practice the state at time $t$ is determined as a direct transformation of the initial state).
 The diffusion process is eventually terminated when $t=T$, by which point all structure is lost and all coordinates follow a Gaussian distribution.
@@
 At run time, the generative model is instantiated with a sample from a Gaussian distribution and the series of denoising steps are applied resulting in a generated sample consisting of atomic coordinates resembling a drug-like molecule.
-The XYZ coordinates can then be interpreted using cheminformatics software to determine the molecular graph.
+The \texttt{XYZ} coordinates can then be interpreted using cheminformatics software to determine the molecular graph.
 
-The EDM was then adapted by introducing SILVR within the denoising process algorithm~\ref{alg:silvr}.
+The EDM was then adapted by introducing SILVR within the denoising process algorithm~\ref{alg:silvr}, as outlined in~\ref{sec:silvr}.
 At run time, each atom of the reference set of coordinates is mapped to an atom in the EDM.
@@
 A small refinement is applied to add information from the reference coordinates to the latent variable of the denoising process~\ref{eq:denoising}.
-This equation has the effect of shrinking the coordinates towards the origin, and then expanding the coordinates out in the direction of the reference.
+This equation has the effect of shrinking the coordinates towards the origin, and then expanding the coordinates out in the direction of the reference, see Figure~\ref{fig:silvr_explanation}.
 Importantly, the extent of this refinement is defined by the SILVR rate $r_{\mathrm{S}}$, with $r_{\mathrm{S}}=0$ providing no additional refinement and $r_{\mathrm{S}}=1$ will result in a total replacement of atoms.
@@
 \subsection{Reference Dataset: Covid-moonshot}
-Reference molecules were selected from the original 23 non-covalent hits of the SARS-CoV-2 Main Protease (Mpro) identified by the XChem fragment screen as part of the Covid Moonshot Project~\cite{moonshot}.
-A more detailed picture of all fragements is presented in Figure~\ref{si:figure}.
-Fragments were visualised using XXX v XXX and combinations were arbitrarily selected as test cases.
-Fragments \texttt{x0072} and \texttt{x0354} were selected for benchmarking the effect of $r_{\mathrm{S}}$ on sampling; \texttt{x1093}, \texttt{x0072}, \texttt{x2193} were selected to represent three significantly overlapping fragments; \texttt{x0434}, \texttt{x0305} and \texttt{x0072}, \textit{x0354} were selected as partially overlapping fragments; and \texttt{x0874}, \texttt{x0397} were selected as two disconnected fragments, resembling a linker design type problem.
-The bonding information of selected fragments was deleted and XYZ coordinates were combined into a single XYZ file.
-Values of $r_{\mathrm{S}}$ were selected and added to the XYZ file to create a reference file containing all experiment setup information.
+Reference molecules were selected from the original 23 non-covalent hits of the SARS-CoV-2 Main Protease (Mpro) identified by the XChem fragment screen as part of the Covid Moonshot Project~\cite{consortium2023open, consortium2021open}.
+A more detailed picture of all fragements is presented in Figure S1 of the SI~\ref{si}.
+Fragments were visualised using NGLview version 3.03 and combinations were arbitrarily selected as test cases.
+Fragments \texttt{x0072} and \texttt{x0354} were selected for benchmarking the effect of $r_{\mathrm{S}}$ on sampling; \texttt{x1093}, \texttt{x0072}, \texttt{x2193} were selected to represent three significantly overlapping fragments; \texttt{x0434}, \texttt{x0305} and \texttt{x0072}, \texttt{x0354} were selected as partially overlapping fragments; and \texttt{x0874}, \texttt{x0397} were selected as two disconnected fragments, resembling a linker design type problem.
+The bonding information of selected fragments was deleted and \texttt{XYZ} coordinates were combined into a single \texttt{XYZ} file.
+Values of $r_{\mathrm{S}}$ were selected and added to the \texttt{XYZ} file to create a reference file containing all experiment setup information.
 Each experiment was sampled 1000 times.
 
-\subsection{Diferent observables were monitored to gauge the success of computational experiments for the generation of new models}
+\subsection{Different observables were monitored to gauge the success of computational experiments for the generation of new models}
 
-Different observables were used to monitor how realistic and reliable new molecules were generated and how well they can fit into the existing binding site of the SARS-CoV-2 main protease.
+Different observables were used to monitor how realistic and reliable newly generated molecules were and how well they can fit into the existing binding site of MPro.
+We looked at the following set of measures:
 
 \subsubsection{Atom stability}
-The accuracy of placement of atoms was determined using the stability metric proposed by Hoogeboom \textit{et al.}.
+The accuracy of placement of atoms was determined using the stability metric proposed by Hoogeboom et al..
 This metric infers bonds and bond orders between atoms by considering their interatomic distances.
@@
 It should be noted that this metric requires the presence of all hydrogens (both explicit and implicit) for an atom to be decided as stable.
-For comparability with other similar published models, this metric was used unmodified.
-The additional metric of “molecular stability” is often reported together with atom stability (if every atom is stable then the whole molecule is stable), however as has been previously identified, large molecules sampled from this GEOM-trained EDM tend to be unstable.
+For comparability with other similar published models, this measure was used unmodified.
+The additional measure of \textit{molecular stability} is often reported together with atom stability (if every atom is stable then the whole molecule is stable), however as has been previously identified, large molecules sampled from this GEOM-trained EDM tend to be unstable.
 
@@
 The RMSD for this pairwise mapping, ignoring atom identities, was calculated to determine the spacial similarity of samples to the reference.
+All RMSD calculations were carried out using \url{https://github.com/charnley/rmsd} version 1.5.1.
 
 \subsubsection{Geometric stability - Auto3D}
-To determine whether samples represent a true molecular geometry, independent minimisation of molecular geometries was performed using Atoms In Molecules Neural Network Potential (AIMNet) with Auto3D.
+To determine whether samples represent a true molecular geometry, an independent minimisation of molecular geometries was performed using Atoms In Molecules Neural Network Potential (AIMNet) with Auto3D.
 All samples were read by RDKit and samples containing more than one molecule were removed from the test set.
 The SMILES string of each molecule was written to a new file, read by Auto3D, and geometry predicted by AIMNet.
-The RMSD between SILVR-generated coordinates and the Auto3D minimised coordinates were calculated with RDKit.
+The RMSD between SILVR-generated coordinates and the Auto3D minimised coordinates were calculated with RDKit version 2022.03.5.
 
@@
 The XYZ coordinates of samples were directly read into OpenEye Tools and the pose was re-scored with \textit{shapegauss}.
-
-\subsubsection{Maybe: Docking of molecules}
 

### 2023-03-16 paper/manuscript.tex
First musings of an introduction

@@
 \title[SILVR: Molecular Generation for binding modes]
-  {SILVR: Guided molecular generation from diffusion models without additional training}
+  {SILVR:Directed Diffusion for Molecule Generation}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \label{sec:intro}
+One of the challenges as part of computer aided drug-design is to sample from the very large space of possible drug-like compounds to find suitable hits for a given target protein.
+It is estimated that there are between $10^23$ to $10^60$ possibly feasible compounds, while only around $10^8$ have been synthesised so far~\cite{polishchuk2013estimation, reymond2012exploring}.
+Different strategies have been employed in the past to try and sample a diverse and synthetically accessible molecular space.
+In particular, methods based on machine-learning have shown vast amount of promise in recent years in this space~\cite{bilodeau2022generative}.
+Various different neural network architectures have been proposed for molecular generation, from Varational autoencoders (VAE)~\cite{kingma2022autoencoding, jin2018junction, ma2018constrained}, to generative adversarial networks (GAN)~\cite{hoffmann2019generating} and normalising flows~\cite{shi2021learning}.
+More recently denoising difusion probailistic models, and particularly equivarient diffusion models have shown promise in moleculear generation~\cite{xu2022geodiff, hoogeboom2022equivariant}.
+
+- Model generation isn't everything
+- How do models fit into hit design, conditioned training needed.
+
+Linker learning with reinforcement learning~\cite{guo2023linkinvent}.
+
 - Equivariant diffusion models have been very promising in predicting 3D structures of drug-like molecules. 
@@
 \label{sec:theory}
-\subsection{Unconditional Equivariant Denoising Diffusion probabilistic models}
-\subsubsection*{Denoising Diffusion probabilistic models (DDPM) as generative models}
-DDPMs are often used as generative models that were developed for the generation of new images~\cite{}.
-More recently the same idea has also been applied to molecular generation~\cite{}.
+\subsection*{Denoising Diffusion probabilistic models (DDPM) as generative models}
+DDPMs are often used as generative models that were developed for the generation of new images~\cite{sohl-dickstein2015deep, ho2020denoising, nichol2021improved}.
+More recently the same idea has also been applied to molecular generation~\cite{hoogeboom2022equivariant}.
 The main idea behind diffusion models is for a neural network to learn the reverse of a diffusion process, often referred to as denoising to \textit{sample} a new image or in our case a new molecule.
-In practice this is done by training a neural network $\phi$ and generating samples $\hat{\mathbf{x}}=\psi(X_t,t)$, with $X_t$ with data $X$, noised up to time $t$.
-Figure~\ref{fig:diffusion_schematic} shows a schematic of the two main parts of this model, the diffusion part where noise is added for each timestep (shown in blue) and the denoising part (shown in yellow) that allows the generation of a new image (molecule).
+In practice this is done by training a neural network $\phi$ and generating samples $p(\mathbf{x_{t-1}}|\mathbf{x}_t)=\mathcal(x_{t-1}; \{\mu_{\theta}(x_t;t), \sigma_t^2\mathbf{I}$ from  a Gaussian transition with learned mean ($\mu_{\theta}(x_t;t)$) and variance ($\sigma_t$) with $\mathbf{x}_t$ with data $X$, noised up to time $t$.
+Figure~\ref{fig:diffusion_schematic} shows a schematic of the two main parts of such a DDPM, the diffusion part where noise is added for each timestep (shown in blue) and the denoising part (shown in yellow) that allows the generation of a new image (molecule).
 
@@
 % Next talk about the denoising step
-$\hat{p}$ is the parametric probability distribution we want to learn. 
-\subsubsection*{Euquivariant diffusion model}
+$p_{\theta}$ is the parametric probability distribution we want to learn. This generative process is given by:
+
+\begin{equation}
+p(\mathbf{x_{t-1}}|\mathbf{x}_t)=\mathcal(x_{t-1}; \{\mu_{\theta}(x_t;t), \sigma_t^2\mathbf{I}.
+\end{equation}
+A sample for denoising timestep $t-1$ can be generated from the following equation given the neural network $\phi$ that has been trained on the diffusion process.
+
+A given samples is generated as:
+\begin{equation}
+ x_{t-1} = \frac{1}{\alpha_t}(x_t-\frac{1-\alpha_t}{\sqrt{1-<\alpha_t>}}\phi(x_t,t) +\sigma z,   
+\end{equation}
+
+with $z\sim\mathcal{N}(0,\mathbf{I})$.
+
+The process is iterated until $t=0$ and as such a new sample is generated.
+
+%%%%%%%
+% EDMs
+%%%%%%%
+\subsection*{Euquivariant diffusion model}
 Unlike in image problems, the orientation of molecules matter under translations and rotations.
@@
 The positions translations and rotations are defined according to the orthogonal matrix: $\mathbf{Rx + t} = (\mathbf{R_{x_1}+t},\ldots, \mathbf{R_{x_N}+t})$.
-Satorras et al.~\cite{} have proposed an E(n) Equivariant Graph Neural Network on which the EDM by Hoogeboom et al.
+Satorras et al.~\cite{satorras2022equivariant} have proposed an E(n) Equivariant Graph Neural Network on which the EDM by Hoogeboom et al.
 builds.
@@
     P_\theta(x_0|c) = \int p_\theta(x_{0:T}|c)Dx_{1:T}
+    p_{\theta}(x_{0:T}|c)=p(x_T)\prod_{t=1}^{T}p_{\theta}(x_{t-1}\x_t,c)
 \end{equation}
+This iterative latenent variable refinement (ILVR) means that the condition can be applied during the denoising step without additional training.
+In the case of Choi et al, they used downsampled reference images as conditioning in order to generate images similar to the reference, but still a newly generated images.
 
@@
 \label{sec:silvr}
-We propose a new method, that combines the iterative variable refinement (ILVR) by Choi et al.~\cite{choi2021ilvr} proposed in an image generation context and incorporates the ideas into the EDM by Hoogeboom et al..
+We propose a new method, that combines ILVR by Choi et al.~\cite{choi2021ilvr} proposed in an image generation context and incorporates the ideas into the EDM by Hoogeboom et al..
 This allows us to generate a selective iterative laten variable refinement (SILVR) procedure in which we introduce a reference molecule into the denoising step.
@@
     \State $\mathbf{z}_{t-1} \gets \kappa (\mathbf{z}_t,t)$
-    \Comment{Compute denosing step}
+    \Comment{Compute denoising step}
     \State Update $\mathbf{z}_{t-1} \gets \mathbf{z}_{t-1} - \alpha_{t-1}r_{\mathrm{S}} \mathbf{z}_{t-1} +r_{\mathrm{S}} \mathbf{\hat{z}}_{t-1}$ \Comment{SILVR equation}
@@
 \EndFor
-\State Add COG (from line 4) to $\mathbf{z}_0$
+\State Add COG($\mathbf{\hat{z}}_0$) to $\mathbf{z}_0$
 \Comment{Move sample back to pocket}
@@
 \end{algorithm}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.5\textwidth]{paper/Figures/Fig2/fig2.jpg}
-    \caption{Schematic illustrating the influence of SILVR of the molecules in the latent space.
-At each denoising step if $r_{mathrm{S}}>0$ the latent space vector get scaled by the SILVR rate.
-This is done from the first denoising timestep (top) until the last one (bottom).
-Repeatedly applying SILVR will result in a molecule that resembles the references (right) and does not (left)}
-    \label{fig:silvr_explanation}
-\end{figure}
 
@@
 \subsection{Conditioning the pre-trained EDM with SILVR}
-The EDM by Hoodgeboom \textit{et al.} was trained on the 30 lowest energy conformations of 430,000 molecules from the GEOM dataset with explicit protons.
+The EDM by Hoodgeboom \textit{et al.} was trained on the 30 lowest energy conformations of 430,000 molecules from the GEOM dataset with explicit protons~\cite{axelrod2022geom}.
 A simplified schematic describing the EDM architecture is shared in scheme~\ref{fig:_architecture} for more details on the model theory and training refer to Hoogeboom et al.~\cite{hoogeboom2022equivariant}.
@@
 The \texttt{XYZ} coordinates can then be interpreted using cheminformatics software to determine the molecular graph.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\textwidth]{paper/Figures/Fig2/fig2.jpg}
+    \caption{Schematic illustrating the influence of SILVR of the molecules in the latent space.
+At each denoising step if $r_{mathrm{S}}>0$ the latent space vector get scaled by the SILVR rate.
+This is done from the first denoising timestep (top) until the last one (bottom).
+Repeatedly applying SILVR will result in a molecule that resembles the references (right) and does not (left)}
+    \label{fig:silvr_explanation}
+\end{figure}
 

### 2023-03-17 paper/SI.tex

@@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% This is a (brief) model paper using the achemso class
+%% The document class accepts keyval options, which should include
+%% the target journal and optionally the manuscript type. 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\documentclass[journal=jacsat,manuscript=article]{achemso}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Place any additional packages needed here.  Only include packages
+%% which are essential, to avoid problems later. Do NOT use any
+%% packages which require e-TeX (for example etoolbox): the e-TeX
+%% extensions are not currently available on the ACS conversion
+%% servers.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\usepackage[version=3]{mhchem} % Formula subscripts using \ce{}
+\usepackage{algorithm}
+\usepackage{algpseudocode}
+\usepackage{svg}
+\usepackage{amsmath}
+\usepackage{hyperref}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%https://www.overleaf.com/project/63f3627b7f8554c152639004
+%% If issues arise when submitting your manuscript, you may want to
+%% un-comment the next line.  This provides information on the
+%% version of every file you have used.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%\listfiles
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Place any additional macros here.  Please use \newcommand* where
+%% possible, and avoid layout-changing macros (which are not used
+%% when typesetting).
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newcommand*\mycommand[1]{\texttt{\emph{#1}}}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Meta-data block
+%% ---------------
+%% Each author should be given as a separate \author command.
+%%
+%% Corresponding authors should have an e-mail given after the author
+%% name as an \email command. Phone and fax numbers can be given
+%% using \phone and \fax, respectively; this information is optional.
+%%
+%% The affiliation of authors is given after the authors; each
+%% \affiliation command applies to all preceding authors not already
+%% assigned an affiliation.
+%%
+%% The affiliation takes an option argument for the short name.  This
+%% will typically be something like "University of Somewhere".
+%%
+%% The \altaffiliation macro should be used for new address, etc.
+%% On the other hand, \alsoaffiliation is used on a per author basis
+%% when authors are associated with multiple institutions.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\author{Nicholas Runcie}
+\altaffiliation{EaSTCHEM School of Chemistry, University of Edinburgh, EH9 3FJ}
+\author{Antonia S.J.S. Mey}
+\altaffiliation{EaSTCHEM School of Chemistry, University of Edinburgh, EH9 3FJ}
+\email{antonia.mey@ed.ac.uk}
+\phone{xxx}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% The document title should be given as usual. Some journals require
+%% a running title from the author: this should be supplied as an
+%% optional argument to \title.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\title[SILVR: Molecular Generation for binding modes]
+  {SILVR: Conditional diffusion model for molecule generation without additional training}
+\title[SILVR: Molecular Generation for binding modes]
+  {Supporting Information for: SILVR: Directed Diffusion for Molecule Generation}
+\begin{document}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-This will usually read something like: ``Experimental procedures and
-characterization data for all new compounds. The class will
-automatically add a sentence pointing to the information on-line:
 \section{Dataset}
-Here have a summary of all the fragment data
+Figure~\ref{fig:fragments} shows all MPro fragments used for this study.
+
+\begin{figure}
+    \centering
+    \includegraphics{}
+    \caption{Caption}
+    \label{fig:fragments}
+\end{figure}
 
@@
 \end{suppinfo}
+
+\end{document}

### 2023-03-17 paper/manuscript.tex
Updates from Overleaf

@@
 \title[SILVR: Molecular Generation for binding modes]
-  {SILVR:Directed Diffusion for Molecule Generation}
+  {SILVR: Directed (Guided?) Diffusion for Molecule Generation}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 One of the challenges as part of computer aided drug-design is to sample from the very large space of possible drug-like compounds to find suitable hits for a given target protein.
-It is estimated that there are between $10^23$ to $10^60$ possibly feasible compounds, while only around $10^8$ have been synthesised so far~\cite{polishchuk2013estimation, reymond2012exploring}.
+It is estimated that there are between $10^{23}$ to $10^{60}$ possibly feasible compounds, while only around $10^8$ have been synthesised so far~\cite{polishchuk2013estimation, reymond2012exploring}.
 Different strategies have been employed in the past to try and sample a diverse and synthetically accessible molecular space.

### 2023-03-17 paper/SI.tex

@@
 
-\section{Dataset}
-Figure~\ref{fig:fragments} shows all MPro fragments used for this study.
+\paragraph{Dataset}
+Figure~\ref{fig:fragments} shows all MPro fragments used for this study from the covid moonshot dataset~\cite{consortium2021open,consortium2023open}.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.8\textwidth]{paper/Figures/FigS1/mpro_ligands_grid_2d.png}
+    \caption{2D Mpro structures from the moonshot dataset.}
+    \label{fig:fragments}
+\end{figure}
+\newpage
+\paragraph{Curated samples from different SILVR rates}
+Figure~\ref{fig:fig_2} contains two examples started from fragments \texttt{xxx} and \texttt{xxx} and resulting samples for different SILVR rates $r_{\mathrm{S}}$.
+\begin{figure}[hb!]
+    \centering
+    \includegraphics[width=\textwidth]{paper/Figures/FigS2/fig_2_sample_vs_rate.png}
+    \caption{Two random samples from the same reference using different SILVR rates.
+All bonds were inferred from \texttt{XYZ} coordinates with OpenBabel.
+All bonds were visualised as single bonds and hydrogen atoms were deleted for clarity.
+Increasing SILVR rate results in sampled atom coordinates coming closer in space, and element type, to the reference while still resembling a truly molecular structure.}
+    \label{fig:fig_2}
+\end{figure}
+\nemapge
+
+\paragraph{Examples of uncurated samples}
 
@@
     \caption{Caption}
-    \label{fig:fragments}
-\end{figure}
-
-\section{Samples from different SILVR rates}
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{paper/Figures/FigS2/fig_2_sample_vs_rate.png}
-    \caption{Two random samples from the same reference using different SILVR rates.
-All bonds were inferred from XYZ coordinates with openbabel.
-All bonds were visualised as single bonds and hydrogen atoms were deleted for clarity.
-Increasing SILVR rate results in sampled atom coordinates coming closer in space, and element type, to the reference while still resembling a true molecular structure.}
-    \label{fig:fig_2}
+    \label{fig:my_label}
 \end{figure}

### 2023-03-18 paper/SI.tex
Updates from Overleaf

@@
 \end{figure}
-\nemapge
+
+\begin{figure}
+    \centering
+    \includegraphics[widht=0.5\textwidth]{paper/Figures/FigS4/FigS4.png}
+    \caption{Fraction of molecules fragmented with respect to the SILVR rate}
+    \label{fig:my_label}
+\end{figure}
+
+\newapge
 
@@
     \centering
-    \includegraphics{}
-    \caption{Caption}
+    \includegraphics[width=0.8\textwidth]{paper/Figures/FigS5/mols_grid_silvr_rate_not_curated.png}
+    \caption{Uncurated samples, showing fragmentation of molecules based on input with two starting structures.}
     \label{fig:my_label}

### 2023-03-18 paper/SI.tex
Finalising the SI

@@
 
+The supporting information consists of a series of figures in addition to the figures in the main paper.
+Figure~\ref{fig:fragments} contains the 2D structures of all fragments used in the experiments as references.
+In Figure~\ref{fig:fig_2} an example of two curated experiments is shown with different SILVR rates $r_{mathrm{s}}$.
+An example of how far atoms, even involved in a reference structure are displaced in the denoising diffusion process is shown in an example in Figure~\ref{fig:atom_displacement} \textbf{A} for a whole molecule generated and in \textbf{B} for the heterocycle.
+The next figure~\ref{fig:fragments} shows on average how many molecules were not-fragmented in a set of 1000 samples for different SILVR rates.
+The final figure~\ref{fig:uncurated_sample} shows a series of 2D samples from an uncurated list of samples, clearly showing examples of fragmented molecules.
 \paragraph{Dataset}
@@
 \newpage
-\paragraph{Curated samples from different SILVR rates}
-Figure~\ref{fig:fig_2} contains two examples started from fragments \texttt{xxx} and \texttt{xxx} and resulting samples for different SILVR rates $r_{\mathrm{S}}$.
+
 \begin{figure}[hb!]
@@
     \centering
+    \includegraphics[width=0.8\textwidth]{paper/Figures/FigS3/FigS3.png}
+    \caption{Example looking at the displacement of reference atoms after diffusion. A: for the whole molecule. B: Zoom in, on the displacement of atoms in the 9aH-Pyrido[1,2-a]pyrazine heterocycle. }
+    \label{fig:atom_displacement}
+\end{figure}
+
+\begin{figure}
+    \centering
     \includegraphics[widht=0.5\textwidth]{paper/Figures/FigS4/FigS4.png}
@@
 
-\paragraph{Examples of uncurated samples}
-
 \begin{figure}
@@
     \caption{Uncurated samples, showing fragmentation of molecules based on input with two starting structures.}
-    \label{fig:my_label}
+    \label{fig:uncurated_sample}
 \end{figure}

### 2023-03-18 paper/manuscript.tex
Finalising the SI

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Supporting Information
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newpage 
+\begin{suppinfo}
+
+The supporting information consists of a series of figures in addition to the figures in the main paper.
+Figure~\ref{fig:fragments} contains the 2D structures of all fragments used in the experiments as references.
+In Figure~\ref{fig:fig_2} an example of two curated experiments is shown with different SILVR rates $r_{mathrm{s}}$.
+An example of how far atoms, even involved in a reference structure are displaced in the denoising diffusion process is shown in an example in Figure~\ref{fig:atom_displacement} \textbf{A} for a whole molecule generated and in \textbf{B} for the heterocycle.
+The next figure~\ref{fig:fragments} shows on average how many molecules were not-fragmented in a set of 1000 samples for different SILVR rates.
+The final figure~\ref{fig:uncurated_sample} shows a series of 2D samples from an uncurated list of samples, clearly showing examples of fragmented molecules.
+\paragraph{Dataset}
+Figure~\ref{fig:fragments} shows all MPro fragments used for this study from the covid moonshot dataset~\cite{consortium2021open,consortium2023open}.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.8\textwidth]{paper/Figures/FigS1/mpro_ligands_grid_2d.png}
+    \caption{2D Mpro structures from the moonshot dataset.}
+    \label{fig:fragments}
+\end{figure}
+\newpage
+
+\begin{figure}[hb!]
+    \centering
+    \includegraphics[width=\textwidth]{paper/Figures/FigS2/fig_2_sample_vs_rate.png}
+    \caption{Two random samples from the same reference using different SILVR rates.
+All bonds were inferred from \texttt{XYZ} coordinates with OpenBabel.
+All bonds were visualised as single bonds and hydrogen atoms were deleted for clarity.
+Increasing SILVR rate results in sampled atom coordinates coming closer in space, and element type, to the reference while still resembling a truly molecular structure.}
+    \label{fig:fig_2}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.8\textwidth]{paper/Figures/FigS3/FigS3.png}
+    \caption{Example looking at the displacement of reference atoms after diffusion. A: for the whole molecule. B: Zoom in, on the displacement of atoms in the 9aH-Pyrido[1,2-a]pyrazine heterocycle. }
+    \label{fig:atom_displacement}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[widht=0.5\textwidth]{paper/Figures/FigS4/FigS4.png}
+    \caption{Fraction of molecules fragmented with respect to the SILVR rate}
+    \label{fig:my_label}
+\end{figure}
+
+\newapge
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.8\textwidth]{paper/Figures/FigS5/mols_grid_silvr_rate_not_curated.png}
+    \caption{Uncurated samples, showing fragmentation of molecules based on input with two starting structures.}
+    \label{fig:uncurated_sample}
+\end{figure}
+
+\end{suppinfo}
+
 \end{document}

### 2023-03-18 paper/manuscript.tex
Updates from Overleaf

@@
 \begin{tocentry}
-
-Some journals require a graphical entry for the Table of Contents.
-This should be laid out ``print ready'' so that the sizing of the
-text is correct.
-
-Inside the \texttt{tocentry} environment, the font used is Helvetica
-8\,pt, as required by \emph{Journal of the American Chemical
-Society}.
-
-The surrounding frame is 9\,cm by 3.5\,cm, which is the maximum
-permitted for  \emph{Journal of the American Chemical Society}
-graphical table of content entries. The box will not resize if the
-content is too big: instead it will overflow the edge of the box.
-
-This box and the associated title will always be printed on a
-separate page at the end of the document.
-
+    \includegraphics{paper/Figures/TOC/TOC_stickman_to_mol.tif}
+    \caption{For Table of Contents Only}
 \end{tocentry}

### 2023-03-18 paper/manuscript.tex

@@
 \begin{tocentry}
-    \includegraphics{paper/Figures/TOC/TOC_stickman_to_mol.tif}
+\centering
+    \includegraphics{paper/Figures/TOC/TOC_stickman_to_mol.jpg}
     \caption{For Table of Contents Only}
+    \label{fig:toc}
 \end{tocentry}
@@
 \label{sec:intro}
-One of the challenges as part of computer aided drug-design is to sample from the very large space of possible drug-like compounds to find suitable hits for a given target protein.
+One of the challenges as part of computer-aided drug design is to sample from a very large space of possible drug-like compounds to find suitable hits for a given target protein.
 It is estimated that there are between $10^{23}$ to $10^{60}$ possibly feasible compounds, while only around $10^8$ have been synthesised so far~\cite{polishchuk2013estimation, reymond2012exploring}.
 Different strategies have been employed in the past to try and sample a diverse and synthetically accessible molecular space.
-In particular, methods based on machine-learning have shown vast amount of promise in recent years in this space~\cite{bilodeau2022generative}.
-Various different neural network architectures have been proposed for molecular generation, from Varational autoencoders (VAE)~\cite{kingma2022autoencoding, jin2018junction, ma2018constrained}, to generative adversarial networks (GAN)~\cite{hoffmann2019generating} and normalising flows~\cite{shi2021learning}.
-More recently denoising difusion probailistic models, and particularly equivarient diffusion models have shown promise in moleculear generation~\cite{xu2022geodiff, hoogeboom2022equivariant}.
+In particular, methods based on machine learning have shown vast promise in recent years in this space~\cite{bilodeau2022generative}.
+Various different neural network architectures have been proposed for molecular generation, from Variational autoencoders (VAE)~\cite{kingma2022autoencoding, jin2018junction, ma2018constrained, agoza2022generating}, to generative adversarial networks (GAN)~\cite{hoffmann2019generating} and normalising flows~\cite{shi2021learning}.
+More recently denoising diffusion probabilistic models and particularly equivariant diffusion models have shown promise in molecular generation~\cite{xu2022geodiff, hoogeboom2022equivariant}.
 
-- Model generation isn't everything
-- How do models fit into hit design, conditioned training needed.
+Typically the objective to achieve is to generate a diverse set of molecules, for a given target protein that are easily synthetically accessible and ideally with a high binding affinity and low toxicity prediction~\cite{some gneric caad}.
+Different machine learning-based  and other strategies have been used to address these ideas.
+From Recurrent Neural networks (RNN)~\cite{xu2021novo,zhang2022novo} and conditional variational autoencoders~\cite{ragoza2022generating} to GANs~\cite{skalic2019target}, various approaches have been used to generate molecules for a binding site where in each case the training is conditioned based on the target protein~\cite{xie2022advances}.
+Some of these models incorporate a ligand score directly~\cite{ragoza2022generating}, while others would require other methods through machine learning~\cite{masters2023deep}, or more traditional docking~\cite{fu2018predictive} and molecular simulation-based affinity prediction methods~\cite{mey2016blinded,mey2020best}.
+Even with a variety of ways to assess the synthetic accessibility of generated compounds~\cite{coley2018scscore, thakkar2021retrosynthetic} often molecules generated in these method are  not easily sythesiable at best or chemically infeasible in the worst cases.
+Another desirable features for a generative model to have are the ability to generate compounds that are extensions to a known fragment hit or link two existing fragments through different linkers with the desire to improve potency, solubility or other properties of existing hits.
+Goa et al.
+have introduced way in which to generate linkers for fragments using reinforcement learning strategies~\cite{guo2023linkinvent}, while Imrie et al.~\cite{imrie2020deep} have used variation atuo encoders on this tasks without the use of protein information.
+More recently Huang et al.~\cite{huang20223dlinker} and Igashov et al.~\cite{igashov2022equivariant} have tackeled this challenge based on equivariant variational auto-encoder and diffusion based models models respectively.
+Each of these models were trained for the specific purpose of linker generation.
+Another objective from a medchem perspective is that once a hit has been found is to generate new compounds that improve this hit to become a good lead compound for optimisation.
 
-Linker learning with reinforcement learning~\cite{guo2023linkinvent}.
-
-- Equivariant diffusion models have been very promising in predicting 3D structures of drug-like molecules. 
-- This does not make them suitable candidates for a target binding site. 
-- We combine equivariant diffusion models~\cite{huang2022mdm} with Iterative latent variable refinement proposed in Denoising Diffusion probabilistic models~\cite{choi2021ilvr}.
-- This allows the generation of new molecules in the shape of the binding site using information from existing fragments.
+The work we present here is a framework for conditioning and existing pretrained equivaraint diffusion model (EDM) towards the task of linker generation between fragments and generating similar compounds to existing hits without specific training towards this task.
+To achieve this we combine the EDM by Hoogeboom et al.~\cite{hoogeboom2022equivariant} with the iterative latent variable refinement method proposed by Choi et al.
+for image generation networks~\cite{choi2021ilvr}.
+This allows the generation of new molecules in the shape of a binding site using information from existing fragments or linking fragments without specific training or knowledge of this task.
 \begin{figure}
@@
 \end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%
+% Theory
+%%%%%%%%%%%%%%%%%%%%%%%
 \section{Theory}

### 2023-03-18 paper/manuscript.tex

@@
 Some of these models incorporate a ligand score directly~\cite{ragoza2022generating}, while others would require other methods through machine learning~\cite{masters2023deep}, or more traditional docking~\cite{fu2018predictive} and molecular simulation-based affinity prediction methods~\cite{mey2016blinded,mey2020best}.
-Even with a variety of ways to assess the synthetic accessibility of generated compounds~\cite{coley2018scscore, thakkar2021retrosynthetic} often molecules generated in these method are  not easily sythesiable at best or chemically infeasible in the worst cases.
+Even with a variety of ways to assess the synthetic accessibility of generated compounds~\cite{coley2018scscore, thakkar2021retrosynthetic} often molecules generated in these method are  not easily sythesisable at best or chemically infeasible in the worst cases.
 Another desirable features for a generative model to have are the ability to generate compounds that are extensions to a known fragment hit or link two existing fragments through different linkers with the desire to improve potency, solubility or other properties of existing hits.
 Goa et al.
-have introduced way in which to generate linkers for fragments using reinforcement learning strategies~\cite{guo2023linkinvent}, while Imrie et al.~\cite{imrie2020deep} have used variation atuo encoders on this tasks without the use of protein information.
-More recently Huang et al.~\cite{huang20223dlinker} and Igashov et al.~\cite{igashov2022equivariant} have tackeled this challenge based on equivariant variational auto-encoder and diffusion based models models respectively.
+have introduced way in which to generate linkers for fragments using reinforcement learning strategies~\cite{guo2023linkinvent}, while Imrie et al.~\cite{imrie2020deep} have used variation auto encoders on this tasks without the use of protein information.
+More recently Huang et al.~\cite{huang20223dlinker} and Igashov et al.~\cite{igashov2022equivariant} have tackled this challenge based on equivariant variational auto-encoder and diffusion based models models respectively.
 Each of these models were trained for the specific purpose of linker generation.
@@
 
-The work we present here is a framework for conditioning and existing pretrained equivaraint diffusion model (EDM) towards the task of linker generation between fragments and generating similar compounds to existing hits without specific training towards this task.
+We present, the selective iterative variable refinement (SILVR), a method for conditioning an existing pre-trained equivaraint diffusion model (EDM) towards the task of linker generation between fragments and generating similar compounds to existing hits without specific training towards this task.
 To achieve this we combine the EDM by Hoogeboom et al.~\cite{hoogeboom2022equivariant} with the iterative latent variable refinement method proposed by Choi et al.
@@
 This allows the generation of new molecules in the shape of a binding site using information from existing fragments or linking fragments without specific training or knowledge of this task.
+
 \begin{figure}
@@
 \end{figure}
+
+Denoising probabilistic diffusion models can be separated into two parts, the diffusion part and the denoising part, as shown in the schematic in Figure~\ref{fig:schematic} with the ideas originating in image generation machine learning problems.
+A neural network is trained to learn the second part of the process, denoising a Gaussian distribution until an image--or in this case--a molecule is generated.
+The idea we propose here is to effectively introduce information of a reference molecule at a given rate to the denoising process of a pre-trained model.
+This, similar to the concept of inpaining~\cite{qin2021image, xie2012image, squires2023artefact}, guides the denoising process at each step towards the reference.
+In this paper we show, that we can generate new compounds that are similar to a given reference and even link fragments using multiple references as input.
+The proposed method is generalizeable to any protein ligand systems with known fragment hits.
+For illustration purposes we use the original 23 X-ray fragment hits from the Covid-Moonshot dataset~\cite{consortium2021open,consortium2023open}.
+In the following, we give an overview of how equivariant difusion models work and introduce how our method SILVR fits into the framework of an existing pre-trained EDM.
+We then show, how the SILVR rate, $r_{\mathrm{S}}$ as modelling parameter can modulate how much of the reference is incorporated in the sample generation and how the original EDM is recovered at  $r_{\mathrm{S}}=0$.
+As part of this we illustrate how fragments can be linked using dummy atoms and how newly generated molecules can fit into the existing binding site according to shape similarity.
+Some of the limitations that arise from the proposed methods are connected to the chosen pre-trained EDM with which we conclude our discussions.
 

### 2023-03-18 paper/manuscript.tex

@@
 The idea we propose here is to effectively introduce information of a reference molecule at a given rate to the denoising process of a pre-trained model.
-This, similar to the concept of inpaining~\cite{qin2021image, xie2012image, squires2023artefact}, guides the denoising process at each step towards the reference.
+This, similar to the concept of inpaining~\cite{qin2021image, xie2012image, squires2023artefact}, or more precisely re-painting~\cite{lugmayr2022repaint}, guides the denoising process at each step towards the reference.
 In this paper we show, that we can generate new compounds that are similar to a given reference and even link fragments using multiple references as input.
@@
 You can rotate the input to the model then apply the diffusion/denoising process and get a structure out, or you can apply the diffusion/denoising process first and then do the same rotation to get the same output.
-Mathematically this means that if we have a set of point $\mathbf{x} = (\mathbf{x}_1,\ldots,\mathbf{x}_N) \in \mathbb{^{N\times 3}}$ and each of these points has a set of scalar feature vectors $\mathbf{h}\in \mathbb{R}^{N\times k}$ associated with it these features are invariant to group transformations.
+Mathematically this means that if we have a set of point $\mathbf{x} = (\mathbf{x}_1,\ldots,\mathbf{x}_N) \in \mathbb{R}^{N\times 3}$ and each of these points has a set of scalar feature vectors $\mathbf{h}\in \mathbb{R}^{N\times k}$ associated with it these features are invariant to group transformations.
 The positions translations and rotations are defined according to the orthogonal matrix: $\mathbf{Rx + t} = (\mathbf{R_{x_1}+t},\ldots, \mathbf{R_{x_N}+t})$.
@@
 
+\section{Summary of figures included in the supporting information}
 The supporting information consists of a series of figures in addition to the figures in the main paper.
@@
 The final figure~\ref{fig:uncurated_sample} shows a series of 2D samples from an uncurated list of samples, clearly showing examples of fragmented molecules.
-\paragraph{Dataset}
-Figure~\ref{fig:fragments} shows all MPro fragments used for this study from the covid moonshot dataset~\cite{consortium2021open,consortium2023open}.
+
+\section{Modified SILVR with dummy atoms}
+The SILVR protocol guides latent atoms according to a one-to-one mapping with reference atoms.
+This model is limited to reference molecules that somewhat overlap.
+Two fragments that are sufficiently far apart will not be successfully linked together.
+A universal EDM protocol for interpreting fragment data would ideally be able to do both fragment merging and linking.
+
+Dummy atoms are latent space atoms that do not have a mapping to a reference atom, and instead are free to explore the latent space unguided.
+It was hoped that in the case of disconnected fragments, these dummy atoms would be able to form a linker.
+In practice, linker formation was observed both with and without dummy atoms, however, the presence of dummy atoms seemed to produce better results (by visual inspection).
+It was also observed that dummy atoms often populate unfilled valences as hydrogen atoms.
+
+The total number of atoms used by the model was set as the sum reference (nr) and dummy atoms (nd).
+The SILVR vector was created such that the first nr indices held the value of rs defined in the experiment protocol, and the subsequent nd indices were set to zero so that SILVR would not be applied to the dummy atoms.
+
+The centre of gravity for the reference coordinates was initially aligned at zero.
+Within the denoising loop, the EDM sampled new coordinates for all atoms.
+The new coordinates for dummy atoms were then added to the reference coordinates, and the modified reference coordinates were re-aligned at zero.
+The SILVR equation was applied as described in the main paper, only refining the atoms mapped to a reference.
+Through each iteration of denoising, the coordinates of the dummy atoms within the reference were continually updated, and the centre of gravity of these coordinates was re-aligned at zero.
+The total shift in the centre of gravity of the reference was tracked and subtracted from the final sampled molecule.
 
@@
 Increasing SILVR rate results in sampled atom coordinates coming closer in space, and element type, to the reference while still resembling a truly molecular structure.}
-    \label{fig:fig_2}
+    \label{fig:examples_curated}
 \end{figure}
@@
     \centering
-    \includegraphics[width=0.8\textwidth]{paper/Figures/FigS3/FigS3.png}
-    \caption{Example looking at the displacement of reference atoms after diffusion. A: for the whole molecule. B: Zoom in, on the displacement of atoms in the 9aH-Pyrido[1,2-a]pyrazine heterocycle. }
-    \label{fig:atom_displacement}
+    \includegraphics[widht=0.5\textwidth]{paper/Figures/FigS3/FigS3.png}
+    \caption{Fraction of molecules fragmented with respect to the SILVR rate}
+    \label{fig:fragmentation}
 \end{figure}
@@
     \centering
-    \includegraphics[widht=0.5\textwidth]{paper/Figures/FigS4/FigS4.png}
-    \caption{Fraction of molecules fragmented with respect to the SILVR rate}
-    \label{fig:my_label}
+    \includegraphics[width=0.8\textwidth]{paper/Figures/FigS5/F.png}
+    \caption{Example looking at the displacement of reference atoms after diffusion. A: for the whole molecule. B: Zoom in, on the displacement of atoms in the 9aH-Pyrido[1,2-a]pyrazine heterocycle. }
+    \label{fig:atom_displacement}
 \end{figure}

### 2023-03-18 paper/manuscript.tex
First skim proofread

@@
 
-Typically the objective to achieve is to generate a diverse set of molecules, for a given target protein that are easily synthetically accessible and ideally with a high binding affinity and low toxicity prediction~\cite{some gneric caad}.
+Typically the objective is to generate a diverse set of molecules, for a given target protein that are easily synthetically accessible and ideally with a high binding affinity and low toxicity prediction~\cite{some gneric caad}.
 Different machine learning-based  and other strategies have been used to address these ideas.
-From Recurrent Neural networks (RNN)~\cite{xu2021novo,zhang2022novo} and conditional variational autoencoders~\cite{ragoza2022generating} to GANs~\cite{skalic2019target}, various approaches have been used to generate molecules for a binding site where in each case the training is conditioned based on the target protein~\cite{xie2022advances}.
-Some of these models incorporate a ligand score directly~\cite{ragoza2022generating}, while others would require other methods through machine learning~\cite{masters2023deep}, or more traditional docking~\cite{fu2018predictive} and molecular simulation-based affinity prediction methods~\cite{mey2016blinded,mey2020best}.
-Even with a variety of ways to assess the synthetic accessibility of generated compounds~\cite{coley2018scscore, thakkar2021retrosynthetic} often molecules generated in these method are  not easily sythesisable at best or chemically infeasible in the worst cases.
+From recurrent neural networks (RNN)~\cite{xu2021novo,zhang2022novo} and conditional variational autoencoders~\cite{ragoza2022generating} to GANs~\cite{skalic2019target}, various approaches have been used to generate molecules for a binding site where in each case the training is conditioned based on the target protein~\cite{xie2022advances}.
+Some of these models incorporate a ligand score directly~\cite{ragoza2022generating}, while others require other methods based on machine learning (ML)~\cite{masters2023deep}, or more traditional docking~\cite{fu2018predictive} and molecular simulation-based affinity prediction methods~\cite{mey2016blinded,mey2020best} for these tasks.
+Even with a variety of ways to assess the synthetic accessibility of generated compounds~\cite{coley2018scscore, thakkar2021retrosynthetic} often molecules generated with these method are  not easily sythesisable at best or chemically infeasible in the worst cases.
 Another desirable features for a generative model to have are the ability to generate compounds that are extensions to a known fragment hit or link two existing fragments through different linkers with the desire to improve potency, solubility or other properties of existing hits.
 Goa et al.
-have introduced way in which to generate linkers for fragments using reinforcement learning strategies~\cite{guo2023linkinvent}, while Imrie et al.~\cite{imrie2020deep} have used variation auto encoders on this tasks without the use of protein information.
-More recently Huang et al.~\cite{huang20223dlinker} and Igashov et al.~\cite{igashov2022equivariant} have tackled this challenge based on equivariant variational auto-encoder and diffusion based models models respectively.
+have introduced way in which to generate linkers for fragments using reinforcement learning strategies~\cite{guo2023linkinvent}, while Imrie et al.~\cite{imrie2020deep} have used variational auto encoders on this tasks without the use of protein information.
+More recently Huang et al.~\cite{huang20223dlinker} and Igashov et al.~\cite{igashov2022equivariant} have tackled this challenge based on equivariant variational auto-encoder and diffusion based models respectively.
 Each of these models were trained for the specific purpose of linker generation.
-Another objective from a medchem perspective is that once a hit has been found is to generate new compounds that improve this hit to become a good lead compound for optimisation.
 
-We present, the selective iterative variable refinement (SILVR), a method for conditioning an existing pre-trained equivaraint diffusion model (EDM) towards the task of linker generation between fragments and generating similar compounds to existing hits without specific training towards this task.
+In this paper, we present SILVR, a selective iterative variable refinement method for conditioning an existing pre-trained equivaraint diffusion model (EDM) towards the task of linker generation between fragments and generating similar compounds to existing hits without specific training towards this task.
 To achieve this we combine the EDM by Hoogeboom et al.~\cite{hoogeboom2022equivariant} with the iterative latent variable refinement method proposed by Choi et al.
@@
 
-Denoising probabilistic diffusion models can be separated into two parts, the diffusion part and the denoising part, as shown in the schematic in Figure~\ref{fig:schematic} with the ideas originating in image generation machine learning problems.
+Denoising probabilistic diffusion models can be separated into two parts, the diffusion part and the denoising part, as shown in the schematic in Figure~\ref{fig:schematic}.
+These ideas originated in image generation machine learning problems, but can also be applied to molecualr generation.
 A neural network is trained to learn the second part of the process, denoising a Gaussian distribution until an image--or in this case--a molecule is generated.
-The idea we propose here is to effectively introduce information of a reference molecule at a given rate to the denoising process of a pre-trained model.
-This, similar to the concept of inpaining~\cite{qin2021image, xie2012image, squires2023artefact}, or more precisely re-painting~\cite{lugmayr2022repaint}, guides the denoising process at each step towards the reference.
-In this paper we show, that we can generate new compounds that are similar to a given reference and even link fragments using multiple references as input.
-The proposed method is generalizeable to any protein ligand systems with known fragment hits.
-For illustration purposes we use the original 23 X-ray fragment hits from the Covid-Moonshot dataset~\cite{consortium2021open,consortium2023open}.
+The idea we propose is to effectively introduce information of a reference molecule at a given rate to the denoising process of a pre-trained model.
+This, similar to the concept of inpaining~\cite{qin2021image, xie2012image, squires2023artefact}, or more precisely re-painting~\cite{lugmayr2022repaint}, which guides the denoising process at each step towards the reference.
+In this paper we show, that we can generate new compounds that are similar to a given reference and even link fragments using multiple references as input with this method.
+This makes it generalizeable to any protein ligand systems with known fragment hits.
+For illustration purposes we use the original 23 X-ray fragment hits for the  SARS-CoV2 Main protease from the Covid-Moonshot dataset~\cite{consortium2021open,consortium2023open}.
 In the following, we give an overview of how equivariant difusion models work and introduce how our method SILVR fits into the framework of an existing pre-trained EDM.
 We then show, how the SILVR rate, $r_{\mathrm{S}}$ as modelling parameter can modulate how much of the reference is incorporated in the sample generation and how the original EDM is recovered at  $r_{\mathrm{S}}=0$.
-As part of this we illustrate how fragments can be linked using dummy atoms and how newly generated molecules can fit into the existing binding site according to shape similarity.
-Some of the limitations that arise from the proposed methods are connected to the chosen pre-trained EDM with which we conclude our discussions.
+As part of this, we illustrate how fragments can be linked using dummy atoms and how newly generated molecules can fit into the existing binding site according to shape similarity.
+Some of the limitations that arise from the proposed methods are connected to the chosen pre-trained EDM which we discuss, as well as making suggestions for future improvements of this model.
 
@@
 For this purpose, equivariant diffusion models (EDMs) are used to account for rotational invarance.
-For this purpose we chose to equivariant diffusion model (EDM) by Hoogeboom et al.~\cite{hoogeboom2022equivariant} as our baseline generative model, as it provides a generative model for new molecules and has equivariance already built into it.
+Therefore, we chose to the EDM by Hoogeboom et al.~\cite{hoogeboom2022equivariant} as our baseline generative model, as it provides a generative model for new molecules and has equivariance already built into it.
 The basic concept behind equivarience is that the model is invariant to rotations and translation in this case the E(3) group.
@@
     P_\theta(x_0|c) = \int p_\theta(x_{0:T}|c)Dx_{1:T}
+\end{equation}
+\begin{equation}
     p_{\theta}(x_{0:T}|c)=p(x_T)\prod_{t=1}^{T}p_{\theta}(x_{t-1}\x_t,c)
@@
 Figure~\ref{fig:silvr_explanation} shows an illustrative example of how the SILVR rate $r_{\mathrm{S}}$ is used to shift the latent space vector $\mathbf{z}_{t-1}$ at any point in the denoising process from $T\ldots t=1$.
-Using the silver equation as defined in the SILVR algorithm~\ref{alg:silvr} line 11, an existing denoising step (green) is brought closer to the reference (purple) in latent space.
+Using the SILVR equation as defined in the SILVR algorithm~\ref{alg:silvr} line 11, an existing denoising step (green) is brought closer to the reference (purple) in latent space.
 
@@
 The EDM by Hoodgeboom \textit{et al.} was trained on the 30 lowest energy conformations of 430,000 molecules from the GEOM dataset with explicit protons~\cite{axelrod2022geom}.
-A simplified schematic describing the EDM architecture is shared in scheme~\ref{fig:_architecture} for more details on the model theory and training refer to Hoogeboom et al.~\cite{hoogeboom2022equivariant}.
+A simplified schematic describing the EDM architecture is shared in scheme~\ref{fig:schematic} for more details on the model theory and training refer to Hoogeboom et al.~\cite{hoogeboom2022equivariant}.
 This model strictly only considers atomic coordinates; all bond information is ignored.
-A more recent version of the EDM also accounts for bond order and SILVR could also be incorporated into this new model~\cite{hvignac2023midi}.
+A more recent version of the EDM also accounts for bond order and SILVR could also be incorporated into this new model~\cite{vignac2023midi}.
 During training, the atomic coordinates (and their element) are passed through a forward diffusion process with iterative addition of Gaussian noise, where the extent of noise added at each step is defined by parameter $\beta$ (N.B.
@@
     \caption{Schematic illustrating the influence of SILVR of the molecules in the latent space.
-At each denoising step if $r_{mathrm{S}}>0$ the latent space vector get scaled by the SILVR rate.
+At each denoising step if $r_{\mathrm{S}}>0$ the latent space vector get scaled by the SILVR rate.
 This is done from the first denoising timestep (top) until the last one (bottom).
-Repeatedly applying SILVR will result in a molecule that resembles the references (right) and does not (left)}
+Repeatedly applying SILVR will result in a molecule that resembles the references for $r_{\mathrm{S}}>0$ (right) and does not for $r_{\mathrm{S}}=0$ (left)}
     \label{fig:silvr_explanation}
@@
 
-The EDM was then adapted by introducing SILVR within the denoising process algorithm~\ref{alg:silvr}, as outlined in~\ref{sec:silvr}.
+The EDM was adapted by introducing SILVR within the denoising process algorithm~\ref{alg:silvr}, as outlined previously.
 At run time, each atom of the reference set of coordinates is mapped to an atom in the EDM.
@@
 By introducing iterative refinement steps, the unconditional EDM can be guided to sample from a smaller region of chemical space that resembles the reference set of coordinates.
-Figure~\ref{fig:} demonstrates this architecture with the example of three disconnected fragments.
+Figure~\ref{fig:schematic} demonstrates this architecture with the example of three disconnected fragments.
 Here, the model generates a single 5-membered ring molecule with each atom maintaining the same element, however, notice each atom has drifted slightly from the reference.
 This is due to the competing effects of SILVR and EDM: the EDM tries to push atoms into a chemically reasonable position, while SILVR pulls atoms towards the reference.
-The resulting samples, therefore, therefore resemble both valid-looking molecules and the reference set of coordinates.
-The ability for reference atoms of fragments to move during generation distinguishes SILVR from standard linker design~\cite{linkerdesign1, hegeboom_linker}.
+The resulting samples, therefore, resemble both valid-looking molecules and the reference set of coordinates.
+The ability for reference atoms of fragments to move during generation distinguishes SILVR from standard linker design~\cite{guo2023linkinvent, igashov2022equivariant}.
 
@@
 Reference molecules were selected from the original 23 non-covalent hits of the SARS-CoV-2 Main Protease (Mpro) identified by the XChem fragment screen as part of the Covid Moonshot Project~\cite{consortium2023open, consortium2021open}.
-A more detailed picture of all fragements is presented in Figure S1 of the SI~\ref{si}.
+A more detailed picture of all fragments is presented in Figure~\ref{fig:moonshot_data} of the Supplementary information(SI)~\ref{si}.
 Fragments were visualised using NGLview version 3.03 and combinations were arbitrarily selected as test cases.
-Fragments \texttt{x0072} and \texttt{x0354} were selected for benchmarking the effect of $r_{\mathrm{S}}$ on sampling; \texttt{x1093}, \texttt{x0072}, \texttt{x2193} were selected to represent three significantly overlapping fragments; \texttt{x0434}, \texttt{x0305} and \texttt{x0072}, \texttt{x0354} were selected as partially overlapping fragments; and \texttt{x0874}, \texttt{x0397} were selected as two disconnected fragments, resembling a linker design type problem.
+Fragments \texttt{x0072} and \texttt{x0354} were selected for benchmarking the effect of $r_{\mathrm{S}}$ on sampling; \texttt{x1093}, \texttt{x0072}, \texttt{x2193} were used to represent three significantly overlapping fragments; \texttt{x0434}, \texttt{x0305} and \texttt{x0072}, \texttt{x0354} were used as partially overlapping fragments; and \texttt{x0874}, \texttt{x0397} were used as two disconnected fragments, resembling a linker design type problem.
 The bonding information of selected fragments was deleted and \texttt{XYZ} coordinates were combined into a single \texttt{XYZ} file.
@@
 
-Different observables were used to monitor how realistic and reliable newly generated molecules were and how well they can fit into the existing binding site of MPro.
+Different observables were used to monitor how realistic and reliable newly generated molecules were and how well they can fit into the existing binding site of Mpro.
 We looked at the following set of measures:
@@
 Once all bonds are defined, the valance of each atom is compared to its expected valance and if these values match, the atom is determined as stable.
-It should be noted that this metric requires the presence of all hydrogens (both explicit and implicit) for an atom to be decided as stable.
+It should be noted that this metric requires the presence of all hydrogens (both explicit and implicit) for an atom to be classed as stable.
 For comparability with other similar published models, this measure was used unmodified.
@@
 The protein receptor file was prepared from the Mpro-x0072 crystal structure with removal of the ligand.
-The XYZ coordinates of samples were directly read into OpenEye Tools and the pose was re-scored with \textit{shapegauss}.
+The \texttt{XYZ} coordinates of samples were directly read into OpenEye Tools and the pose was re-scored with \textit{shapegauss}.
 
@@
 As expected at no conditioning random samples are generated that do not resemble the reference fragments.
-At low values of $r_{\mathrm{S}}$ (< 0.0025) the sampled molecules only show an approximate agreement in orientation.
+At low values of $r_{\mathrm{S}}$ ($< 0.0025$) the sampled molecules only show an approximate agreement in orientation.
 At medium values of $r_{\mathrm{S}}$ ($0.0025 \le r_{\mathrm{S}} < 0.01$) the resulting samples begin to produce key structural features like ring systems and heteroatoms at positions seen in the reference.
@@
 To assess the stability and variability of generated molecules we looked at three different metrics, as discussed in the methods section~\ref{sec:methods}.
-We generated 1000 samples at different $r_{\mathrm{S}}$ from the two fragments \textit{X0072} and \textit{X0354} as a reference.
+We generated 1000 samples at different $r_{\mathrm{S}}$ from the two fragments \texttt{X0072} and \texttt{X0354} as a reference.
 Figure~\ref{fig:measures} summarises the findings from these experiments with box plots generated across the 1000 samples.
@@
     \caption{Validation measures of SILVR model using fragments \texttt{x0072} and \texttt{x0354} as reference coordinates.
-A) Ratio of stable atoms - an atom is determined as stable if the valence matches the expected valence for the element B) RMSD from reference - the calculated RMSD between the reference and sample, using an absolute one-to-one mapping ignore atom identity C) OpenEye measure “shapegauss” - a Gaussian scoring function describing the shape fit between Mpro and samples, ignoring chemical interactions D) Geometry stability - AIMNet geometry optimisation was completed with Auto3D using the SMILES string of each sample.
+A: Ratio of stable atoms - an atom is determined as stable if the valence matches the expected valence for the element B: RMSD from reference - the calculated RMSD between the reference and sample, using an absolute one-to-one mapping ignore atom identity C: OpenEye measure “shapegauss” - a Gaussian scoring function describing the shape fit between Mpro and samples, ignoring chemical interactions D: Geometry stability - AIMNet geometry optimisation was completed with Auto3D using the SMILES string of each sample.
 RMSD was calculated between the predicted geometry and the sampled geometry using RDKit.}
@@
 \end{figure}
-The similarity of samples to their reference set of coordinates was determined by RMSD, with a clear inverse correlation observed between rs and RMSD, as seen in Figure~\ref{fig:measures} b.
+The similarity of samples to their reference set of coordinates was determined by RMSD, with a clear inverse correlation observed between $r_{\mathrm{S}}$ and RMSD, as seen in Figure~\ref{fig:measures} B.
 This indicates that the extent of guidance of atoms towards the reference set of coordinates can be fine-tuned by varying $r_{\mathrm{S}}$.
@@
 For this purpose a separate geometry optimisation protocol was devised using the SMILES strings of the generated molecules and the RMSD between the generated molecule and the geometrically optimised molecule calculated.
-The results of this are found in Figure~\ref{fig:measures} d.
+The results of this are found in Figure~\ref{fig:measures} D.
 At low to medium values of $r_{\mathrm{S}}$ ($< 0.01$), the average RMSD values all fall in the $1.6 - 1.7$ \AA~  range, which is arguably good, however still leaves sufficient room for improvement.
@@
 \subsection{Generated samples with SILVR fit the binding site of Mpro}
-As one of the main motivators for SILVR is to be able to generate new molecules that fit directly into a binding site based on input fragements, we measured shape complementarity between newly generated samples and the Mpro binding site.
+As one of the main motivators for SILVR is to be able to generate new molecules that fit directly into a binding site based on input fragments, we measured shape complementarity between newly generated samples and the Mpro binding site.
 We used the OpenEye shapegauss scoring function for this purpose.
@@
 In addition to the experiments using two fragments and generating 1000 samples, we also looked at different combinations of fragments and resulting molecules.
-In general, the trends of Figure~\ref{fig:measures} were preserved for all experiments with slight overall variations.
+In general, the trends of Figure~\ref{fig:measures} were preserved for all experiments.
 Here we present three cases we investigated in detail.
@@
 
-Figures~\ref{fig:samples}A and~\ref{fig:samples}B demonstrate a superposition of three significantly overlapping fragments that result in generated molecules that fit the Mpro binding site well.
+Figures~\ref{fig:samples} A and~\ref{fig:samples} B demonstrate a superposition of three significantly overlapping fragments that result in generated molecules that fit the Mpro binding site well.
 Scrutinising sample A with $r_{\mathrm{S}}=0.005$, we can see the azaindole fused ring system has been interpreted as a pyrrole ring, the ketone transformed into an amide (maintaining the same carbonyl position), the sulfonyl group vanished, and the overlapping atoms have transformed into a fused ring system.
@@
 
-Figure~\ref{fig:samples}C shows a merger of two partially overlapping fragments with $r_{\mathrm{S}}=0.005$.
+Figure~\ref{fig:samples} C shows a merger of two partially overlapping fragments with $r_{\mathrm{S}}=0.005$.
 While the urea group was successfully preserved, the 6-membered ring shrank to a 5-membered heterocycle.
@@
 
-In contrast, Figure~\ref{fig:samples}D shows a stricter merging of two fragments, with $r_{\mathrm{S}}=0.01$.
+In contrast, Figure~\ref{fig:samples} D shows a stricter merging of two fragments, with $r_{\mathrm{S}}=0.01$.
 Visibly, the scaffold of the lower fragment has been maintained while the top fragment has contributed to a fused ring.
@@
 Being able to reliably link fragments that sit in a binding site of a protein is crucial for the design of potential new drugs.
-Being able to do so without retraining a model as was done by~\cite{huang20223dlinker}.
+Here we demonstrate how this can be done  without retraining and no training for the specific task of linker design~\cite{huang20223dlinker,imrie2020deep}.
 Using conditioning through SILVR allows the generation of linkers between fragments without the need to retrain the EDM.
 This is illustrated in the example of Figure~\ref{fig:samples}E and F.
-While it was possible to use SILVR as described in the Theory section~\ref{sec:tehory}, better results were obtained with the addition of \textit{dummy atoms}.
+While it was possible to use SILVR as described in the Theory section, better results were obtained with the addition of \textit{dummy atoms}.
 These are atoms which are present in the EDM without a mapping to a reference atom, and so are free to explore the whole coordinate space without guidance from SILVR.
@@
 
-\section{Discussion and Outlook}
+\section{Discussion and outlook}
 \label{sec:discussion}
@@
 The samples generated by SILVR are often of poorer quality than the samples selected in Figure~\ref{fig:samples}.
-Across all samples around half of the samples were determined by RDKit to be fragmented, meaning the sample contained two or more distinct molecular graphs.
+Across all samples around half of the samples were determined by RDKit to be fragmented, meaning the sample contained two or more distinct molecular graphs (See the uncurated list of samples in Figure~\ref{fig:uncurated_sample} in the SI).
 It was observed qualitatively that fragmented samples typically contained corrupted structures (multiple fragmentations, linear carbon chains, flattened rings, etc).
 We believe this fragmentation is triggered during intermediate steps of denoising, resulting in an unstable latent representation and subsequently poor EDM inference.
-Fragmentation becomes a particular issue for linker design type SILVR tasks (Figure~\ref{fig:samples}E and F), where the reference coordinates direct the latent variables away from each other, triggering fragmentation.
+Fragmentation becomes a particular issue for linker design type SILVR tasks (Figure~\ref{fig:samples} E and F), where the reference coordinates direct the latent variables away from each other, triggering fragmentation.
 For these experiments, 65\% of all samples were fragmented.
@@
 \subsection{There is no control over the retention of functional groups from the reference structure}
-When applied to a drug design context, the conservation of key functional groups in exact spacial positions is crucial for the maintenance of protein-ligand interactions.
+When applied in a drug-design context, the conservation of key functional groups in exact spacial positions is crucial to maintain desirable protein-ligand interactions.
 The series of molecules in Figure S2 shows a loss of the sulfonyl chloride group present in the reference, which may be undesirable.
@@
 The simple conditioning against a reference set of molecules means that the model can be used for tasks of fragment linker design, as well as generating new molecules that fit into an existing binding pocket without any specific training needed towards these tasks.
+It is also generalizable as it works on ligands with no information needed from the protein in its current form.
 In the future, given improvements in EDMs that produce more realistic and synthetically accessible molecules this method can cheaply generate structures exploring new chemical space with desired conditioning towards existing fragment hits.
@@
 Figure~\ref{fig:fragments} contains the 2D structures of all fragments used in the experiments as references.
-In Figure~\ref{fig:fig_2} an example of two curated experiments is shown with different SILVR rates $r_{mathrm{s}}$.
+In Figure~\ref{fig:examples_curated} an example of two curated experiments is shown with different SILVR rates $r_{mathrm{s}}$.
+Next, Figure~\ref{fig:fragments}, shows on average how many molecules were not-fragmented in a set of 1000 samples for different SILVR rates.
+Figure~\ref{fig:SA_QED} includes synthetical accessibility and QED.
 An example of how far atoms, even involved in a reference structure are displaced in the denoising diffusion process is shown in an example in Figure~\ref{fig:atom_displacement} \textbf{A} for a whole molecule generated and in \textbf{B} for the heterocycle.
-The next figure~\ref{fig:fragments} shows on average how many molecules were not-fragmented in a set of 1000 samples for different SILVR rates.
-The final figure~\ref{fig:uncurated_sample} shows a series of 2D samples from an uncurated list of samples, clearly showing examples of fragmented molecules.
+The final Figure~\ref{fig:uncurated_sample} shows a series of 2D samples from an uncurated list of samples, clearly showing examples of fragmented molecules.
 
@@
     \caption{2D Mpro structures from the moonshot dataset.}
-    \label{fig:fragments}
+    \label{fig:moonshot_data}
 \end{figure}
@@
     \centering
-    \includegraphics[widht=0.5\textwidth]{paper/Figures/FigS3/FigS3.png}
+    \includegraphics[widht=0.5\textwidth]{paper/Figures/FigS3/FigS4.png}
     \caption{Fraction of molecules fragmented with respect to the SILVR rate}
@@
     \centering
-    \includegraphics[width=0.8\textwidth]{paper/Figures/FigS5/F.png}
+    \includegraphics[widht=0.5\textwidth]{paper/Figures/FigS4/si_fig_2d_measures.png}
+    \caption{Write caption for SA and QED}
+    \label{fig:SA_QED}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.8\textwidth]{paper/Figures/FigS5/FigS3.png}
     \caption{Example looking at the displacement of reference atoms after diffusion. A: for the whole molecule. B: Zoom in, on the displacement of atoms in the 9aH-Pyrido[1,2-a]pyrazine heterocycle. }

### 2023-03-19 paper/manuscript.tex
Another proof read

@@
 \title[SILVR: Molecular Generation for binding modes]
-  {SILVR: Directed (Guided?) Diffusion for Molecule Generation}
+  {SILVR: Guided Diffusion for Molecule Generation}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \label{sec:intro}
-One of the challenges as part of computer-aided drug design is to sample from a very large space of possible drug-like compounds to find suitable hits for a given target protein.
-It is estimated that there are between $10^{23}$ to $10^{60}$ possibly feasible compounds, while only around $10^8$ have been synthesised so far~\cite{polishchuk2013estimation, reymond2012exploring}.
+One challenge  of computer-aided drug design is the ability to sample from a very large space of possible drug-like compounds to find suitable hits for a given target protein.
+It is estimated that there are between $10^{23}$ to $10^{60}$ possible feasible compounds, while only around $10^8$ have been synthesised so far~\cite{polishchuk2013estimation, reymond2012exploring}.
 Different strategies have been employed in the past to try and sample a diverse and synthetically accessible molecular space.
-In particular, methods based on machine learning have shown vast promise in recent years in this space~\cite{bilodeau2022generative}.
-Various different neural network architectures have been proposed for molecular generation, from Variational autoencoders (VAE)~\cite{kingma2022autoencoding, jin2018junction, ma2018constrained, agoza2022generating}, to generative adversarial networks (GAN)~\cite{hoffmann2019generating} and normalising flows~\cite{shi2021learning}.
-More recently denoising diffusion probabilistic models and particularly equivariant diffusion models have shown promise in molecular generation~\cite{xu2022geodiff, hoogeboom2022equivariant}.
+In particular, methods based on machine learning (ML) have shown vast promise in this space in recent years~\cite{bilodeau2022generative}.
+Various neural network architectures have been proposed for molecular generation, from Variational autoencoders (VAE)~\cite{kingma2022autoencoding, jin2018junction, ma2018constrained, agoza2022generating}, to generative adversarial networks (GAN)~\cite{hoffmann2019generating} and normalising flows~\cite{shi2021learning}.
+More recently denoising diffusion probabilistic models, and particularly equivariant diffusion models, have shown promise in molecular generation~\cite{xu2022geodiff, hoogeboom2022equivariant}.
+All of these were conceived primarily to generate new molecules, but being able to generate chemically varied molecules is only the first hurdle for identifying new drug candidates.
 
-Typically the objective is to generate a diverse set of molecules, for a given target protein that are easily synthetically accessible and ideally with a high binding affinity and low toxicity prediction~\cite{some gneric caad}.
+Typically the objective is to generate a diverse set of molecules for a given target protein that are easily synthetically accessible and ideally with a high binding affinity and low toxicity prediction~\cite{prieto-martinez2019chapter}.
 Different machine learning-based  and other strategies have been used to address these ideas.
@@
 
-In this paper, we present SILVR, a selective iterative variable refinement method for conditioning an existing pre-trained equivaraint diffusion model (EDM) towards the task of linker generation between fragments and generating similar compounds to existing hits without specific training towards this task.
+In this paper, we present SILVR, a selective iterative variable refinement (SILVR) method for conditioning an existing pre-trained equivariant diffusion model (EDM) towards the task of linker generation between fragments and generating similar compounds to existing hits without specific training towards this task.
 To achieve this we combine the EDM by Hoogeboom et al.~\cite{hoogeboom2022equivariant} with the iterative latent variable refinement method proposed by Choi et al.
@@
     \includegraphics[width=\textwidth]{paper/Figures/Fig1/Schematic.png}
-    \caption{Schematic of the equivariant diffusion model with  SILVR indicated for every denoising step.}
+    \caption{Schematic of the equivariant diffusion model with  SILVR indicated for every denoising step.
+Here a reference in blue on the left are 3 small fragments.
+They are evolved over time $\t$ in the diffusion process to resemble a Gaussian distribution at $t=T$.
+The $\beta$ added represents the noise added at each step and the multiple arrows in grey show the speed up from time $t=2$ to $t=T$.
+As atoms effectively 'diffuse' they can be preceived as changing position.
+To generate a new sample (left most yellow box), making use of the reference fragments, the EDM is sampled at each denoising step, as indicated by the EDM arrows, but now the reference at that same level of noise is added back in through SILVR at a given rate $r_{\mathbf{S}}$, until a new sample that resembles the reference is genearted (following the bottom row along the yellow arrow).}
     \label{fig:schematic}
@@
 A neural network is trained to learn the second part of the process, denoising a Gaussian distribution until an image--or in this case--a molecule is generated.
-The idea we propose is to effectively introduce information of a reference molecule at a given rate to the denoising process of a pre-trained model.
-This, similar to the concept of inpaining~\cite{qin2021image, xie2012image, squires2023artefact}, or more precisely re-painting~\cite{lugmayr2022repaint}, which guides the denoising process at each step towards the reference.
-In this paper we show, that we can generate new compounds that are similar to a given reference and even link fragments using multiple references as input with this method.
+The idea we propose is to introduce information of a reference molecule at a given rate to the denoising process of a pre-trained model.
+This is similar to the concept of inpaining~\cite{qin2021image, xie2012image, squires2023artefact}, or more precisely re-painting~\cite{lugmayr2022repaint}, which guides the denoising process at each step towards the reference.
+In this paper we show, that we can generate new compounds that are similar to a given reference and even link fragments using multiple references as input.
 This makes it generalizeable to any protein ligand systems with known fragment hits.
@@
 As part of this, we illustrate how fragments can be linked using dummy atoms and how newly generated molecules can fit into the existing binding site according to shape similarity.
-Some of the limitations that arise from the proposed methods are connected to the chosen pre-trained EDM which we discuss, as well as making suggestions for future improvements of this model.
 
@@
 \end{equation},
-for any $t>s$. The parameters $\alpha_t \in \mathbb{R}^{+}$ contains the amount of retained signal and $\sigma_t \in \mathbb{R}^+$ represents the variance and thus the amount of white noise added. 
+for any $t>s$. The parameters $\alpha_t \in \mathbb{R}^{+}$ contains the amount of retained signal and $\sigma_t \in \mathbb{R}^{+}$ represents the variance and thus the amount of white noise added. 
 Different people have looked at different noise schedules~\cite{sohl-dickstein2015deep, ho2020denoising}.
@@
 %%%%%%%
-\subsection*{Euquivariant diffusion model}
-Unlike in image problems, the orientation of molecules matter under translations and rotations.
-For this purpose, equivariant diffusion models (EDMs) are used to account for rotational invarance.
-Therefore, we chose to the EDM by Hoogeboom et al.~\cite{hoogeboom2022equivariant} as our baseline generative model, as it provides a generative model for new molecules and has equivariance already built into it.
+\subsection*{Equivariant diffusion model}
+Unlike in image problems, the orientation of molecules matters under translations and rotations.
+For this purpose, equivariant diffusion models (EDMs) are used to account for rotational invariance.
+Therefore, we chose the EDM by Hoogeboom et al.~\cite{hoogeboom2022equivariant} as our baseline generative model, as it provides a generative model for new molecules and has equivariance already built into it.
 The basic concept behind equivarience is that the model is invariant to rotations and translation in this case the E(3) group.
@@
 You can rotate the input to the model then apply the diffusion/denoising process and get a structure out, or you can apply the diffusion/denoising process first and then do the same rotation to get the same output.
-Mathematically this means that if we have a set of point $\mathbf{x} = (\mathbf{x}_1,\ldots,\mathbf{x}_N) \in \mathbb{R}^{N\times 3}$ and each of these points has a set of scalar feature vectors $\mathbf{h}\in \mathbb{R}^{N\times k}$ associated with it these features are invariant to group transformations.
+Mathematically this means that if we have a set of point $\mathbf{x} = (\mathbf{x}_1,\ldots,\mathbf{x}_N) \in \mathbb{R}^{N\times 3}$ and each of these points has an associated set of scalar feature vectors $\mathbf{h}\in \mathbb{R}^{N\times k}$ which are invariant to group transformations.
 The positions translations and rotations are defined according to the orthogonal matrix: $\mathbf{Rx + t} = (\mathbf{R_{x_1}+t},\ldots, \mathbf{R_{x_N}+t})$.
@@
 \label{sec:silvr}
-We propose a new method, that combines ILVR by Choi et al.~\cite{choi2021ilvr} proposed in an image generation context and incorporates the ideas into the EDM by Hoogeboom et al..
-This allows us to generate a selective iterative laten variable refinement (SILVR) procedure in which we introduce a reference molecule into the denoising step.
+We propose a new method, that combines ILVR by Choi et al.~\cite{choi2021ilvr} proposed in an image generation context and the EDM by Hoogeboom et al..
+This allows us to generate a selective iterative latent variable refinement (SILVR) procedure in which we introduce a reference molecule into the denoising step.
 The reference used can be a small fragment, multiple fragments or even include dummy atoms that do not exist in the original reference.
-As a result we propose a modification of Algorithm 2 by Hoogeboom et al.~\cite{hoogeboom2022equivariant} to result in the SILVR algorithm~\ref{alg:silvr}.
+As a result, we propose the following algorithm for SILVR~\ref{alg:silvr}.
 We are considering latent space variables $\mathbf{z}=[\mathbf{z}^{(x)},\mathbf{z}^{(h)}]$ for the standard denoising process and  $\mathbf{\hat{z}} = [\mathbf{\hat{z}}^{(y)},\mathbf{\hat{z}}^{(h)}]$, for the set of reference coordinates given by $\mathbf{y}$.
@@
 \subsection{Conditioning the pre-trained EDM with SILVR}
-The EDM by Hoodgeboom \textit{et al.} was trained on the 30 lowest energy conformations of 430,000 molecules from the GEOM dataset with explicit protons~\cite{axelrod2022geom}.
+The EDM by Hoodgeboom \textit{et al.} was trained on the 30 lowest energy conformations of 430,000 molecules from the GEOM dataset with explicit hydrogens~\cite{axelrod2022geom}.
 A simplified schematic describing the EDM architecture is shared in scheme~\ref{fig:schematic} for more details on the model theory and training refer to Hoogeboom et al.~\cite{hoogeboom2022equivariant}.
@@
     \caption{Schematic illustrating the influence of SILVR of the molecules in the latent space.
-At each denoising step if $r_{\mathrm{S}}>0$ the latent space vector get scaled by the SILVR rate.
+At each denoising step if $r_{\mathrm{S}}>0$ the latent space vector gets scaled by the SILVR rate.
 This is done from the first denoising timestep (top) until the last one (bottom).
@@
 That is, the amount of structure remaining from the reference should match the amount of structure formed by the generative process.
-A small refinement is applied to add information from the reference coordinates to the latent variable of the denoising process~\ref{eq:denoising}.
-This equation has the effect of shrinking the coordinates towards the origin, and then expanding the coordinates out in the direction of the reference, see Figure~\ref{fig:silvr_explanation}.
+A small refinement is applied to add information from the reference coordinates to the latent variable of the denoising process.
+This equation has the effect of shrinking the coordinates towards the origin and then expanding the coordinates out in the direction of the reference, see Figure~\ref{fig:silvr_explanation}.
 Importantly, the extent of this refinement is defined by the SILVR rate $r_{\mathrm{S}}$, with $r_{\mathrm{S}}=0$ providing no additional refinement and $r_{\mathrm{S}}=1$ will result in a total replacement of atoms.
@@
 Reference molecules were selected from the original 23 non-covalent hits of the SARS-CoV-2 Main Protease (Mpro) identified by the XChem fragment screen as part of the Covid Moonshot Project~\cite{consortium2023open, consortium2021open}.
-A more detailed picture of all fragments is presented in Figure~\ref{fig:moonshot_data} of the Supplementary information(SI)~\ref{si}.
-Fragments were visualised using NGLview version 3.03 and combinations were arbitrarily selected as test cases.
-Fragments \texttt{x0072} and \texttt{x0354} were selected for benchmarking the effect of $r_{\mathrm{S}}$ on sampling; \texttt{x1093}, \texttt{x0072}, \texttt{x2193} were used to represent three significantly overlapping fragments; \texttt{x0434}, \texttt{x0305} and \texttt{x0072}, \texttt{x0354} were used as partially overlapping fragments; and \texttt{x0874}, \texttt{x0397} were used as two disconnected fragments, resembling a linker design type problem.
+A more detailed picture of all fragments is presented in FigureS~\ref{fig:moonshot_data} of the Supplementary information(SI).
+Fragments were visualised using NGLview version 3.03 and combinations were arbitrarily selected as test cases for the following different experimental setups beyond trying to understand similarity between refernce and new sample.
+We looked at the following scenarios:
+\begin{enumerate}
+    \item Using three distinct fragments with substantial overlap as a reference to generate new samples
+    \item Using two fragments with a slight overlap to generate a single new sample.
+    \item And using two fragments that are disconnected to investigate linker generation
+\end{enumerate}
+Fragments \texttt{x0072} and \texttt{x0354} were selected for benchmarking the effect of $r_{\mathrm{S}}$ on sampling; \texttt{x1093}, \texttt{x0072}, \texttt{x2193} were randomly chosen to represent three significantly overlapping fragments; \texttt{x0434}, \texttt{x0305} and \texttt{x0072}, \texttt{x0354} were used as partially overlapping fragments; and \texttt{x0874}, \texttt{x0397} were used as two disconnected fragments, resembling a linker design type problem.
 The bonding information of selected fragments was deleted and \texttt{XYZ} coordinates were combined into a single \texttt{XYZ} file.
@@
 
-\subsection{Different observables were monitored to gauge the success of computational experiments for the generation of new models}
+\subsection{Different observables were used for the performance assessment of SILVR}
 
@@
 This metric infers bonds and bond orders between atoms by considering their interatomic distances.
-Once all bonds are defined, the valance of each atom is compared to its expected valance and if these values match, the atom is determined as stable.
+Once all bonds are defined, the valence of each atom is compared to its expected valence and if these values match, the atom is determined as stable.
 It should be noted that this metric requires the presence of all hydrogens (both explicit and implicit) for an atom to be classed as stable.
@@
 \subsubsection{Shape similarity of sample to protein}
-The agreement in the shape of samples and the main protease binding site was determined using OpenEye Tools Gaussian scoring function \textit{Shapegauss}.
+The agreement in the shape of samples and the main protease binding site was determined using openeye-toolkits version 2022.2.2 Gaussian scoring function \textit{Shapegauss}~\cite{openeyescientificsoftwareinc.oedocking,kelley2015posit}.
 This scoring function measures the shape similarity between the ligand and receptor by considering each heavy atom as a Gaussian function~\cite{mcgann2003gaussian}.
@@
 The protein receptor file was prepared from the Mpro-x0072 crystal structure with removal of the ligand.
-The \texttt{XYZ} coordinates of samples were directly read into OpenEye Tools and the pose was re-scored with \textit{shapegauss}.
+The \texttt{XYZ} coordinates of samples were directly read into openeye-toolkit and the pose was re-scored with \textit{Shapegauss}.
 
@@
 In the following, we will demonstrate how SILVR can be used to generate conditioned samples to a reference using a pre-trained EDM without additional training.
-Some of the flaws in the presented results stem from the underlying EDM, but retraining an improved EDM is out of the scope of this work.
 The main questions we set out to answer with SILVR were:
@@
 \subsection{The SILVR rate $r_{\mathrm{S}}$ effectively modulates similarity to the reference structures}
-Qualitatively, the generated molecular samples from the recondition EDM  show a clear resemblance to their reference structures, with similarity increasing with $r_{\mathrm{S}}$.
-Figure S2, in the SI, shows two example samples started from fragments xxx and xxx over a range of SILVR rates between $r_{\mathrm{S}}=0$ to $r_{\mathrm{S}}=0.02$.
+Qualitatively, the generated molecular samples from the condition EDM  show a clear resemblance to their reference structures, with similarity increasing with $r_{\mathrm{S}}$.
+Figure S~\ref{fig:examples_curated}, in the SI, shows two example samples started from fragments \texttt{x0072} and \texttt{x0354} over a range of SILVR rates between $r_{\mathrm{S}}=0$ to $r_{\mathrm{S}}=0.02$.
 As expected at no conditioning random samples are generated that do not resemble the reference fragments.
@@
 At very high values of $r_{\mathrm{S}}$ ($> 0.02$) there is a very high similarity between samples and the reference, however, most structures no longer resemble valid molecules.
-It seems best molecules are formed at intermediate values of $r_{\mathrm{S}}$ ($0.05 < r_{\mathrm{S}} < 0.01$) offering a trade-off between similarity to the reference, sampling diversity, and molecular likeness.
+Best molecules are formed at intermediate values of $r_{\mathrm{S}}$ ($0.05 < r_{\mathrm{S}} < 0.01$) offering a trade-off between similarity to the reference, sampling diversity, and molecular likeness.
 This is further validated by looking at stability measures.
@@
 \subsection{Intermediate SILVR rates produce stable and varied molecules}
-To assess the stability and variability of generated molecules we looked at three different metrics, as discussed in the methods section~\ref{sec:methods}.
-We generated 1000 samples at different $r_{\mathrm{S}}$ from the two fragments \texttt{X0072} and \texttt{X0354} as a reference.
+To assess the stability and variability of generated molecules we looked at three different metrics, as discussed in the methods section.
+We generated 1000 samples at different $r_{\mathrm{S}}$ using fragments \texttt{X0072} and \texttt{X0354} as a reference.
 Figure~\ref{fig:measures} summarises the findings from these experiments with box plots generated across the 1000 samples.
@@
 are not a single connected molecular graph or not, with respect to $r_{\mathrm{S}}$.
-This is presented in Figure S4 in the SI.
+This is presented in Figure S\ref{fig:fragmentation} in the SI.
 At an intermediate $r_{\mathrm{S}} = 0.025$ just over 50\% of the generated samples are not fragmented meaning that in the worst-case scenario, only one in two generated molecules can be analysed further.
 The subsequent analysis is carried out only on whole molecular graphs.
-Figure~\ref{fig:measures} a looks at the \textit{atom stability} measure as introduced by Hoogeboom et al.~\cite{hoogeboom2022equivariant}.
+Figure~\ref{fig:measures} A looks at the \textit{atom stability} measure as introduced by Hoogeboom et al.~\cite{hoogeboom2022equivariant}.
 Samples generated at low $r_{\mathrm{S}}$ tend to have similar stabilities to each other, samples start becoming less stable around $r_{\mathrm{S}}=0.005$, and become totally unstable at $r_{\mathrm{S}}=0.02$.
-This trend can largely be explained due to issues around protons.
-The atom stability measure calculates whether the valance of each atom matches what is expected for that atom, however, the measure requires the presence of explicit protons.
-A carbon skeleton with appropriate C-C bond lengths would be determined as \textit{unstable} unless each carbon was populated with explicit protons.
+This trend can largely be explained due to issues around hydrogens.
+The atom stability measure calculates whether the valance of each atom matches what is expected for that atom, however, the measure requires the presence of explicit hydrogens.
+A carbon skeleton with appropriate C-C bond lengths would be determined as \textit{unstable} unless each carbon was populated with explicit hydrogens.
 In the case of high $r_{\mathrm{S}}$ values, the SILVR method pulls atom types strongly towards the reference.
-Since there are no protons in the reference, all atoms are mapped to heavy atoms, and therefore most atoms are unable to satisfy a full valence.
-Adding protons explicitly to the molecules, through OpenBable or RDkit, is a way of improving this measure.
+Since there are no hydrogens in the reference, all atoms are mapped to heavy atoms, and therefore most atoms are unable to satisfy a full valence.
+Adding hydrogens explicitly to the molecules, through OpenBable or RDkit, is a way of improving this measure.
 
@@
 The results of this are found in Figure~\ref{fig:measures} D.
-At low to medium values of $r_{\mathrm{S}}$ ($< 0.01$), the average RMSD values all fall in the $1.6 - 1.7$ \AA~  range, which is arguably good, however still leaves sufficient room for improvement.
+At low to medium values of $r_{\mathrm{S}}$ ($< 0.01$), the average RMSD values all fall in the $1.6 - 1.7$ \AA~  range.
 Importantly, no difference is seen between the control set ($r_{\mathrm{S}}=0$) and the SILVR samples ($0 < r_{\mathrm{S}} < 0.01$) indicating the quality of generated molecules was not impeded by the SILVR protocol.
@@
 As one of the main motivators for SILVR is to be able to generate new molecules that fit directly into a binding site based on input fragments, we measured shape complementarity between newly generated samples and the Mpro binding site.
-We used the OpenEye shapegauss scoring function for this purpose.
-From Figure~\ref{fig:measures} c it can be seen that the shape complementarity of samples improves with increasing $r_{\mathrm{S}}$, demonstrating that SILVR can produce ligands of binding site geometry when guided by the coordinates of fragment molecules.
+We used the OpenEye \textit{Shapegauss} scoring function for this purpose~\cite{kelley2015posit}.
+From Figure~\ref{fig:measures} C it can be seen that the shape complementarity of samples improves with increasing $r_{\mathrm{S}}$, demonstrating that SILVR can produce ligands of binding site geometry when guided by the coordinates of fragment molecules.
 The lack of plot data for $r_{\mathrm{S}}=0.02$ and $r_{\mathrm{S}}=0.03$ was because all shapegauss calculations failed.
@@
 In general, the trends of Figure~\ref{fig:measures} were preserved for all experiments.
-Here we present three cases we investigated in detail.
-\begin{enumerate}
-    \item Using three fragments with substantial overlap as a reference.
-    \item Using two fragments with a slight overlap. 
-    \item Using two fragments that are disconnected to investigate linker generation. This scenario will be presented in detail in section~\ref{sec:linker}.
-\end{enumerate}
+In the follow we present three cases we investigated in detail, the test case using 3 fragments as a reference with substantial overlap, using two fragments with some overlap, and two disconnected fragments for the linker generation experiment.
 
@@
 Of particular interest is the formation of the fused ring system.
-At first glance, it might be assumed that reference atoms map to the sample atom closest in space, however in actuality they travel up to 1.7 \AA~to arrive at their final position (SI figure 4).
-In this case, the nitrogen atoms observed in the fused ring are directly obtained from the nitrogen atoms in the reference, however, their final position is one bond’s length from their reference.
+At first glance, it might be assumed that reference atoms map to the sample atom closest in space, however in actuality they travel up to 1.7 \AA~to arrive at their final position (Figure S\ref{fig:atom_displacement}) In this case, the nitrogen atoms observed in the fused ring are directly obtained from the nitrogen atoms in the reference, however, their final position is one bond’s length from their reference.
 This shows the flexibility of each sample atom to explore within a radius (defined by $r_{\mathrm{S}}$) of the reference atom.
@@
 \label{sec:discussion}
-SILVR, as presented, represents a promising proof of concept method in which a general EDM can be conditioned to generate samples that resemble a reference structure without any additional training needed.
-We could show that SILVR can complete both fragment merging and linking type tasks, without any \textit{a priori} knowledge of these design challenges.
+SILVR, as presented, represents method in which a general EDM can be conditioned to generate samples that resemble a reference structure without any additional training needed.
+We showed that SILVR can complete both fragment merging and linking type tasks, without any \textit{a priori} knowledge of these design challenges.
 Considering all results with respect to the control EDM ($r_{\mathrm{S}}=0.0$), we show that at intermediate values of $r_{\mathrm{S}}$ the SILVR protocol produces molecules of equal quality to that of the unmodified EDM, while also guiding molecules towards reference structures.
@@
 Next, Figure~\ref{fig:fragments}, shows on average how many molecules were not-fragmented in a set of 1000 samples for different SILVR rates.
-Figure~\ref{fig:SA_QED} includes synthetical accessibility and QED.
+Figure~\ref{fig:SA_QED} includes synthetical accessibility(~\cite{ertl2009estimation}) and QED~\cite{bickerton2012quantifying, wildman1999prediction}.
 An example of how far atoms, even involved in a reference structure are displaced in the denoising diffusion process is shown in an example in Figure~\ref{fig:atom_displacement} \textbf{A} for a whole molecule generated and in \textbf{B} for the heterocycle.
@@
     \centering
-    \includegraphics[widht=0.5\textwidth]{paper/Figures/FigS4/si_fig_2d_measures.png}
+    \includegraphics[width=0.8\textwidth]{paper/Figures/FigS4/si_fig_2d_measures.png}
     \caption{Write caption for SA and QED}
@@
     \centering
-    \includegraphics[width=0.8\textwidth]{paper/Figures/FigS5/mols_grid_silvr_rate_not_curated.png}
+    \includegraphics[width=0.8\textwidth]{paper/Figures/FigS6/mols_grid_silvr_rate_not_curated.png}
     \caption{Uncurated samples, showing fragmentation of molecules based on input with two starting structures.}

### 2023-03-19 paper/manuscript.tex

@@
 More recently denoising diffusion probabilistic models, and particularly equivariant diffusion models, have shown promise in molecular generation~\cite{xu2022geodiff, hoogeboom2022equivariant}.
-All of these were conceived primarily to generate new molecules, but being able to generate chemically varied molecules is only the first hurdle for identifying new drug candidates.
+All of these were conceived primarily to generate new molecules, however, being able to generate chemically varied molecules is only the first hurdle for identifying new drug candidates.
 
-Typically the objective is to generate a diverse set of molecules for a given target protein that are easily synthetically accessible and ideally with a high binding affinity and low toxicity prediction~\cite{prieto-martinez2019chapter}.
-Different machine learning-based  and other strategies have been used to address these ideas.
+Typically the objective is to generate a diverse set of molecules for a given target protein that are easily synthetically accessible, and ideally with high binding affinity and low toxicity prediction~\cite{prieto-martinez2019chapter}.
+Different machine learning-based, and non-machine-learning-based, strategies have been used to address these ideas.
 From recurrent neural networks (RNN)~\cite{xu2021novo,zhang2022novo} and conditional variational autoencoders~\cite{ragoza2022generating} to GANs~\cite{skalic2019target}, various approaches have been used to generate molecules for a binding site where in each case the training is conditioned based on the target protein~\cite{xie2022advances}.
@@
     \caption{Validation measures of SILVR model using fragments \texttt{x0072} and \texttt{x0354} as reference coordinates.
-A: Ratio of stable atoms - an atom is determined as stable if the valence matches the expected valence for the element B: RMSD from reference - the calculated RMSD between the reference and sample, using an absolute one-to-one mapping ignore atom identity C: OpenEye measure “shapegauss” - a Gaussian scoring function describing the shape fit between Mpro and samples, ignoring chemical interactions D: Geometry stability - AIMNet geometry optimisation was completed with Auto3D using the SMILES string of each sample.
-RMSD was calculated between the predicted geometry and the sampled geometry using RDKit.}
+A: Ratio of stable atoms - an atom is determined as stable if the valence matches the expected valence for the element B: RMSD from reference - the calculated RMSD between the reference and sample, using an absolute one-to-one mapping ignore atom identity C: OpenEye measure \texit{Shapegauss} - a Gaussian scoring function describing the shape fit between Mpro and samples, ignoring chemical interactions D: Geometry stability - AIMNet geometry optimisation was completed with Auto3D using the SMILES string of each sample.
+RMSD was calculated between the predicted geometry and the sampled geometry using RDKit.
+Lines indicated the sample median and circles the sample mean.}
     \label{fig:measures}

### 2023-03-19 paper/manuscript.tex
nicholas readthrough

@@
 \usepackage{hyperref}
+\usepackage{amssymb}
 
@@
 One challenge  of computer-aided drug design is the ability to sample from a very large space of possible drug-like compounds to find suitable hits for a given target protein.
-It is estimated that there are between $10^{23}$ to $10^{60}$ possible feasible compounds, while only around $10^8$ have been synthesised so far~\cite{polishchuk2013estimation, reymond2012exploring}.
+It is estimated that there are between $10^{23}$ to $10^{60}$ feasible compounds, while only around $10^8$ have been synthesised so far~\cite{polishchuk2013estimation, reymond2012exploring}.
 Different strategies have been employed in the past to try and sample a diverse and synthetically accessible molecular space.
@@
 From recurrent neural networks (RNN)~\cite{xu2021novo,zhang2022novo} and conditional variational autoencoders~\cite{ragoza2022generating} to GANs~\cite{skalic2019target}, various approaches have been used to generate molecules for a binding site where in each case the training is conditioned based on the target protein~\cite{xie2022advances}.
-Some of these models incorporate a ligand score directly~\cite{ragoza2022generating}, while others require other methods based on machine learning (ML)~\cite{masters2023deep}, or more traditional docking~\cite{fu2018predictive} and molecular simulation-based affinity prediction methods~\cite{mey2016blinded,mey2020best} for these tasks.
-Even with a variety of ways to assess the synthetic accessibility of generated compounds~\cite{coley2018scscore, thakkar2021retrosynthetic} often molecules generated with these method are  not easily sythesisable at best or chemically infeasible in the worst cases.
-Another desirable features for a generative model to have are the ability to generate compounds that are extensions to a known fragment hit or link two existing fragments through different linkers with the desire to improve potency, solubility or other properties of existing hits.
+Some of these models incorporate a ligand score directly~\cite{ragoza2022generating}, while others require methods based on machine learning (ML)~\cite{masters2023deep}, or more traditional docking~\cite{fu2018predictive} and molecular simulation-based affinity prediction methods~\cite{mey2016blinded,mey2020best}.
+Even with a variety of ways to assess the synthetic accessibility of generated compounds~\cite{coley2018scscore, thakkar2021retrosynthetic}, molecules generated with these methods are often not easily synthesizable and in the worst cases can be chemically infeasible.
+A desirable feature for a generative model for drug design is the ability to generate compounds that are extensions of known fragment hits, or connect two existing fragments through "linkers", with the aim of making a new compound with improved pharmacochemical properties such as potency or solubility.
 Goa et al.
-have introduced way in which to generate linkers for fragments using reinforcement learning strategies~\cite{guo2023linkinvent}, while Imrie et al.~\cite{imrie2020deep} have used variational auto encoders on this tasks without the use of protein information.
-More recently Huang et al.~\cite{huang20223dlinker} and Igashov et al.~\cite{igashov2022equivariant} have tackled this challenge based on equivariant variational auto-encoder and diffusion based models respectively.
-Each of these models were trained for the specific purpose of linker generation.
+have introduced a way to generate linkers for fragments using reinforcement learning strategies~\cite{guo2023linkinvent}, while Imrie et al.~\cite{imrie2020deep} have used variational auto encoders on this task without the use of protein information.
+More recently Huang et al.~\cite{huang20223dlinker} and Igashov et al.~\cite{igashov2022equivariant} have tackled this challenge using equivariant variational auto-encoder and diffusion based models respectively.
+Each of these models were explicitly trained for the specific purpose of linker generation.
 
-In this paper, we present SILVR, a selective iterative variable refinement (SILVR) method for conditioning an existing pre-trained equivariant diffusion model (EDM) towards the task of linker generation between fragments and generating similar compounds to existing hits without specific training towards this task.
+In this paper, we present SILVR, a selective iterative variable refinement (SILVR) method for conditioning an existing pre-trained equivariant diffusion model (EDM) towards the task of fragment merging and linker generation, yielding compounds similar to existing hits without specific training towards this task.
 To achieve this we combine the EDM by Hoogeboom et al.~\cite{hoogeboom2022equivariant} with the iterative latent variable refinement method proposed by Choi et al.
 for image generation networks~\cite{choi2021ilvr}.
-This allows the generation of new molecules in the shape of a binding site using information from existing fragments or linking fragments without specific training or knowledge of this task.
+This allows the generation of new molecules in the shape of a binding site using information from existing fragment hits, without specific training or knowledge of this task.
 
@@
     \caption{Schematic of the equivariant diffusion model with  SILVR indicated for every denoising step.
-Here a reference in blue on the left are 3 small fragments.
-They are evolved over time $\t$ in the diffusion process to resemble a Gaussian distribution at $t=T$.
-The $\beta$ added represents the noise added at each step and the multiple arrows in grey show the speed up from time $t=2$ to $t=T$.
-As atoms effectively 'diffuse' they can be preceived as changing position.
-To generate a new sample (left most yellow box), making use of the reference fragments, the EDM is sampled at each denoising step, as indicated by the EDM arrows, but now the reference at that same level of noise is added back in through SILVR at a given rate $r_{\mathbf{S}}$, until a new sample that resembles the reference is genearted (following the bottom row along the yellow arrow).}
+Here the reference in blue on the left is 3 small fragments.
+They are evolved over time $t$ in the diffusion process to resemble a Gaussian distribution at $t=T$.
+The $\beta$ added represents the noise added at each step, and the multiple arrows in grey show the speed up from time $t=2$ to $t=T$.
+As atoms effectively 'diffuse' they can be perceived as changing position.
+To generate a new sample (left most yellow box), making use of the reference fragments, the EDM is sampled at each denoising step, as indicated by the EDM arrows, but now the reference at that same level of noise is added back in through SILVR at a given rate $r_{\mathbf{S}}$, until a new sample that resembles the reference is generated (following the bottom row along the yellow arrow).}
     \label{fig:schematic}
@@
 
-Denoising probabilistic diffusion models can be separated into two parts, the diffusion part and the denoising part, as shown in the schematic in Figure~\ref{fig:schematic}.
-These ideas originated in image generation machine learning problems, but can also be applied to molecualr generation.
+Denoising probabilistic diffusion models can be separated into two parts, the diffusion process and the denoising process, as shown in the schematic in Figure~\ref{fig:schematic}.
+These ideas originated in image generation machine learning problems, but can also be applied to molecular generation.
 A neural network is trained to learn the second part of the process, denoising a Gaussian distribution until an image--or in this case--a molecule is generated.
@@
 This is similar to the concept of inpaining~\cite{qin2021image, xie2012image, squires2023artefact}, or more precisely re-painting~\cite{lugmayr2022repaint}, which guides the denoising process at each step towards the reference.
-In this paper we show, that we can generate new compounds that are similar to a given reference and even link fragments using multiple references as input.
-This makes it generalizeable to any protein ligand systems with known fragment hits.
+In this paper we show that we can generate new compounds that are similar to a given reference and even link fragments using multiple references as input.
+This makes it generalizable to any protein-ligand systems with known fragment hits.
 For illustration purposes we use the original 23 X-ray fragment hits for the  SARS-CoV2 Main protease from the Covid-Moonshot dataset~\cite{consortium2021open,consortium2023open}.
 In the following, we give an overview of how equivariant difusion models work and introduce how our method SILVR fits into the framework of an existing pre-trained EDM.
-We then show, how the SILVR rate, $r_{\mathrm{S}}$ as modelling parameter can modulate how much of the reference is incorporated in the sample generation and how the original EDM is recovered at  $r_{\mathrm{S}}=0$.
+We then show, how the SILVR rate, $r_{\mathrm{S}}$ as a modelling parameter can modulate how much of the reference is incorporated in the sample generation and how the original EDM is recovered at  $r_{\mathrm{S}}=0$.
 As part of this, we illustrate how fragments can be linked using dummy atoms and how newly generated molecules can fit into the existing binding site according to shape similarity.
@@
 DDPMs are often used as generative models that were developed for the generation of new images~\cite{sohl-dickstein2015deep, ho2020denoising, nichol2021improved}.
-More recently the same idea has also been applied to molecular generation~\cite{hoogeboom2022equivariant}.
-The main idea behind diffusion models is for a neural network to learn the reverse of a diffusion process, often referred to as denoising to \textit{sample} a new image or in our case a new molecule.
-In practice this is done by training a neural network $\phi$ and generating samples $p(\mathbf{x_{t-1}}|\mathbf{x}_t)=\mathcal(x_{t-1}; \{\mu_{\theta}(x_t;t), \sigma_t^2\mathbf{I}$ from  a Gaussian transition with learned mean ($\mu_{\theta}(x_t;t)$) and variance ($\sigma_t$) with $\mathbf{x}_t$ with data $X$, noised up to time $t$.
-Figure~\ref{fig:diffusion_schematic} shows a schematic of the two main parts of such a DDPM, the diffusion part where noise is added for each timestep (shown in blue) and the denoising part (shown in yellow) that allows the generation of a new image (molecule).
+More recently, the same idea has also been applied to molecular generation~\cite{hoogeboom2022equivariant}.
+The main idea behind diffusion models is for a neural network to learn the reverse of a diffusion process, often referred to as denoising, to \textit{sample} a new image or in our case a new molecule.
+In practice, this is done by training a neural network $\phi$ and generating samples $p(\mathbf{x}_{t-1}|\mathbf{x}_t)=\mathcal{N}(x_{t-1}; \mu_{\theta}(x_t;t), \sigma_t^2\mathbf{I})$ from  a Gaussian transition with learned mean ($\mu_{\theta}(x_t;t)$) and variance ($\sigma_t$) with $\mathbf{x}_t$ being the data, noised up to time $t$.
+Figure~\ref{fig:schematic} shows a schematic of the two main parts of such a DDPM, the diffusion part where noise is added for each timestep (shown in blue) and the denoising part (shown in yellow) that allows the generation of a new image (molecule).
 
-The \textit{forward} part of the diffusion model is a Markov process, where you start out with a set of data $\mathbf{x}$ and add noise to this data over a time interval t = [0,\ldots,T] according to the following distribution:
+The \textit{forward} part of the diffusion model is a Markov process: noise is added to a set of data $\mathbf{x}$ and over a time interval $t = [0,\ldots,T]$ according to the following distribution:
 \begin{equation}
-  q(\mathbf{x_{1:T}})|\mathbf{x_0} = q(x_0)\prod_{t=1}^T q(\mathbf{x}_t|\mathbf{x}_{t-1})
-\end{equation}.
+  q(\mathbf{x}_{1:T}|\mathbf{x_0}) = q(x_0)\prod_{t=1}^T q(\mathbf{x}_t|\mathbf{x}_{t-1}).
+\end{equation}
 The product of conditional probabilities $q(..)$ can be modelled as a Gaussian Kernel given by:
 \begin{equation}
-    q(x_t, x_s)=\mathcal{N}(x_t|\alpha_t/\alpha_s, \sigma_t^2-\frac{\alpha_t}{\alpha_s}\sigma^2_s
-\end{equation},
-for any $t>s$. The parameters $\alpha_t \in \mathbb{R}^{+}$ contains the amount of retained signal and $\sigma_t \in \mathbb{R}^{+}$ represents the variance and thus the amount of white noise added. 
+    q(\mathbf{x}_t, \mathbf{x}_s)=\mathcal{N}(x_t|\frac{\alpha_t}{\alpha_s}, \sigma_t^2-\frac{\alpha_t}{\alpha_s}\sigma^2_s),
+\end{equation}
+for any $t>s$.
+The parameters $\alpha_t \in \mathbb{R}^{+}$ contains the amount of retained signal and $\sigma_t \in \mathbb{R}^{+}$ represents the variance and thus the amount of white noise added.
+The parameter $\alpha_t$ also directly relates to $\beta$ in Figure~\ref{fig:schematic} with $\alpha_t:=1-\beta_t$ and $\langle\alpha\rangle_t:=\prod_{s=1}^t\alpha_s$.
+$\beta_t$ is a fixed variance schedule which adds noise with each timestep $t$.
 Different people have looked at different noise schedules~\cite{sohl-dickstein2015deep, ho2020denoising}.
@@
 \begin{equation}
-p(\mathbf{x_{t-1}}|\mathbf{x}_t)=\mathcal(x_{t-1}; \{\mu_{\theta}(x_t;t), \sigma_t^2\mathbf{I}.
+p_{\theta}(\mathbf{x}_{t-1}|\mathbf{x}_t)=\mathcal{N}(x_{t-1}; \mu_{\theta}(x_t;t), \sigma_t^2\mathbf{I}).
 \end{equation}
@@
 \begin{equation}
- x_{t-1} = \frac{1}{\alpha_t}(x_t-\frac{1-\alpha_t}{\sqrt{1-<\alpha_t>}}\phi(x_t,t) +\sigma z,   
+ x_{t-1} = \frac{1}{\alpha_t}(x_t-\frac{1-\alpha_t}{\sqrt{1-<\alpha_t>}}\phi(x_t,t)) +\sigma \mathbf{z},   
 \end{equation}
 
-with $z\sim\mathcal{N}(0,\mathbf{I})$.
+with $\mathbf{z}\sim\mathcal{N}(0,\mathbf{I})$.
 
-The process is iterated until $t=0$ and as such a new sample is generated.
+The process is iterated until $t=1$ and as such a new sample is generated.
 
@@
 \subsection*{Equivariant diffusion model}
-Unlike in image problems, the orientation of molecules matters under translations and rotations.
+Unlike for image problems, the orientation of molecules doew not matter under translations and rotations.
 For this purpose, equivariant diffusion models (EDMs) are used to account for rotational invariance.
 Therefore, we chose the EDM by Hoogeboom et al.~\cite{hoogeboom2022equivariant} as our baseline generative model, as it provides a generative model for new molecules and has equivariance already built into it.
-The basic concept behind equivarience is that the model is invariant to rotations and translation in this case the E(3) group.
+The basic concept behind equivarience is that the model is invariant to rotations and translation, in this case the E(3) group.
 This means that scalar (features such as atom types) and vector node properties (such as the positions) will be invariant to group transformations.
 As a result the order in which a rotation is applied does not matter.
-You can rotate the input to the model then apply the diffusion/denoising process and get a structure out, or you can apply the diffusion/denoising process first and then do the same rotation to get the same output.
+The input to the model can be rotated and diffusion/denoising  applied to get a structure, or the diffusion/denoising process can be applied first followed by the same rotation to get the same output.
 Mathematically this means that if we have a set of point $\mathbf{x} = (\mathbf{x}_1,\ldots,\mathbf{x}_N) \in \mathbb{R}^{N\times 3}$ and each of these points has an associated set of scalar feature vectors $\mathbf{h}\in \mathbb{R}^{N\times k}$ which are invariant to group transformations.
@@
 \end{equation}
-This iterative latenent variable refinement (ILVR) means that the condition can be applied during the denoising step without additional training.
-In the case of Choi et al, they used downsampled reference images as conditioning in order to generate images similar to the reference, but still a newly generated images.
+This iterative latent variable refinement (ILVR) means that the condition can be applied during the denoising steps without additional training.
+In the case of Choi et al, they used downsampled reference images as conditioning in order to generate images similar to the reference, but still newly generated images.
 
@@
 We propose a new method, that combines ILVR by Choi et al.~\cite{choi2021ilvr} proposed in an image generation context and the EDM by Hoogeboom et al..
-This allows us to generate a selective iterative latent variable refinement (SILVR) procedure in which we introduce a reference molecule into the denoising step.
-The reference used can be a small fragment, multiple fragments or even include dummy atoms that do not exist in the original reference.
+This allows us to generate a selective iterative latent variable refinement (SILVR) procedure in which we introduce a reference molecule into the denoising process.
+The reference used can be a small fragment, multiple fragments, or even include dummy atoms that do not exist in the original reference.
 As a result, we propose the following algorithm for SILVR~\ref{alg:silvr}.
-We are considering latent space variables $\mathbf{z}=[\mathbf{z}^{(x)},\mathbf{z}^{(h)}]$ for the standard denoising process and  $\mathbf{\hat{z}} = [\mathbf{\hat{z}}^{(y)},\mathbf{\hat{z}}^{(h)}]$, for the set of reference coordinates given by $\mathbf{y}$.
+We consider latent space variables $\mathbf{z}=[\mathbf{z}^{(x)},\mathbf{z}^{(h)}]$ for the standard denoising process and  $\mathbf{\hat{z}} = [\mathbf{\hat{z}}^{(y)},\mathbf{\hat{z}}^{(h)}]$, for the set of reference coordinates given by $\mathbf{y}$.
 The vector $\mathbf{h}$ contains all the scalar node properties of the equivariant EDM.
@@
 \State Subtract center of gravity (COG) from $\mathbf{\hat{z}}_0$
+\Comment{Center reference at zero}
 \State Sample $z_T \sim \mathcal{N}(\mathbf{0,I})$
@@
 \State Add COG($\mathbf{\hat{z}}_0$) to $\mathbf{z}_0$
-\Comment{Move sample back to pocket}
+\Comment{Move sample to original position of reference}
 \State Sample $\mathbf{x,h} \sim p(\mathbf{x,h|z_0})$ 
@@
 
-The core of the new method is the addition of a refinement process within the denoising part during run time.
+The core of the new method is the addition of a refinement step within the denoising process during runtime.
 The resulting SILVR model produces conditional samples without any conditional training when generating new molecules.
@@
 The EDM by Hoodgeboom \textit{et al.} was trained on the 30 lowest energy conformations of 430,000 molecules from the GEOM dataset with explicit hydrogens~\cite{axelrod2022geom}.
-A simplified schematic describing the EDM architecture is shared in scheme~\ref{fig:schematic} for more details on the model theory and training refer to Hoogeboom et al.~\cite{hoogeboom2022equivariant}.
+For more details on the model theory and training refer to Hoogeboom et al.~\cite{hoogeboom2022equivariant}.
 This model strictly only considers atomic coordinates; all bond information is ignored.
@@
 During training, the atomic coordinates (and their element) are passed through a forward diffusion process with iterative addition of Gaussian noise, where the extent of noise added at each step is defined by parameter $\beta$ (N.B.
-scheme~\ref{fig:schematic} shows the diffusion process as a Markov chain, however in practice the state at time $t$ is determined as a direct transformation of the initial state).
+Figure~\ref{fig:schematic} shows the diffusion process as a Markov chain, however in practice the state at time $t$ is determined as a direct transformation of the initial state).
 The diffusion process is eventually terminated when $t=T$, by which point all structure is lost and all coordinates follow a Gaussian distribution.
 An equivariant graph neural network (EGNN) is then trained to predict the reverse process, predicting the previous state in the sequence given any state.
-At run time, the generative model is instantiated with a sample from a Gaussian distribution and the series of denoising steps are applied resulting in a generated sample consisting of atomic coordinates resembling a drug-like molecule.
+At runtime, the generative model is instantiated with a sample from a Gaussian distribution and the series of denoising steps are applied resulting in a generated sample consisting of atomic coordinates resembling a drug-like molecule.
 The \texttt{XYZ} coordinates can then be interpreted using cheminformatics software to determine the molecular graph.
@@
 
-The EDM was adapted by introducing SILVR within the denoising process algorithm~\ref{alg:silvr}, as outlined previously.
-At run time, each atom of the reference set of coordinates is mapped to an atom in the EDM.
+The EDM was adapted by introducing SILVR within the denoising process, algorithm~\ref{alg:silvr}, as outlined previously.
+At runtime, each atom of the reference set of coordinates is mapped to an atom in the EDM.
 The reference coordinates are then translated such that their center of mass is at the origin.
@@
 This equation has the effect of shrinking the coordinates towards the origin and then expanding the coordinates out in the direction of the reference, see Figure~\ref{fig:silvr_explanation}.
-Importantly, the extent of this refinement is defined by the SILVR rate $r_{\mathrm{S}}$, with $r_{\mathrm{S}}=0$ providing no additional refinement and $r_{\mathrm{S}}=1$ will result in a total replacement of atoms.
+Importantly, the extent of this refinement is defined by the SILVR rate $r_{\mathrm{S}}$, with $r_{\mathrm{S}}=0$ providing no additional refinement and $r_{\mathrm{S}}=1$ resulting in a total replacement of atoms.
 The diffusion of the reference to $t=\tau$ requires the sampling of a Gaussian; while in principle this sampling could be performed only once prior to the denoising loop, in practice better results were obtained when the Gaussian was repeatedly sampled during the denoising process.
@@
 
-\subsubsection{SILVR to reference}
+\subsubsection{RMSD to reference}
 The SILVR algorithm creates a one-to-one mapping between reference atom coordinates and sample atom coordinates.
@@
 At very high values of $r_{\mathrm{S}}$ ($> 0.02$) there is a very high similarity between samples and the reference, however, most structures no longer resemble valid molecules.
-Best molecules are formed at intermediate values of $r_{\mathrm{S}}$ ($0.05 < r_{\mathrm{S}} < 0.01$) offering a trade-off between similarity to the reference, sampling diversity, and molecular likeness.
+The best molecules are formed at intermediate values of $r_{\mathrm{S}}$ ($0.05 \le r_{\mathrm{S}} \le 0.01$) offering a trade-off between similarity to the reference, sampling diversity, and molecular likeness.
 This is further validated by looking at stability measures.
@@
 The zeroth test we made with the generated samples was looking at how many generated molecules were fragmented i.e.
-are not a single connected molecular graph or not, with respect to $r_{\mathrm{S}}$.
+are not a single connected molecular graph, with respect to $r_{\mathrm{S}}$.
 This is presented in Figure S\ref{fig:fragmentation} in the SI.
@@
     \caption{Validation measures of SILVR model using fragments \texttt{x0072} and \texttt{x0354} as reference coordinates.
-A: Ratio of stable atoms - an atom is determined as stable if the valence matches the expected valence for the element B: RMSD from reference - the calculated RMSD between the reference and sample, using an absolute one-to-one mapping ignore atom identity C: OpenEye measure \texit{Shapegauss} - a Gaussian scoring function describing the shape fit between Mpro and samples, ignoring chemical interactions D: Geometry stability - AIMNet geometry optimisation was completed with Auto3D using the SMILES string of each sample.
+A: Ratio of stable atoms - an atom is determined as stable if the valence matches the expected valence for the element B: RMSD from reference - the calculated RMSD between the reference and sample, using an absolute one-to-one mapping ignore atom identity C: OpenEye measure \textit{Shapegauss} - a Gaussian scoring function describing the shape fit between Mpro and samples, ignoring chemical interactions D: Geometry stability - AIMNet geometry optimisation was completed with Auto3D using the SMILES string of each sample.
 RMSD was calculated between the predicted geometry and the sampled geometry using RDKit.
@@
 In general, the trends of Figure~\ref{fig:measures} were preserved for all experiments.
-In the follow we present three cases we investigated in detail, the test case using 3 fragments as a reference with substantial overlap, using two fragments with some overlap, and two disconnected fragments for the linker generation experiment.
+In the following, we present three cases we investigated in detail, the test case using 3 fragments as a reference with substantial overlap, using two fragments with some overlap, and two disconnected fragments for the linker generation experiment.
 
@@
 \label{sec:discussion}
-SILVR, as presented, represents method in which a general EDM can be conditioned to generate samples that resemble a reference structure without any additional training needed.
+SILVR, as presented, represents a method in which a general EDM can be conditioned to generate samples that resemble a reference structure, without any additional training needed.
 We showed that SILVR can complete both fragment merging and linking type tasks, without any \textit{a priori} knowledge of these design challenges.
 Considering all results with respect to the control EDM ($r_{\mathrm{S}}=0.0$), we show that at intermediate values of $r_{\mathrm{S}}$ the SILVR protocol produces molecules of equal quality to that of the unmodified EDM, while also guiding molecules towards reference structures.
-We, therefore, claim that if a diffusion model can be successfully trained to produce random high-quality drug-like structures, SILVR will provide molecular designs towards desired regions of chemical space without harming the quality of molecules.
+We, therefore, claim that if a diffusion model can be successfully trained to produce random high-quality drug-like structures, SILVR will provide molecular designs from desired regions of chemical space without harming the quality of molecules.
 Our method poses a direct interface between crystallographic fragment data and \textit{de novo} molecular generation.
-However, there are a few ways in which the current method can be improved further but we deem out of scope for this work.
+There are a few ways in which the current method can be improved further, but we deem these out of scope for this work.
 
@@
 a fragment from an X-ray fragment screen.
-The SILVR rate $r_{\mathrm{S}}$ allows the tuning of 'how much of the reference' molecule should be taken into account when generating new molecules, with medium values of $r_{\mathrm{S}}$ around 0.005 giving the best results.
+The SILVR rate $r_{\mathrm{S}}$ allows the tuning of 'how much of the reference' molecule should be taken into account when generating new molecules, with medium values of $r_{\mathrm{S}}$ around 0.005 to 0.01 giving the best results.
 The simple conditioning against a reference set of molecules means that the model can be used for tasks of fragment linker design, as well as generating new molecules that fit into an existing binding pocket without any specific training needed towards these tasks.
-It is also generalizable as it works on ligands with no information needed from the protein in its current form.
-In the future, given improvements in EDMs that produce more realistic and synthetically accessible molecules this method can cheaply generate structures exploring new chemical space with desired conditioning towards existing fragment hits.
+This method is also generalizable, as it works on ligands with no information needed from the protein in its current form.
+In the future, given improvements in EDMs that produce more realistic and synthetically accessible molecules, this method can cheaply generate structures exploring new chemical space with desired conditioning towards existing fragment hits.
 
@@
 The supporting information consists of a series of figures in addition to the figures in the main paper.
-Figure~\ref{fig:fragments} contains the 2D structures of all fragments used in the experiments as references.
+Figure~\ref{fig:moonshot_data} contains the 2D structures of all fragments used in the experiments as references.
 In Figure~\ref{fig:examples_curated} an example of two curated experiments is shown with different SILVR rates $r_{mathrm{s}}$.
-Next, Figure~\ref{fig:fragments}, shows on average how many molecules were not-fragmented in a set of 1000 samples for different SILVR rates.
+Next, Figure~\ref{fig:fragmentation}, shows on average how many molecules were not-fragmented in a set of 1000 samples for different SILVR rates.
 Figure~\ref{fig:SA_QED} includes synthetical accessibility(~\cite{ertl2009estimation}) and QED~\cite{bickerton2012quantifying, wildman1999prediction}.
@@
     \includegraphics[widht=0.5\textwidth]{paper/Figures/FigS3/FigS4.png}
-    \caption{Fraction of molecules fragmented with respect to the SILVR rate}
+    \caption{Fraction of molecules not fragmented with respect to the SILVR rate}
     \label{fig:fragmentation}
@@
     \includegraphics[width=0.8\textwidth]{paper/Figures/FigS4/si_fig_2d_measures.png}
-    \caption{Write caption for SA and QED}
+    \caption{A: The synthetic accessibility score (SA score) estimates the synthetic feasibility of molecules based on fragments contributions.~\cite{ertl2009estimation} This was calculated for all non-fragmented samples using an RDKit implementation of the scoring function SAScorer.
+A lower score indicates an easier-to-synthesise molecule: most catalogue and bioactive molecules fall in the range 2-5, while a score greater than 7 represents the upper end of complexity for natural products.~\cite{ertl2009estimation} B: The Quantitative Estimate of Druglikeness (QED) score combines a selection of descriptors such as molecular weight and calculated Log(P) to estimate the drug-likeness of a molecule.~\cite{bickerton2012quantifying, wildman1999prediction} This was calculated using RDKit with default settings across all non-fragmented samples.
+A higher score indicates a more drug-like molecule.}
     \label{fig:SA_QED}
@@
     \includegraphics[width=0.8\textwidth]{paper/Figures/FigS5/FigS3.png}
-    \caption{Example looking at the displacement of reference atoms after diffusion. A: for the whole molecule. B: Zoom in, on the displacement of atoms in the 9aH-Pyrido[1,2-a]pyrazine heterocycle. }
+    \caption{Example looking at the displacement of reference atoms after SILVR denosing.
+A: for the whole molecule.
+B: Zoom in, on the displacement of atoms in the 9aH-Pyrido[1,2-a]pyrazine heterocycle.
+}
     \label{fig:atom_displacement}
@@
 
-\newapge
+\begin{figure}
+    \centering\includegraphics[width=0.8\textwidth]{paper/Figures/FigS7/dummy_atoms_different_rates.png}
+    \caption{Effect of dummy atoms on the atom stability for linker design type experiment with reference fragments x0874 and x0397.
+A: $r_{\mathrm{S}} = 0.005$ and B:$r_{\mathrm{S}} = 0.01$.
+Lines represent sample median, circles sample mean.}
+    \label{fig:dummy_atoms}
+\end{figure}
+
+\newpage
 

### 2023-04-12 paper/manuscript.tex
small fixes in theory

@@
 \email{antonia.mey@ed.ac.uk}
-\phone{xxx}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title[SILVR: Molecular Generation for binding modes]
-  {SILVR: Conditional diffusion model for molecule generation without additional training}
+%\title[SILVR: Molecular Generation for binding modes]
+ % {SILVR: Conditional diffusion model for molecule generation without additional training}
 \title[SILVR: Molecular Generation for binding modes]
@@
 \begin{abstract}
-I'll be an abstract when I grow up. 
+Computationally generating novel synthetically accessible compounds with high affinity and low toxicity is a great challenge in drug-design.
+Machine-learning models beyond conventional pharmacophoric methods have shown promise in generating novel small molecule compounds, but require significant tuning for a specific protein target.
+Here, we introduce a method called selective iterative latent variable refinement (SILVR) for conditioning an existing, unconditionally trained, diffusion-based equivariant generative model to generate new molecules that fit into a binding site based on fragment hits.
+We use the SARS-CoV-2 Main Protease fragments from Diamond X-Chem that form part of the COVID Moonshot project as a reference dataset for conditioning the molecular generation.
+The SILVR rate controls the extent of conditioning and we show that moderate SILVR rates make it possible to generate new molecules of similar shape to the original fragments, meaning that the new molecules fit the binding site without knowledge of the protein.
+We can also merge up to 3 fragments into a new molecule without affecting the quality of molecules generated by the underlying generative model.
+Our method is generalizable to any protein target with known fragments and any diffusion based model for molecule generation.
 \end{abstract}
@@
 \label{sec:intro}
-One challenge  of computer-aided drug design is the ability to sample from a very large space of possible drug-like compounds to find suitable hits for a given target protein.
+Sampling from a very large space of possible drug-like compounds to find suitable hits for a given target protein is a open challenge in drug-design.
 It is estimated that there are between $10^{23}$ to $10^{60}$ feasible compounds, while only around $10^8$ have been synthesised so far~\cite{polishchuk2013estimation, reymond2012exploring}.
-Different strategies have been employed in the past to try and sample a diverse and synthetically accessible molecular space.
+Different strategies have been used   to try and sample a diverse and synthetically accessible molecular space from pharmacophore search~\cite{schwab2010conformations} to machine learning-based methods.
 In particular, methods based on machine learning (ML) have shown vast promise in this space in recent years~\cite{bilodeau2022generative}.
-Various neural network architectures have been proposed for molecular generation, from Variational autoencoders (VAE)~\cite{kingma2022autoencoding, jin2018junction, ma2018constrained, agoza2022generating}, to generative adversarial networks (GAN)~\cite{hoffmann2019generating} and normalising flows~\cite{shi2021learning}.
+Various neural network architectures have been proposed for molecular generation, from variational autoencoders (VAE)~\cite{kingma2022autoencoding, jin2018junction, ma2018constrained, ragoza2022generating}, to generative adversarial networks (GAN)~\cite{hoffmann2019generating} and normalising flows~\cite{shi2021learning}.
 More recently denoising diffusion probabilistic models, and particularly equivariant diffusion models, have shown promise in molecular generation~\cite{xu2022geodiff, hoogeboom2022equivariant}.
@@
 
-Typically the objective is to generate a diverse set of molecules for a given target protein that are easily synthetically accessible, and ideally with high binding affinity and low toxicity prediction~\cite{prieto-martinez2019chapter}.
-Different machine learning-based, and non-machine-learning-based, strategies have been used to address these ideas.
+Typically the objective is to generate a diverse set of molecules for a given target protein that are easily synthetically accessible, and ideally with high binding affinity and low predicted toxicity~\cite{prieto-martinez2019chapter}.
+Different machine-learning-based, and non-machine-learning-based, strategies have been used to address these ideas.
 From recurrent neural networks (RNN)~\cite{xu2021novo,zhang2022novo} and conditional variational autoencoders~\cite{ragoza2022generating} to GANs~\cite{skalic2019target}, various approaches have been used to generate molecules for a binding site where in each case the training is conditioned based on the target protein~\cite{xie2022advances}.
@@
 Even with a variety of ways to assess the synthetic accessibility of generated compounds~\cite{coley2018scscore, thakkar2021retrosynthetic}, molecules generated with these methods are often not easily synthesizable and in the worst cases can be chemically infeasible.
-A desirable feature for a generative model for drug design is the ability to generate compounds that are extensions of known fragment hits, or connect two existing fragments through "linkers", with the aim of making a new compound with improved pharmacochemical properties such as potency or solubility.
-Goa et al.
-have introduced a way to generate linkers for fragments using reinforcement learning strategies~\cite{guo2023linkinvent}, while Imrie et al.~\cite{imrie2020deep} have used variational auto encoders on this task without the use of protein information.
-More recently Huang et al.~\cite{huang20223dlinker} and Igashov et al.~\cite{igashov2022equivariant} have tackled this challenge using equivariant variational auto-encoder and diffusion based models respectively.
+
+Fragment-based drug discovery is an approach where a library of small molecular fragments ($<$300 Da) is screened against a target~\cite{hajduk2007decade, kumar2012fragment, bian2018computational}.
+These fragments are selected such that they present promiscuous binding, allowing exploration of many types of interactions a drug-like molecule could adopt within a given target.
+Individual fragments can not be drugs in and of themselves as they do not possess enough intermolecular interactions to achieve a sufficient binding affinity with a target, however, by considering an ensemble of known fragment hits, new high affinity binders can be constructed by merging and linking known fragments together and elaborating on singular fragments.
+An array of screening methods exist for determining if fragments bind a target, however here we focus on X-ray crystallography techniques.
+Protein drug targets can either be co-crystallized with fragments, or be crystalised unbound and subsequently soaked in a fragment solution.
+These crystals can be resolved by X-ray crystallography, with the results showing a high-quality electron density map revealing the exact binding geometry and interactions a fragment obtains with the given target.
+% A couple of sentences on the covid Moonshot dataset here to set the scene. 
+An application of generative models is therefore in the interpretation of such fragment data for the automated design of high affinity binders.
+Goa \textit{et al.} have introduced a way to generate linkers for fragments using reinforcement learning strategies~\cite{guo2023linkinvent}, while Imrie \textit{et al.}~\cite{imrie2020deep} have used variational auto encoders on this task without the use of protein information.
+More recently Huang \textit{et al.}~\cite{huang20223dlinker} and Igashov \textit{et al.}~\cite{igashov2022equivariant} have tackled this challenge using equivariant variational auto-encoder and diffusion based models respectively.
 Each of these models were explicitly trained for the specific purpose of linker generation.
 
-In this paper, we present SILVR, a selective iterative variable refinement (SILVR) method for conditioning an existing pre-trained equivariant diffusion model (EDM) towards the task of fragment merging and linker generation, yielding compounds similar to existing hits without specific training towards this task.
-To achieve this we combine the EDM by Hoogeboom et al.~\cite{hoogeboom2022equivariant} with the iterative latent variable refinement method proposed by Choi et al.
+In this paper, we present SILVR, a selective iterative latent variable refinement (SILVR) method for conditioning an existing pre-trained equivariant diffusion model (EDM) towards the task of fragment merging and linker generation, yielding compounds similar to existing hits without specific training towards this task.
+To achieve this we combine the EDM by Hoogeboom \textit{et al.}~\cite{hoogeboom2022equivariant} with the iterative latent variable refinement method proposed by Choi et al.
 for image generation networks~\cite{choi2021ilvr}.
@@
     \includegraphics[width=\textwidth]{paper/Figures/Fig1/Schematic.png}
-    \caption{Schematic of the equivariant diffusion model with  SILVR indicated for every denoising step.
+    \caption{Schematic of the equivariant diffusion model with  selective iterative latent variable refinement (SILVR) indicated for every denoising step.
 Here the reference in blue on the left is 3 small fragments.
 They are evolved over time $t$ in the diffusion process to resemble a Gaussian distribution at $t=T$.
-The $\beta$ added represents the noise added at each step, and the multiple arrows in grey show the speed up from time $t=2$ to $t=T$.
+The $\beta$ added represents the noise added at each step, and the multiple arrows in grey show the omitted steps from time $t=2$ to $t=T$.
 As atoms effectively 'diffuse' they can be perceived as changing position.
-To generate a new sample (left most yellow box), making use of the reference fragments, the EDM is sampled at each denoising step, as indicated by the EDM arrows, but now the reference at that same level of noise is added back in through SILVR at a given rate $r_{\mathbf{S}}$, until a new sample that resembles the reference is generated (following the bottom row along the yellow arrow).}
+To generate a new sample of a single molecule that links all fragments (left most yellow box), making use of the reference fragments, the EDM is sampled at each denoising step, as indicated by the EDM arrows, but now the reference at that same level of noise is added back in through SILVR at a given rate $r_{\mathbf{S}}$, until a new sample that resembles the reference is generated (following the bottom row along the yellow arrow).}
     \label{fig:schematic}
@@
 These ideas originated in image generation machine learning problems, but can also be applied to molecular generation.
-A neural network is trained to learn the second part of the process, denoising a Gaussian distribution until an image--or in this case--a molecule is generated.
-The idea we propose is to introduce information of a reference molecule at a given rate to the denoising process of a pre-trained model.
-This is similar to the concept of inpaining~\cite{qin2021image, xie2012image, squires2023artefact}, or more precisely re-painting~\cite{lugmayr2022repaint}, which guides the denoising process at each step towards the reference.
-In this paper we show that we can generate new compounds that are similar to a given reference and even link fragments using multiple references as input.
-This makes it generalizable to any protein-ligand systems with known fragment hits.
-For illustration purposes we use the original 23 X-ray fragment hits for the  SARS-CoV2 Main protease from the Covid-Moonshot dataset~\cite{consortium2021open,consortium2023open}.
-In the following, we give an overview of how equivariant difusion models work and introduce how our method SILVR fits into the framework of an existing pre-trained EDM.
+A neural network is trained to learn the second part of the process, denoising a Gaussian distribution until an image---or in this case, a molecule---is generated.
+The idea we propose is to introduce information from a reference molecule at a given rate to the denoising process of a pre-trained model.
+This is similar to the concept of inpainting~\cite{qin2021image, xie2012image, squires2023artefact}, or more precisely re-painting~\cite{lugmayr2022repaint}, which guides the denoising process at each step towards the reference.
+In this paper we show that we can generate new compounds that are similar to a given reference, and link fragments together, using multiple superimposed fragments as input.
+This method is generalizable to any protein-ligand systems with known fragment hits.
+For illustration purposes we use the original 23 X-ray fragment hits for the  SARS-CoV-2 Main protease from the COVID Moonshot dataset~\cite{consortium2023open}.
+In the following, we give an overview of how equivariant diffusion models work and introduce how our method SILVR fits into the framework of an existing pre-trained EDM.
 We then show, how the SILVR rate, $r_{\mathrm{S}}$ as a modelling parameter can modulate how much of the reference is incorporated in the sample generation and how the original EDM is recovered at  $r_{\mathrm{S}}=0$.
@@
 \label{sec:theory}
-\subsection*{Denoising Diffusion probabilistic models (DDPM) as generative models}
+\subsection*{Denoising diffusion probabilistic models (DDPM) as generative models}
 DDPMs are often used as generative models that were developed for the generation of new images~\cite{sohl-dickstein2015deep, ho2020denoising, nichol2021improved}.
@@
 The main idea behind diffusion models is for a neural network to learn the reverse of a diffusion process, often referred to as denoising, to \textit{sample} a new image or in our case a new molecule.
-In practice, this is done by training a neural network $\phi$ and generating samples $p(\mathbf{x}_{t-1}|\mathbf{x}_t)=\mathcal{N}(x_{t-1}; \mu_{\theta}(x_t;t), \sigma_t^2\mathbf{I})$ from  a Gaussian transition with learned mean ($\mu_{\theta}(x_t;t)$) and variance ($\sigma_t$) with $\mathbf{x}_t$ being the data, noised up to time $t$.
+In practice, this is done by training a neural network $\phi$ and generating samples $p(\mathbf{x}_{t-1}|\mathbf{x}_t)=\mathcal{N}(x_{t-1}; \mu_{\theta}(x_t;t), \sigma_t^2\mathbf{I})$ from  a Gaussian transition kernel with learned mean ($\mu_{\theta}(x_t;t)$) and variance ($\sigma_t$), with $\mathbf{x}_t$ being the data noised up to time $t$.
 Figure~\ref{fig:schematic} shows a schematic of the two main parts of such a DDPM, the diffusion part where noise is added for each timestep (shown in blue) and the denoising part (shown in yellow) that allows the generation of a new image (molecule).
 
-The \textit{forward} part of the diffusion model is a Markov process: noise is added to a set of data $\mathbf{x}$ and over a time interval $t = [0,\ldots,T]$ according to the following distribution:
+The \textit{forward} part of the diffusion model is a Markov process: noise is added to a set of data $\mathbf{x}$ and over a time interval $t = [1,\ldots,T]$ according to the following distribution:
 \begin{equation}
@@
 $\beta_t$ is a fixed variance schedule which adds noise with each timestep $t$.
-Different people have looked at different noise schedules~\cite{sohl-dickstein2015deep, ho2020denoising}.
+Different researchers have examined different noise schedules~\cite{sohl-dickstein2015deep, ho2020denoising}.
 % Next talk about the denoising step
@@
 \end{equation}
-A sample for denoising timestep $t-1$ can be generated from the following equation given the neural network $\phi$ that has been trained on the diffusion process.
-
-A given samples is generated as:
+A sample for denoising timestep $t-1$ can be generated from the following equation given the neural network $\phi$ that has been trained on the diffusion process: 
 \begin{equation}
- x_{t-1} = \frac{1}{\alpha_t}(x_t-\frac{1-\alpha_t}{\sqrt{1-<\alpha_t>}}\phi(x_t,t)) +\sigma \mathbf{z},   
+ x_{t-1} = \frac{1}{\alpha_t}(x_t-\frac{1-\alpha_t}{\sqrt{1-\langle\alpha_t\rangle}}\phi(x_t,t)) +\sigma \mathbf{z},   
 \end{equation}
@@
 
-The process is iterated until $t=1$ and as such a new sample is generated.
+The process is iterated until $t=1$ and as such A new sample $x_0$ of the denoising process---which is intended to represent a proposed molecule design---is generated.
 
@@
 %%%%%%%
-\subsection*{Equivariant diffusion model}
-Unlike for image problems, the orientation of molecules doew not matter under translations and rotations.
+\subsection*{Equivariant diffusion model (EDM)}
+Unlike for image problems, the orientation of molecules does not matter under translations and rotations.
 For this purpose, equivariant diffusion models (EDMs) are used to account for rotational invariance.
-Therefore, we chose the EDM by Hoogeboom et al.~\cite{hoogeboom2022equivariant} as our baseline generative model, as it provides a generative model for new molecules and has equivariance already built into it.
+Therefore, we chose the EDM by Hoogeboom \textit{et al.}~\cite{hoogeboom2022equivariant} as our baseline generative model, as it provides a generative model for new molecules and has equivariance already built into it.
 The basic concept behind equivarience is that the model is invariant to rotations and translation, in this case the E(3) group.
@@
 The positions translations and rotations are defined according to the orthogonal matrix: $\mathbf{Rx + t} = (\mathbf{R_{x_1}+t},\ldots, \mathbf{R_{x_N}+t})$.
-Satorras et al.~\cite{satorras2022equivariant} have proposed an E(n) Equivariant Graph Neural Network on which the EDM by Hoogeboom et al.
-builds.
+Satorras \textit{et al.}~\cite{satorras2022equivariant} have proposed an E(n) Equivariant Graph Neural Network on which the EDM by Hoogeboom \textit{et al.} builds.
 For full details on the architecture and their code see~\cite{hoogeboom2022equivariant}.
@@
 \subsection{Iterative Latent Variable Refinement as a conditioning tool}
-Choi et al.~\cite{choi2021ilvr} introduced a way to condition DDPMs by sampling their images from the conditional distribution $p(x_0|c)$, with the condition c:
+Choi \textit{et al.}~\cite{choi2021ilvr} introduced a way to condition DDPMs by sampling their images from the conditional distribution $p(x_0|c)$, with the condition c:
 \begin{equation}
@@
 This iterative latent variable refinement (ILVR) means that the condition can be applied during the denoising steps without additional training.
-In the case of Choi et al, they used downsampled reference images as conditioning in order to generate images similar to the reference, but still newly generated images.
+In the case of Choi et al, they used downsampled reference images as conditioning in order to generate novel images similar to the reference.
 
@@
 %%%%
-\subsection{Selective Iterative Latent Variable Refinement}
+\subsection{Selective Iterative Latent Variable Refinement (SILVR)}
 \label{sec:silvr}
-We propose a new method, that combines ILVR by Choi et al.~\cite{choi2021ilvr} proposed in an image generation context and the EDM by Hoogeboom et al..
-This allows us to generate a selective iterative latent variable refinement (SILVR) procedure in which we introduce a reference molecule into the denoising process.
-The reference used can be a small fragment, multiple fragments, or even include dummy atoms that do not exist in the original reference.
+We propose a new method, that combines ILVR by Choi \textit{et al.}~\cite{choi2021ilvr} proposed in an image generation context and the EDM by Hoogeboom \textit{et al.}.
+This allows us to generate a selective iterative latent variable refinement (SILVR) procedure in which we introduce information of a reference molecule into the denoising process.
+We describe this method as "selective" due to the ability to guide individual atoms at independent rates.
+The reference can be a single molecule or a series of superimposed fragments, and additional unguided dummy atoms can also be defined at the beginning of the denoising process.
 As a result, we propose the following algorithm for SILVR~\ref{alg:silvr}.
@@
 To illustrate the usefulness of this run-time modification, we show how SILVR can be used in the context of fragment-based drug design.
-The goal is to produce molecules that resemble a binding site based on exiting fragments and their set of atomic coordinates.
+The goal is to produce molecules that are complementary to a binding site based on exiting fragments and their set of atomic coordinates.
 %As stated before, the selective iterative latent variable refinement (SILVR) method makes use of the pre-trained equivariant diffusion model (EDM) as introduced by Hoogeboom \textit{et al.}~\cite{hoogeboom2022equivariant}.
@@
 The EDM by Hoodgeboom \textit{et al.} was trained on the 30 lowest energy conformations of 430,000 molecules from the GEOM dataset with explicit hydrogens~\cite{axelrod2022geom}.
-For more details on the model theory and training refer to Hoogeboom et al.~\cite{hoogeboom2022equivariant}.
-This model strictly only considers atomic coordinates; all bond information is ignored.
-A more recent version of the EDM also accounts for bond order and SILVR could also be incorporated into this new model~\cite{vignac2023midi}.
-During training, the atomic coordinates (and their element) are passed through a forward diffusion process with iterative addition of Gaussian noise, where the extent of noise added at each step is defined by parameter $\beta$ (N.B.
-Figure~\ref{fig:schematic} shows the diffusion process as a Markov chain, however in practice the state at time $t$ is determined as a direct transformation of the initial state).
+For more details on the model theory and training refer to Hoogeboom \textit{et al.}~\cite{hoogeboom2022equivariant}.
+This model strictly only considers atomic coordinates (heavy atoms and hydrogen atoms), while all bond and molecular graph information is ignored.
+A more recently proposed version of the EDM incorporates molecular graph generation within the denoising diffusion model, and we believe it should be possible to incorporate SILVR into the denoising loop of this new model following the same principles we describe in this paper~\cite{vignac2023midi}.
+During training the atomic coordinates, and a one-hot encoding of their element, are passed through a forward diffusion process with iterative addition of Gaussian noise; both the coordinates and one-hot vector get diffused during this process.
+The extent of noise added at each step is defined by parameter $\beta$ (N.B.\ Figure~\ref{fig:schematic} shows the diffusion process as a Markov chain, however in practice the state at time $t$ can be efficiently computed as a direct transformation of the initial state).
 The diffusion process is eventually terminated when $t=T$, by which point all structure is lost and all coordinates follow a Gaussian distribution.
@@
 At runtime, the generative model is instantiated with a sample from a Gaussian distribution and the series of denoising steps are applied resulting in a generated sample consisting of atomic coordinates resembling a drug-like molecule.
-The \texttt{XYZ} coordinates can then be interpreted using cheminformatics software to determine the molecular graph.
+The resulting Cartesian coordinates can then be interpreted using cheminformatics software to determine the molecular graph.
 
@@
 The EDM was adapted by introducing SILVR within the denoising process, algorithm~\ref{alg:silvr}, as outlined previously.
-At runtime, each atom of the reference set of coordinates is mapped to an atom in the EDM.
+At runtime, each atom of the reference set of coordinates is mapped to an atom in the EDM; this is simply achieved by constructing a reference tensor with the same shape as the EDM latent tensor, with the mapping being on a row-by-row basis.
 The reference coordinates are then translated such that their center of mass is at the origin.
@@
 
-\subsection{Reference Dataset: Covid-moonshot}
-Reference molecules were selected from the original 23 non-covalent hits of the SARS-CoV-2 Main Protease (Mpro) identified by the XChem fragment screen as part of the Covid Moonshot Project~\cite{consortium2023open, consortium2021open}.
-A more detailed picture of all fragments is presented in FigureS~\ref{fig:moonshot_data} of the Supplementary information(SI).
-Fragments were visualised using NGLview version 3.03 and combinations were arbitrarily selected as test cases for the following different experimental setups beyond trying to understand similarity between refernce and new sample.
+\subsection{Reference Dataset: COVID Moonshot}
+Reference molecules were selected from the original 23 non-covalent hits of the SARS-CoV-2 Main Protease (Mpro) identified by the XChem fragment screen~\cite{douangamath2020crystallographic} as part of the Covid Moonshot Project~\cite{consortium2023open, consortium2021open}.
+A more detailed picture of all fragments is presented in Figure~\ref{fig:moonshot_data} of the Supplementary information(SI).
+Fragments were visualised using NGLview version 3.03 and combinations were arbitrarily selected as test cases for the following different experimental settings for trying to understand the similarity between the reference and the new sample.
 We looked at the following scenarios:
@@
 \end{enumerate}
-Fragments \texttt{x0072} and \texttt{x0354} were selected for benchmarking the effect of $r_{\mathrm{S}}$ on sampling; \texttt{x1093}, \texttt{x0072}, \texttt{x2193} were randomly chosen to represent three significantly overlapping fragments; \texttt{x0434}, \texttt{x0305} and \texttt{x0072}, \texttt{x0354} were used as partially overlapping fragments; and \texttt{x0874}, \texttt{x0397} were used as two disconnected fragments, resembling a linker design type problem.
-The bonding information of selected fragments was deleted and \texttt{XYZ} coordinates were combined into a single \texttt{XYZ} file.
+Fragments \texttt{x0072} and \texttt{x0354} were selected for benchmarking the effect of the SILVR rate $r_{\mathrm{S}}$ on sampling; \texttt{x1093}, \texttt{x0072}, \texttt{x2193} were randomly chosen to represent three significantly overlapping fragments; \texttt{x0434}, \texttt{x0305} and \texttt{x0072}, \texttt{x0354} were used as partially overlapping fragments; and \texttt{x0874}, \texttt{x0397} were used as two disconnected fragments, resembling a linker design type problem.
+The bonding information of selected fragments was deleted and cartesian coordinates were combined into a single \texttt{XYZ} file.
 Values of $r_{\mathrm{S}}$ were selected and added to the \texttt{XYZ} file to create a reference file containing all experiment setup information.
@@
 \subsubsection{Atom stability}
-The accuracy of placement of atoms was determined using the stability metric proposed by Hoogeboom et al..
+The accuracy of placement of atoms was determined using the stability metric proposed by Hoogeboom \textit{et al.}~\cite{hoogeboom2022equivariant}.
 This metric infers bonds and bond orders between atoms by considering their interatomic distances.
 Once all bonds are defined, the valence of each atom is compared to its expected valence and if these values match, the atom is determined as stable.
-It should be noted that this metric requires the presence of all hydrogens (both explicit and implicit) for an atom to be classed as stable.
+It should be noted that this measure requires the explicit presence of all hydrogens for an atom to be classed as stable.
 For comparability with other similar published models, this measure was used unmodified.
@@
 The SILVR algorithm creates a one-to-one mapping between reference atom coordinates and sample atom coordinates.
-The RMSD for this pairwise mapping, ignoring atom identities, was calculated to determine the spacial similarity of samples to the reference.
+The RMSD for this pairwise mapping, ignoring atom identities, was calculated to determine the spatial similarity of samples to the reference.
 All RMSD calculations were carried out using \url{https://github.com/charnley/rmsd} version 1.5.1.
@@
 
-\subsubsection{Shape similarity of sample to protein}
-The agreement in the shape of samples and the main protease binding site was determined using openeye-toolkits version 2022.2.2 Gaussian scoring function \textit{Shapegauss}~\cite{openeyescientificsoftwareinc.oedocking,kelley2015posit}.
-This scoring function measures the shape similarity between the ligand and receptor by considering each heavy atom as a Gaussian function~\cite{mcgann2003gaussian}.
+\subsubsection{Shape complementarity of sample to protein}
+The agreement in the shape of samples and the main protease binding site was determined using openeye-toolkits version 2022.2.2 Gaussian scoring function \textit{shapegauss}~\cite{openeyescientificsoftwareinc.oedocking,kelley2015posit}.
+This scoring function measures the shape complementarity between the ligand and receptor by considering each heavy atom as a Gaussian function~\cite{mcgann2003gaussian}.
 The most favourable score occurs “when two atoms touch but do not overlap”.
-This metric does not consider any intermolecular interactions.
+This metric does not consider any intermolecular interactions beyond shape complementarity.
 The protein receptor file was prepared from the Mpro-x0072 crystal structure with removal of the ligand.
-The \texttt{XYZ} coordinates of samples were directly read into openeye-toolkit and the pose was re-scored with \textit{Shapegauss}.
+The \texttt{XYZ} coordinates of samples were directly read into openeye-toolkit and the pose was re-scored with \textit{shapegauss}.
 
@@
 Qualitatively, the generated molecular samples from the condition EDM  show a clear resemblance to their reference structures, with similarity increasing with $r_{\mathrm{S}}$.
-Figure S~\ref{fig:examples_curated}, in the SI, shows two example samples started from fragments \texttt{x0072} and \texttt{x0354} over a range of SILVR rates between $r_{\mathrm{S}}=0$ to $r_{\mathrm{S}}=0.02$.
+Figure S~\ref{fig:examplese_curated}, in the SI, shows two example samples started from fragments \texttt{x0072} and \texttt{x0354} over a range of SILVR rates between $r_{\mathrm{S}}=0$ to $r_{\mathrm{S}}=0.02$.
 As expected at no conditioning random samples are generated that do not resemble the reference fragments.
@@
 \subsection{Intermediate SILVR rates produce stable and varied molecules}
-To assess the stability and variability of generated molecules we looked at three different metrics, as discussed in the methods section.
+To assess the stability and variability of generated molecules we looked at four different metrics, as discussed in the methods section.
 We generated 1000 samples at different $r_{\mathrm{S}}$ using fragments \texttt{X0072} and \texttt{X0354} as a reference.
@@
 This is presented in Figure S\ref{fig:fragmentation} in the SI.
-At an intermediate $r_{\mathrm{S}} = 0.025$ just over 50\% of the generated samples are not fragmented meaning that in the worst-case scenario, only one in two generated molecules can be analysed further.
+At an intermediate $r_{\mathrm{S}} = 0.025$ just over 50\% of the generated samples are not fragmented meaning that, only one in two generated molecules can be analysed further.
 The subsequent analysis is carried out only on whole molecular graphs.
-Figure~\ref{fig:measures} A looks at the \textit{atom stability} measure as introduced by Hoogeboom et al.~\cite{hoogeboom2022equivariant}.
-Samples generated at low $r_{\mathrm{S}}$ tend to have similar stabilities to each other, samples start becoming less stable around $r_{\mathrm{S}}=0.005$, and become totally unstable at $r_{\mathrm{S}}=0.02$.
+Figure~\ref{fig:measures} A looks at the \textit{atom stability} measure as introduced by Hoogeboom \textit{et al.}~\cite{hoogeboom2022equivariant}.
+Samples generated at low $r_{\mathrm{S}}$ tend to have similar atom stability, samples start becoming less stable around $r_{\mathrm{S}}=0.005$, and become totally unstable at $r_{\mathrm{S}}=0.02$.
 This trend can largely be explained due to issues around hydrogens.
@@
 Since there are no hydrogens in the reference, all atoms are mapped to heavy atoms, and therefore most atoms are unable to satisfy a full valence.
-Adding hydrogens explicitly to the molecules, through OpenBable or RDkit, is a way of improving this measure.
+Adding hydrogens explicitly to the molecules, through OpenBabel or RDkit, is a way of improving this measure.
 
-\begin{figure}[h!]
+\begin{figure}[htb!]
     \centering
@@
     \caption{Validation measures of SILVR model using fragments \texttt{x0072} and \texttt{x0354} as reference coordinates.
-A: Ratio of stable atoms - an atom is determined as stable if the valence matches the expected valence for the element B: RMSD from reference - the calculated RMSD between the reference and sample, using an absolute one-to-one mapping ignore atom identity C: OpenEye measure \textit{Shapegauss} - a Gaussian scoring function describing the shape fit between Mpro and samples, ignoring chemical interactions D: Geometry stability - AIMNet geometry optimisation was completed with Auto3D using the SMILES string of each sample.
+A: Ratio of stable atoms - an atom is determined as stable if the valence matches the expected valence for the element B: RMSD from reference - the calculated RMSD between the reference and sample, using an absolute one-to-one mapping ignore atom identity with low RMSD meaning molecules are similar to the reference and high RMSD they are not.
+C: OpenEye measure \textit{shapegauss} - a Gaussian scoring function describing the shape fit between Mpro and samples, ignoring chemical interactions.
+A lower score means...
+D: Geometry stability - AIMNet geometry optimisation was completed with Auto3D using the SMILES string of each sample.
 RMSD was calculated between the predicted geometry and the sampled geometry using RDKit.
@@
 As one of the main motivators for SILVR is to be able to generate new molecules that fit directly into a binding site based on input fragments, we measured shape complementarity between newly generated samples and the Mpro binding site.
-We used the OpenEye \textit{Shapegauss} scoring function for this purpose~\cite{kelley2015posit}.
+We used the OpenEye \textit{shapegauss} scoring function for this purpose~\cite{kelley2015posit}.
 From Figure~\ref{fig:measures} C it can be seen that the shape complementarity of samples improves with increasing $r_{\mathrm{S}}$, demonstrating that SILVR can produce ligands of binding site geometry when guided by the coordinates of fragment molecules.
-The lack of plot data for $r_{\mathrm{S}}=0.02$ and $r_{\mathrm{S}}=0.03$ was because all shapegauss calculations failed.
+The lack of plot data for $r_{\mathrm{S}}=0.02$ and $r_{\mathrm{S}}=0.03$ was because all \textit{shapegauss} calculations failed.
 We believe this was due to the atom coordinates representing highly strained and internally clashing molecules, and the scoring algorithm either failed to read the molecules or identified them as bad conformations.
 
-\begin{figure}
+\begin{figure}[!htb]
     \centering
@@
 
-\subsection{There is no control over the retention of functional groups from the reference structure}
+\subsection{The retention of functional groups from the reference structure is challenging}
 When applied in a drug-design context, the conservation of key functional groups in exact spacial positions is crucial to maintain desirable protein-ligand interactions.
@@
 To answer structure-based drug design, a scoring function of protein surface potentials could be used to guide atoms towards energetically favoured pocket regions; issues of synthetic accessibility might be solved by clever design of refinement methods; and interfacing with alternative data sources (such as NMR) might be achievable.
+
+\subsection{Placements of hydrogens and dummy atoms needs additional trials}
+An EDM with explicity hydrogens will improve the overall models.
+At the moment there is a mixing of explicit and implicity hydrogens depending on need for analysis and input.
+An optimal model can account for hydrogens both explicityly and implicityly allowing for scoring of either.
+In addition using dummy atoms strategically for growing certain parts of a fragment is something to be explored further in the future.
 
@@
 All data for the experiments carried out and instructions on how to reproduce this work can be found at~\url{https://github.com/meyresearch/SILVR}.
-An updated version of Hoogeboom et al.
-EDM that includes SILVR can be found at~\url{https://github.com/nichrun/e3_diffusion_for_molecules}.
+An updated version of Hoogeboom \textit{et al.} EDM that includes SILVR can be found at~\url{https://github.com/nichrun/e3_diffusion_for_molecules}.
 
@@
     \centering
-    \includegraphics[widht=0.5\textwidth]{paper/Figures/FigS3/FigS4.png}
+    \includegraphics[width=0.5\textwidth]{paper/Figures/FigS3/fraction_whole_samples.png}
     \caption{Fraction of molecules not fragmented with respect to the SILVR rate}

### 2023-04-13 paper/manuscript.tex
Updates from Overleaf

@@
 Machine-learning models beyond conventional pharmacophoric methods have shown promise in generating novel small molecule compounds, but require significant tuning for a specific protein target.
-Here, we introduce a method called selective iterative latent variable refinement (SILVR) for conditioning an existing, unconditionally trained, diffusion-based equivariant generative model to generate new molecules that fit into a binding site based on fragment hits.
+Here, we introduce a method called selective iterative latent variable refinement (SILVR) for conditioning an existing a diffusion-based equivariant generative model without retraining.
+The model allows the generation of new molecules that fit into a binding site of a protein based on fragment hits.
 We use the SARS-CoV-2 Main Protease fragments from Diamond X-Chem that form part of the COVID Moonshot project as a reference dataset for conditioning the molecular generation.
@@
 Typically the objective is to generate a diverse set of molecules for a given target protein that are easily synthetically accessible, and ideally with high binding affinity and low predicted toxicity~\cite{prieto-martinez2019chapter}.
-Different machine-learning-based, and non-machine-learning-based, strategies have been used to address these ideas.
-From recurrent neural networks (RNN)~\cite{xu2021novo,zhang2022novo} and conditional variational autoencoders~\cite{ragoza2022generating} to GANs~\cite{skalic2019target}, various approaches have been used to generate molecules for a binding site where in each case the training is conditioned based on the target protein~\cite{xie2022advances}.
-Some of these models incorporate a ligand score directly~\cite{ragoza2022generating}, while others require methods based on machine learning (ML)~\cite{masters2023deep}, or more traditional docking~\cite{fu2018predictive} and molecular simulation-based affinity prediction methods~\cite{mey2016blinded,mey2020best}.
+A plethora of methods has been devised for the generation of molecules, but also assessment of binding affinity such as more traditional docking~\cite{fu2018predictive} and molecular simulation-based affinity prediction methods~\cite{mey2016blinded,mey2020best} and machine learning methods more recently.
+Machine learning has lead to an explosion of methods for molecular generation, from recurrent neural networks (RNN)~\cite{xu2021novo,zhang2022novo} and conditional variational autoencoders~\cite{ragoza2022generating} to GANs~\cite{skalic2019target}, various approaches have been used to generate molecules for a binding site where in each case the training is conditioned based on the target protein~\cite{xie2022advances}.
+Some of these models incorporate a ligand score directly~\cite{ragoza2022generating}, while others require methods based on machine learning (ML)~\cite{masters2023deep}, or more conventional methods.
 Even with a variety of ways to assess the synthetic accessibility of generated compounds~\cite{coley2018scscore, thakkar2021retrosynthetic}, molecules generated with these methods are often not easily synthesizable and in the worst cases can be chemically infeasible.
@@
 In this paper, we present SILVR, a selective iterative latent variable refinement (SILVR) method for conditioning an existing pre-trained equivariant diffusion model (EDM) towards the task of fragment merging and linker generation, yielding compounds similar to existing hits without specific training towards this task.
-To achieve this we combine the EDM by Hoogeboom \textit{et al.}~\cite{hoogeboom2022equivariant} with the iterative latent variable refinement method proposed by Choi et al.
-for image generation networks~\cite{choi2021ilvr}.
+To achieve this we combine the EDM by Hoogeboom \textit{et al.}~\cite{hoogeboom2022equivariant} with the iterative latent variable refinement method proposed by Choi et al.~\cite{choi2021ilvr} for image generation networks.
 This allows the generation of new molecules in the shape of a binding site using information from existing fragment hits, without specific training or knowledge of this task.
 
-\begin{figure}
+\begin{figure}[h!]
     \centering
@@
 This is similar to the concept of inpainting~\cite{qin2021image, xie2012image, squires2023artefact}, or more precisely re-painting~\cite{lugmayr2022repaint}, which guides the denoising process at each step towards the reference.
-In this paper we show that we can generate new compounds that are similar to a given reference, and link fragments together, using multiple superimposed fragments as input.
+In this paper, we show that we can generate new compounds that are similar to a given reference, and link fragments together, using multiple superimposed fragments as input.
 This method is generalizable to any protein-ligand systems with known fragment hits.
@@
     \centering
-    \includegraphics[width=\textwidth]{paper/Figures/Fig3/fig_3_metrics.png}
+    \includegraphics[width=\textwidth]{paper/Figures/Fig3/fig_3_metrics_violins.png}
     \caption{Validation measures of SILVR model using fragments \texttt{x0072} and \texttt{x0354} as reference coordinates.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\bibliographystyle{unsrt}
 \bibliography{SILVR}

### 2023-04-17 paper/manuscript.tex
Updates from Overleaf

@@
 \usepackage{amssymb}
+\usepackage[
+ %eprint=false,
+ %style=plain
+]{biblatex}
+%\ExecuteBibliographyOptions[misc]{eprint=true}
+\addbibresource{SILVR.bib}
 
@@
     \centering
-    \includegraphics[width=\textwidth]{paper/Figures/Fig1/Schematic.png}
+    \includegraphics[width=\textwidth]{paper/Figures/Fig1/Schematic_new.png}
     \caption{Schematic of the equivariant diffusion model with  selective iterative latent variable refinement (SILVR) indicated for every denoising step.
@@
 We then show, how the SILVR rate, $r_{\mathrm{S}}$ as a modelling parameter can modulate how much of the reference is incorporated in the sample generation and how the original EDM is recovered at  $r_{\mathrm{S}}=0$.
-As part of this, we illustrate how fragments can be linked using dummy atoms and how newly generated molecules can fit into the existing binding site according to shape similarity.
+As part of this, we illustrate how fragments can be linked using dummy atoms and how newly generated molecules can fit into the existing binding site according to shape complementarity.
 
@@
 
-The \textit{forward} part of the diffusion model is a Markov process: noise is added to a set of data $\mathbf{x}$ and over a time interval $t = [1,\ldots,T]$ according to the following distribution:
+The \textit{forward} part of the diffusion model is a Markov process: noise is added to a set of data $\mathbf{x}_0$ and over a time interval $t = [1,\ldots,T]$ according to the following distribution:
 \begin{equation}
-  q(\mathbf{x}_{1:T}|\mathbf{x_0}) = q(x_0)\prod_{t=1}^T q(\mathbf{x}_t|\mathbf{x}_{t-1}).
+  q(\mathbf{x}_{1:T}|\mathbf{x}_0) = q(x_0)\prod_{t=1}^T q(\mathbf{x}_t|\mathbf{x}_{t-1}).
 \end{equation}
@@
 for any $t>s$.
-The parameters $\alpha_t \in \mathbb{R}^{+}$ contains the amount of retained signal and $\sigma_t \in \mathbb{R}^{+}$ represents the variance and thus the amount of white noise added.
+The parameters $\alpha_t \in \mathbb{R}^{+}$ specifies the amount of retained signal and $\sigma_t \in \mathbb{R}^{+}$ represents the variance and thus the amount of white noise added.
 The parameter $\alpha_t$ also directly relates to $\beta$ in Figure~\ref{fig:schematic} with $\alpha_t:=1-\beta_t$ and $\langle\alpha\rangle_t:=\prod_{s=1}^t\alpha_s$.
@@
 
-The process is iterated until $t=1$ and as such A new sample $x_0$ of the denoising process---which is intended to represent a proposed molecule design---is generated.
+The process is iterated until $t=1$ and as such a new sample $x_0$ of the denoising process---which is intended to represent a proposed molecule design---is generated.
 
@@
 \subsection*{Equivariant diffusion model (EDM)}
-Unlike for image problems, the orientation of molecules does not matter under translations and rotations.
-For this purpose, equivariant diffusion models (EDMs) are used to account for rotational invariance.
-Therefore, we chose the EDM by Hoogeboom \textit{et al.}~\cite{hoogeboom2022equivariant} as our baseline generative model, as it provides a generative model for new molecules and has equivariance already built into it.
+In the previous section introduced the neural network $\phi(\mthabf{x},t)$.
+In practice, it makes sense to use an euqivariant graph neural network, as it is a data efficient way to learn about molecules.
+If a model has rotational and translational equivarience, it means a neural network does not need to learn orientations and transations of molecules from scratch.
+We chose the EDM by Hoogeboom \textit{et al.}~\cite{hoogeboom2022equivariant} as our baseline generative model, as it provides a generative model for new molecules and has equivariance already built into it.
+Furthermore, it has all code and pretrained weights available online~\cite{githubhoogeboom}.
 The basic concept behind equivarience is that the model is invariant to rotations and translation, in this case the E(3) group.
@@
 Mathematically this means that if we have a set of point $\mathbf{x} = (\mathbf{x}_1,\ldots,\mathbf{x}_N) \in \mathbb{R}^{N\times 3}$ and each of these points has an associated set of scalar feature vectors $\mathbf{h}\in \mathbb{R}^{N\times k}$ which are invariant to group transformations.
-The positions translations and rotations are defined according to the orthogonal matrix: $\mathbf{Rx + t} = (\mathbf{R_{x_1}+t},\ldots, \mathbf{R_{x_N}+t})$.
+The position translations and rotations are defined according to the orthogonal matrix: $\mathbf{Rx + t} = (\mathbf{R_{x_1}+t},\ldots, \mathbf{R_{x_N}+t})$.
 Satorras \textit{et al.}~\cite{satorras2022equivariant} have proposed an E(n) Equivariant Graph Neural Network on which the EDM by Hoogeboom \textit{et al.} builds.
 For full details on the architecture and their code see~\cite{hoogeboom2022equivariant}.
+%Something on latent variables Z.
+
 %%%%
@@
 \begin{equation}
-    p_{\theta}(x_{0:T}|c)=p(x_T)\prod_{t=1}^{T}p_{\theta}(x_{t-1}\x_t,c)
+    p_{\theta}(x_{0:T}|c)=p(x_T)\prod_{t=1}^{T}p_{\theta}(x_{t-1}x_t,c)
 \end{equation}
@@
 \State Compute $\mathbf{\hat{z}}_0$ from $\mathbf{y, h}$, such that $[\mathbf{z}^{(y)} , \mathbf{z}^{(h)} ] = f (\mathbf{y, h})$ is E(3). 
-\State Subtract center of gravity (COG) from $\mathbf{\hat{z}}_0$
+\State Subtract center of geometry (COG) from $\mathbf{\hat{z}}_0$
 \Comment{Center reference at zero}
@@
 
-The core of the new method is the addition of a refinement step within the denoising process during runtime.
+The core of the new method is the addition of a refinement step within the denoising process during runtime of any pre-trained E(3) EDM.
 The resulting SILVR model produces conditional samples without any conditional training when generating new molecules.
@@
 The goal is to produce molecules that are complementary to a binding site based on exiting fragments and their set of atomic coordinates.
-%As stated before, the selective iterative latent variable refinement (SILVR) method makes use of the pre-trained equivariant diffusion model (EDM) as introduced by Hoogeboom \textit{et al.}~\cite{hoogeboom2022equivariant}.
 
@@
 The EDM by Hoodgeboom \textit{et al.} was trained on the 30 lowest energy conformations of 430,000 molecules from the GEOM dataset with explicit hydrogens~\cite{axelrod2022geom}.
+One appealing feature of GEOM is that it contains drug-like molecules including 6000 compounds for SARS-CoV-2 targets making this an appealing dataset for generating new molecules that combine or expand fragments for a SARS-CoV-2 target.
 For more details on the model theory and training refer to Hoogeboom \textit{et al.}~\cite{hoogeboom2022equivariant}.
 This model strictly only considers atomic coordinates (heavy atoms and hydrogen atoms), while all bond and molecular graph information is ignored.
-A more recently proposed version of the EDM incorporates molecular graph generation within the denoising diffusion model, and we believe it should be possible to incorporate SILVR into the denoising loop of this new model following the same principles we describe in this paper~\cite{vignac2023midi}.
+A more recently proposed version of the EDM incorporates molecular graph generation within the denoising diffusion model~\cite{vignac2023midi} improving the quality and potentially also synthetically accessible space of newly proposed molecules.
+We believe SILVR can be added into the denoising loop of this new model following the same principles we propose here.
 During training the atomic coordinates, and a one-hot encoding of their element, are passed through a forward diffusion process with iterative addition of Gaussian noise; both the coordinates and one-hot vector get diffused during this process.
@@
     \includegraphics[width=0.5\textwidth]{paper/Figures/Fig2/fig2.jpg}
-    \caption{Schematic illustrating the influence of SILVR of the molecules in the latent space.
+    \caption{Schematic illustrating the influence of applying the SILVR equation~\ref{eq:SILVR} to molecules in the latent space.
 At each denoising step if $r_{\mathrm{S}}>0$ the latent space vector gets scaled by the SILVR rate.
@@
 The EDM was adapted by introducing SILVR within the denoising process, algorithm~\ref{alg:silvr}, as outlined previously.
-At runtime, each atom of the reference set of coordinates is mapped to an atom in the EDM; this is simply achieved by constructing a reference tensor with the same shape as the EDM latent tensor, with the mapping being on a row-by-row basis.
-The reference coordinates are then translated such that their center of mass is at the origin.
-The reference coordinates are then diffused to the same timestep as that of the denoising process.
+At runtime, each atom of the reference set of coordinates is mapped to an atom in the EDM; this is achieved by constructing a reference tensor with the same shape as the EDM latent tensor, with the mapping being on a row-by-row basis.
+The reference coordinates are then translated such that their center of geometry is at the origin and diffused to the same timestep as that of the denoising process.
 That is, the amount of structure remaining from the reference should match the amount of structure formed by the generative process.
@@
 The diffusion of the reference to $t=\tau$ requires the sampling of a Gaussian; while in principle this sampling could be performed only once prior to the denoising loop, in practice better results were obtained when the Gaussian was repeatedly sampled during the denoising process.
-Once denoising is complete, sampled molecules can be translated back to the same coordinates as the reference by reverting the initial centre of gravity transformation; in the case of fragment data, this has the effect of returning samples to the binding site of the protein.
+Once denoising is complete, sampled molecules can be translated back to the same coordinates as the reference by reverting the initial centre of geometry transformation; in the case of fragment data, this has the effect of returning samples to the binding site of the protein.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\bibliographystyle{unsrt}
-\bibliography{SILVR}
+%\bibliographystyle{hunsrt}
+\printbibliography
 
@@
 
-The centre of gravity for the reference coordinates was initially aligned at zero.
+The centre of geometry for the reference coordinates was initially aligned at zero.
 Within the denoising loop, the EDM sampled new coordinates for all atoms.
@@
 The SILVR equation was applied as described in the main paper, only refining the atoms mapped to a reference.
-Through each iteration of denoising, the coordinates of the dummy atoms within the reference were continually updated, and the centre of gravity of these coordinates was re-aligned at zero.
-The total shift in the centre of gravity of the reference was tracked and subtracted from the final sampled molecule.
+Through each iteration of denoising, the coordinates of the dummy atoms within the reference were continually updated, and the centre of geometry of these coordinates was re-aligned at zero.
+The total shift in the centre of geometry of the reference was tracked and subtracted from the final sampled molecule.
 

### 2023-04-18 paper/manuscript.tex
Updates from Overleaf

@@
 The model allows the generation of new molecules that fit into a binding site of a protein based on fragment hits.
-We use the SARS-CoV-2 Main Protease fragments from Diamond X-Chem that form part of the COVID Moonshot project as a reference dataset for conditioning the molecular generation.
+We use the SARS-CoV-2 Main protease fragments from Diamond X-Chem that form part of the COVID Moonshot project as a reference dataset for conditioning the molecular generation.
 The SILVR rate controls the extent of conditioning and we show that moderate SILVR rates make it possible to generate new molecules of similar shape to the original fragments, meaning that the new molecules fit the binding site without knowledge of the protein.
@@
 Here the reference in blue on the left is 3 small fragments.
-They are evolved over time $t$ in the diffusion process to resemble a Gaussian distribution at $t=T$.
-The $\beta$ added represents the noise added at each step, and the multiple arrows in grey show the omitted steps from time $t=2$ to $t=T$.
+They are evolved over time $t$ in the diffusion process to resemble a Gaussian distribution at $t=T$, see Equation~\ref{eqn:noising}.
+The $\beta$ added represents the noise added at each step, and the dots show the omitted steps from time $t=3$ to $t=T$.
 As atoms effectively 'diffuse' they can be perceived as changing position.
-To generate a new sample of a single molecule that links all fragments (left most yellow box), making use of the reference fragments, the EDM is sampled at each denoising step, as indicated by the EDM arrows, but now the reference at that same level of noise is added back in through SILVR at a given rate $r_{\mathbf{S}}$, until a new sample that resembles the reference is generated (following the bottom row along the yellow arrow).}
+To generate a new sample a sample is generated from $p_{\theta}(x)$ according to Equation~\ref{eqn:denoising}, this distribution is from the learned EDM.
+At each denoising step, a set of reference fragments ($\mathbf{y}_t$) at that same level of noise $t$ is used which is indicated by the SILVR arrows to condition the EDM.
+This is controlled through SILVR at a given rate $r_{\mathbf{S}}$, until a new sample that resembles the reference is generated (following the bottom row along the yellow boxes and EDM arrows).}
     \label{fig:schematic}
@@
 In practice, this is done by training a neural network $\phi$ and generating samples $p(\mathbf{x}_{t-1}|\mathbf{x}_t)=\mathcal{N}(x_{t-1}; \mu_{\theta}(x_t;t), \sigma_t^2\mathbf{I})$ from  a Gaussian transition kernel with learned mean ($\mu_{\theta}(x_t;t)$) and variance ($\sigma_t$), with $\mathbf{x}_t$ being the data noised up to time $t$.
-Figure~\ref{fig:schematic} shows a schematic of the two main parts of such a DDPM, the diffusion part where noise is added for each timestep (shown in blue) and the denoising part (shown in yellow) that allows the generation of a new image (molecule).
+Figure~\ref{fig:schematic} shows a schematic of the two main parts of such a DDPM, the diffusion part, where noise is added for each timestep (shown in blue) and the denoising part (shown in yellow) that allows the generation of a new image (molecule).
 
@@
   q(\mathbf{x}_{1:T}|\mathbf{x}_0) = q(x_0)\prod_{t=1}^T q(\mathbf{x}_t|\mathbf{x}_{t-1}).
+  \label{eqn:noising}
 \end{equation}
-The product of conditional probabilities $q(..)$ can be modelled as a Gaussian Kernel given by:
+The product of conditional probabilities $q(\mathbf{x}_t | \mathbf{x}_{t-1})$ can be modelled as a Gaussian Kernel given by:
 \begin{equation}
-    q(\mathbf{x}_t, \mathbf{x}_s)=\mathcal{N}(x_t|\frac{\alpha_t}{\alpha_s}, \sigma_t^2-\frac{\alpha_t}{\alpha_s}\sigma^2_s),
+q(\mathbf{x}_t | \mathbf{x}_{t-1}) = \mathcal{N}(\mathbf{x}_t;\sqrt{1-\beta_t}\mathbf{x}_{t-1}, \beta_t\mathbf{I}),
+\end{equation}
+where the mean of the normal distribution is given by $ \sqrt{1-\beta_t}$ and $\beta_t$ a fixed variance. 
+In order to diffuse directly to timestep $s$ in the diffusion process the following shorthand is possible:
+\begin{equation}
+    q(\mathbf{x}_t, \mathbf{x}_s)=\prod_{t=1}^s q(\mathbf{x}_t|\mathbf{x}_{t-1})=\mathcal{N}(\mathbf{x}_t|\frac{\alpha_t}{\alpha_s}, \sigma_t^2-\frac{\alpha_t}{\alpha_s}\sigma^2_s),
 \end{equation}
@@
 % Next talk about the denoising step
-$p_{\theta}$ is the parametric probability distribution we want to learn. This generative process is given by:
+What we are actually interested in is learning the reverse process, i.e.
+the denoising and generating $\mathbf{x}$, however the reverse of the process $q(\mathbf{x}_{t-1}|\mathbf{x}_{t=T})$ is intractable.
+The DDPM learns this reverse transitions $p_{\theta}(\hat{\mathbf{x}}_{t-1}|\mathbf{x}_{t=T})$, which is also a Gaussian transition kernel.
+This generative (or denoising) process is given by:
 
 \begin{equation}
-p_{\theta}(\mathbf{x}_{t-1}|\mathbf{x}_t)=\mathcal{N}(x_{t-1}; \mu_{\theta}(x_t;t), \sigma_t^2\mathbf{I}).
+p_{\theta}(\hat{\mathbf{x}}_{t-1}|\mathbf{x}_{t=T})=\mathcal{N}(\hat{\mathbf{x}}_{t-1}; \mu_{\theta}(\mathbf{x}_t;T), \sigma_t^2\mathbf{I}).
+\label{eqn:denoising}
 \end{equation}
-A sample for denoising timestep $t-1$ can be generated from the following equation given the neural network $\phi$ that has been trained on the diffusion process: 
+$\mu_{\theta}$, is the learned mean and $\sigma$ represents a fixed variance for this transition process.
+A sample for denoising timestep $t-1$ can be generated from the following equation given the neural network $\phi(\mathbf{x}_t,t)$ that has been trained on the diffusion process: 
 \begin{equation}
- x_{t-1} = \frac{1}{\alpha_t}(x_t-\frac{1-\alpha_t}{\sqrt{1-\langle\alpha_t\rangle}}\phi(x_t,t)) +\sigma \mathbf{z},   
+ \hat{\mathbf{x}}_{t-1} = \frac{1}{\alpha_t}(\mathbf{x}_t-\frac{1-\alpha_t}{\sqrt{1-\langle\alpha_t\rangle}}\phi(\mathbf{x}_t,t)) +\sigma \rho,   
 \end{equation}
 
-with $\mathbf{z}\sim\mathcal{N}(0,\mathbf{I})$.
+with $\rho\sim\mathcal{N}(0,\mathbf{I})$.
 
-The process is iterated until $t=1$ and as such a new sample $x_0$ of the denoising process---which is intended to represent a proposed molecule design---is generated.
+The process is iterated until $t=1$ and as such a new sample $\mathbf{x}_0$ of the denoising process---which is intended to represent a proposed molecule design---is generated.
 
@@
 \subsection*{Equivariant diffusion model (EDM)}
-In the previous section introduced the neural network $\phi(\mthabf{x},t)$.
+In the previous section introduced the neural network $\phi(\mathbf{x}_t,t)$.
 In practice, it makes sense to use an euqivariant graph neural network, as it is a data efficient way to learn about molecules.
@@
 Satorras \textit{et al.}~\cite{satorras2022equivariant} have proposed an E(n) Equivariant Graph Neural Network on which the EDM by Hoogeboom \textit{et al.} builds.
+The actual diffusion process of the Hoogeboom model relies taking place on a set of points $M$ $\{\{\mathbf{x}_i,\mathbf{h}_i\}\}_{i\ldots M}$ on a latent variable that combines atom coordinates $\mathbf{x}_i$ and node features of each atom $\mathbf{h}_i$, such that: $\mathbf{z}_t = [\mathbf{z}_t^{(x), \mathbf{z}_t^{(h)}}$.
+The diffusion process similar to that of equation~\ref{eqn:noising} can be written as:
+\begin{equation}
+    q(\mathbf{z}_t| \mathbf{x},\mathbf{h}) = \mathcal{N}_{xh}(\mathbf{z}_t|\alpha_t[\mathbf{x,h}],\sigma^2_t\mathbf{I}].
+\end{equation}
+Similarly for the genreative denoising process the distribution can be written as:
+\begin{equation}
+    p_\theta(\mathbf{z}_s| \mathbf{z}_t) = \mathcal{N}_{xh}(\mathbf{z}_s | \mu_{\theta t\rightarrow s}( [ \mathbf{\hat{x},\hat{h}}],\mathbf{z}_t),\sigma_{t\rightarrow s}^2\mathbf{I}).
+\end{equation}
+This is the equivlanet of equation~\ref{eqn:denoising}, using $\mathbf{\hat{x}}, \mathbf{\hat{h}}$ as the data variables that are estimated by the neural network.
+The neural network $\phi(\mathbf{z}_t,t)$ outputs an auxillary variable $\hat\epsilon = [\hat{\epsilon}^{(x)},\hat{\epsilon}^{(h}]$, from which $\mathbf{\hat{x}}, \mathbf{\hat{h}}$ can be recovered as:
+\begin{equation}
+    [\mathbf{\hat{x}}, \mathbf{\hat{h}}] = \frac{\mathbf{z}_t}{\alpha_t}-\hat{\epsilon}_t\frac{\sigma_t}{\alpha_t}.
+\end{equation}
+We use the notation $\kappa$ to generate a sample [\mathbf{\hat{x}}, \mathbf{\hat{h}}] from the EDM. 
 For full details on the architecture and their code see~\cite{hoogeboom2022equivariant}. 
@@
 \subsection{Iterative Latent Variable Refinement as a conditioning tool}
-Choi \textit{et al.}~\cite{choi2021ilvr} introduced a way to condition DDPMs by sampling their images from the conditional distribution $p(x_0|c)$, with the condition c:
+Conditioning diffusion models is often desirable to for example generate similar images to that of an original input images, or in the example of Hoogeboom generating molecules in the presence of an external electric field resulting in more polarizable molecules.
+However, this conditioning requires retraining the network to accommodate for this condition.
+Choi \textit{et al.}~\cite{choi2021ilvr} introduced a way to condition DDPMs without having to retrain the neural network.
+In the generative process it is possible to introduce a condition $c$ using a conditional distribution $p(x_0|c)$:
 \begin{equation}
-    P_\theta(x_0|c) = \int p_\theta(x_{0:T}|c)Dx_{1:T}
+    p_\theta(x_0|c) = \int p_\theta(x_{0:T}|c)dx_{1:T}
 \end{equation}
 \begin{equation}
-    p_{\theta}(x_{0:T}|c)=p(x_T)\prod_{t=1}^{T}p_{\theta}(x_{t-1}x_t,c)
+    p_{\theta}[x_{\{0:T\}}|c]=p(x_T)\prod_{t=1}^{T}p_{\theta}(x_{t-1}|x_t,c)
+\end{equation}
+Their trick is to use a reference image $\mathbf{y}$ and place it in the same downsampled fillter $\psi(\mathbf{y})$ as the generated image $\psi(\mathbf{\hat{x}}_0)$.
+This means the original image and the reference are  in the same latent space.
+So in each denoising step the proposal distribution is matched with that of the reference $y$ noised up to the appropriate timestep.
+An unconditional distribution at $t$ is generated first:
+\begin{equation}
+\mathbf{\hat{x}}_{t-1} \sim p_{\theta}( \mathbf{\hat{x}}|\mathbf{x}_t),
+\end{equation}
+then this  new sample is 'adjusted' according to:
+\begin{equation}
+    \mathbf{\hat{x}}^{\prime}_{t-1} = \psi(\mathbf{y}_{t-1}) + (I-\psi)(\mathbf{\hat{x}}_{t-1})
 \end{equation}
 This iterative latent variable refinement (ILVR) means that the condition can be applied during the denoising steps without additional training.
-In the case of Choi et al, they used downsampled reference images as conditioning in order to generate novel images similar to the reference.
+In the case of Choi \textit{et al.}~\cite{choi2021ilvr}, they used downsampled reference images as conditioning in order to generate novel images similar to the reference.
 
@@
 \label{sec:silvr}
-We propose a new method, that combines ILVR by Choi \textit{et al.}~\cite{choi2021ilvr} proposed in an image generation context and the EDM by Hoogeboom \textit{et al.}.
+We propose a new method, that combines ideas from ILVR by Choi \textit{et al.}~\cite{choi2021ilvr} proposed in an image generation context and the EDM by Hoogeboom \textit{et al.}~\cite{hoogeboom2022equivariant}.
 This allows us to generate a selective iterative latent variable refinement (SILVR) procedure in which we introduce information of a reference molecule into the denoising process.
@@
 The reference can be a single molecule or a series of superimposed fragments, and additional unguided dummy atoms can also be defined at the beginning of the denoising process.
-As a result, we propose the following algorithm for SILVR~\ref{alg:silvr}.
-We consider latent space variables $\mathbf{z}=[\mathbf{z}^{(x)},\mathbf{z}^{(h)}]$ for the standard denoising process and  $\mathbf{\hat{z}} = [\mathbf{\hat{z}}^{(y)},\mathbf{\hat{z}}^{(h)}]$, for the set of reference coordinates given by $\mathbf{y}$.
-The vector $\mathbf{h}$ contains all the scalar node properties of the equivariant EDM.
+We consider latent space variables $\mathbf{z}=[\mathbf{z}^{(x)},\mathbf{z}^{(h)}]$ for the standard denoising process and  $\mathbf{\tilde{z}} = [\mathbf{z}^{(y)},\mathbf{z}^{(h)}]$, for the set of reference coordinates given by $\mathbf{y}$.
+The vector $\mathbf{h}$ contains all the scalar node properties of the equivariant EDM for the reference.
+Similarly to Choi \texit{et al.}, we update the diffusion process at noise level $t$ in the latent space $\mathbf{z}$, with the reference $\mathbf{\tilde{z}}$ using a factor, or vector if used at variable rates for different atoms $r_{\mathrm{S}}$.
+We call $r_{\mathrm{S}}$ the SILVR rate and this leads to an overall update or conditioning towards a reference at each step in the generative denoising process according to the SILVR equation~\ref{eqn:SILVR}.
+\begin{equation}
+\mathbf{z}_{t-1}^{\prime} = \mathbf{z}_{t-1}-\alpha_{t-1}r_{\mathrm{S}}\mathbf{z}_{t-1}+r_{\mathrm{S}}\mathbf{\tilde{z}}_{t-1}
+    \label{eqn:SILVR}
+\end{equation}
+As a result, we propose the following algorithm for the generation of condition samples according to SILVR~\ref{alg:silvr}.
 \begin{algorithm}
@@
 
-\State Compute $\mathbf{\hat{z}}_0$ from $\mathbf{y, h}$, such that $[\mathbf{z}^{(y)} , \mathbf{z}^{(h)} ] = f (\mathbf{y, h})$ is E(3). 
-\State Subtract center of geometry (COG) from $\mathbf{\hat{z}}_0$
+\State Compute $\mathbf{\tilde{z}}_0$ from $\mathbf{y, h}$, such that $[\mathbf{z}^{(y)} , \mathbf{z}^{(h)} ] = f (\mathbf{y, h})$ is E(3). 
+\State Subtract center of geometry (COG) from $\mathbf{\tilde{z}}_0$
 \Comment{Center reference at zero}
@@
     \State Subtract COG from $\boldsymbol{\epsilon}^{(x)}$
-    \State $\mathbf{\hat{z}}_{t-1} = \alpha_{t-1} \mathbf{\hat{z}}_{0} +\sigma_{t-1}\times\boldsymbol{\epsilon}$
+    \State $\mathbf{\hat{z}}_{t-1} = \alpha_{t-1} \mathbf{\tilde{z}}_{0} +\sigma_{t-1}\times\boldsymbol{\epsilon}$
     \Comment{Noise reference to t=t-1}
@@
     \Comment{Compute denoising step}
-    \State Update $\mathbf{z}_{t-1} \gets \mathbf{z}_{t-1} - \alpha_{t-1}r_{\mathrm{S}} \mathbf{z}_{t-1} +r_{\mathrm{S}} \mathbf{\hat{z}}_{t-1}$ \Comment{SILVR equation}
+    \State Update $\mathbf{z}^{\prime}_{t-1} \gets \mathbf{z}_{t-1} - \alpha_{t-1}r_{\mathrm{S}} \mathbf{z}_{t-1} +r_{\mathrm{S}} \mathbf{\tilde{z}}_{t-1}$ \Comment{SILVR equation}
     \State Subtract COG from updated $\mathbf{z}_{t-1}$
@@
 The resulting SILVR model produces conditional samples without any conditional training when generating new molecules.
-Figure~\ref{fig:silvr_explanation} shows an illustrative example of how the SILVR rate $r_{\mathrm{S}}$ is used to shift the latent space vector $\mathbf{z}_{t-1}$ at any point in the denoising process from $T\ldots t=1$.
-Using the SILVR equation as defined in the SILVR algorithm~\ref{alg:silvr} line 11, an existing denoising step (green) is brought closer to the reference (purple) in latent space.
+Figure~\ref{fig:silvr_explanation} shows an illustrative example of how the SILVR rate $r_{\mathrm{S}}$ is used to shift the latent space vector $\mathbf{z}_{t-1}$ at any point in the denoising process from $T\ldots t=1$, here in 2D for illustration purposes.
+Using the SILVR equation~\ref{eqn:SILVR}, an existing denoising step (light blue) is brought closer to the reference (purple) in the latent space according to a scaled version of purple using $r_{\mathrm{S}}$ (green).
 
@@
     \includegraphics[width=0.5\textwidth]{paper/Figures/Fig2/fig2.jpg}
-    \caption{Schematic illustrating the influence of applying the SILVR equation~\ref{eq:SILVR} to molecules in the latent space.
+    \caption{Schematic illustrating the influence of applying the SILVR equation~\ref{eqn:SILVR} to molecules in the latent space.
 At each denoising step if $r_{\mathrm{S}}>0$ the latent space vector gets scaled by the SILVR rate.
@@
 \subsection{Reference Dataset: COVID Moonshot}
-Reference molecules were selected from the original 23 non-covalent hits of the SARS-CoV-2 Main Protease (Mpro) identified by the XChem fragment screen~\cite{douangamath2020crystallographic} as part of the Covid Moonshot Project~\cite{consortium2023open, consortium2021open}.
+Reference molecules were selected from the original 23 non-covalent hits of the SARS-CoV-2 Main Protease (Mpro) identified by the XChem fragment screen~\cite{douangamath2020crystallographic} as part of the COVID Moonshot Project~\cite{consortium2023open, consortium2021open}.
 A more detailed picture of all fragments is presented in Figure~\ref{fig:moonshot_data} of the Supplementary information(SI).
@@
 \begin{enumerate}
-    \item Using three distinct fragments with substantial overlap as a reference to generate new samples
+    \item Using three distinct fragments with substantial overlap as a reference to generate new samples.
     \item Using two fragments with a slight overlap to generate a single new sample.
-    \item And using two fragments that are disconnected to investigate linker generation
+    \item And using two fragments that are disconnected to investigate linker generation.
 \end{enumerate}
@@
 \subsubsection{Shape complementarity of sample to protein}
-The agreement in the shape of samples and the main protease binding site was determined using openeye-toolkits version 2022.2.2 Gaussian scoring function \textit{shapegauss}~\cite{openeyescientificsoftwareinc.oedocking,kelley2015posit}.
+The agreement in the shape of samples and the MPro binding site was determined using openeye-toolkit version 2022.2.2 Gaussian scoring function \textit{shapegauss}~\cite{openeyescientificsoftwareinc.oedocking,kelley2015posit}.
 This scoring function measures the shape complementarity between the ligand and receptor by considering each heavy atom as a Gaussian function~\cite{mcgann2003gaussian}.
@@
 The protein receptor file was prepared from the Mpro-x0072 crystal structure with removal of the ligand.
-The \texttt{XYZ} coordinates of samples were directly read into openeye-toolkit and the pose was re-scored with \textit{shapegauss}.
+The \texttt{XYZ} coordinates of samples were directly read into the openeye-toolkit and the pose was re-scored with \textit{shapegauss}.
 
@@
     \item Do the generated samples of new molecules still fit into the Mpro binding site?
-    \item Can we link molecule fragments without additional training?
+    \item Can we link molecule fragments without incorporating binding site information as part of additional training?
 \end{enumerate}
@@
 \subsection{The SILVR rate $r_{\mathrm{S}}$ effectively modulates similarity to the reference structures}
-Qualitatively, the generated molecular samples from the condition EDM  show a clear resemblance to their reference structures, with similarity increasing with $r_{\mathrm{S}}$.
+Qualitatively, the generated molecular samples from the conditioned EDM  show a clear resemblance to their reference structures, with similarity increasing with $r_{\mathrm{S}}$.
 Figure S~\ref{fig:examplese_curated}, in the SI, shows two example samples started from fragments \texttt{x0072} and \texttt{x0354} over a range of SILVR rates between $r_{\mathrm{S}}=0$ to $r_{\mathrm{S}}=0.02$.
@@
 At low values of $r_{\mathrm{S}}$ ($< 0.0025$) the sampled molecules only show an approximate agreement in orientation.
-At medium values of $r_{\mathrm{S}}$ ($0.0025 \le r_{\mathrm{S}} < 0.01$) the resulting samples begin to produce key structural features like ring systems and heteroatoms at positions seen in the reference.
+At medium values of $r_{\mathrm{S}}$ ($0.0025 \le r_{\mathrm{S}} < 0.01$) the resulting samples begin to produce key structural features such as ring systems and heteroatoms at positions seen in the reference.
 At high values of $r_{\mathrm{S}}$ ( $\ge 0.01$) the resulting samples have a very high resemblance to the reference with most structural features in correct positions, however, the diversity of samples is significantly reduced and structures start to become chemically less reasonable.
@@
 We generated 1000 samples at different $r_{\mathrm{S}}$ using fragments \texttt{X0072} and \texttt{X0354} as a reference.
-Figure~\ref{fig:measures} summarises the findings from these experiments with box plots generated across the 1000 samples.
+Figure~\ref{fig:measures} summarises the findings from these experiments with violin plots generated across the 1000 samples.
 The zeroth test we made with the generated samples was looking at how many generated molecules were fragmented i.e.
@@
 
-\begin{figure}[htb!]
+\begin{figure}[ht!]
     \centering
-    \includegraphics[width=\textwidth]{paper/Figures/Fig3/fig_3_metrics_violins.png}
+    \includegraphics[width=0.8\textwidth]{paper/Figures/Fig3/fig_3_metrics_violins.png}
     \caption{Validation measures of SILVR model using fragments \texttt{x0072} and \texttt{x0354} as reference coordinates.
@@
 C: OpenEye measure \textit{shapegauss} - a Gaussian scoring function describing the shape fit between Mpro and samples, ignoring chemical interactions.
-A lower score means...
-D: Geometry stability - AIMNet geometry optimisation was completed with Auto3D using the SMILES string of each sample.
+A lower score means a better shape fit of the molecule D: Geometry stability - AIMNet geometry optimisation was completed with Auto3D using the SMILES string of each sample.
 RMSD was calculated between the predicted geometry and the sampled geometry using RDKit.
@@
 From Figure~\ref{fig:measures} C it can be seen that the shape complementarity of samples improves with increasing $r_{\mathrm{S}}$, demonstrating that SILVR can produce ligands of binding site geometry when guided by the coordinates of fragment molecules.
-The lack of plot data for $r_{\mathrm{S}}=0.02$ and $r_{\mathrm{S}}=0.03$ was because all \textit{shapegauss} calculations failed.
+The lack of data for $r_{\mathrm{S}}=0.02$ and $r_{\mathrm{S}}=0.03$ was because all \textit{shapegauss} calculations failed.
 We believe this was due to the atom coordinates representing highly strained and internally clashing molecules, and the scoring algorithm either failed to read the molecules or identified them as bad conformations.
 
-\begin{figure}[!htb]
+\begin{figure}[!ht]
     \centering
-    \includegraphics[width=\textwidth]{paper/Figures/Fig4/fig_4_ref_sample_bound.png}
+    \includegraphics[width=0.8\textwidth]{paper/Figures/Fig4/fig_4_ref_sample_bound.png}
     \caption{Examples of generated molecules from different experiments testing different overlap models.
@@
 This is illustrated in the example of Figure~\ref{fig:samples}E and F.
-While it was possible to use SILVR as described in the Theory section, better results were obtained with the addition of \textit{dummy atoms}.
+While it was possible to use SILVR as described in the theory section, better results were obtained with the addition of \textit{dummy atoms}.
 These are atoms which are present in the EDM without a mapping to a reference atom, and so are free to explore the whole coordinate space without guidance from SILVR.
@@
 
+% Something on dummy atoms here also reference figure: fig:dummy_atoms
+
 \section{Discussion and outlook}
 \label{sec:discussion}
-SILVR, as presented, represents a method in which a general EDM can be conditioned to generate samples that resemble a reference structure, without any additional training needed.
+SILVR, as presented, represents a method in which a general equivariant diffusion model (EDM) can be conditioned to generate samples that resemble a reference structure, without any additional training needed.
 We showed that SILVR can complete both fragment merging and linking type tasks, without any \textit{a priori} knowledge of these design challenges.
@@
 
-\subsection{The inclusion of electrostatic information from proteins will likely improve the newly generated models}
-To answer structure-based drug design, a scoring function of protein surface potentials could be used to guide atoms towards energetically favoured pocket regions; issues of synthetic accessibility might be solved by clever design of refinement methods; and interfacing with alternative data sources (such as NMR) might be achievable.
+%\subsection{The inclusion of electrostatic information from proteins will likely improve the newly generated models}
+%To answer structure-based drug design, a scoring function of protein surface potentials could be used to guide atoms towards energetically favoured pocket regions; issues of synthetic accessibility might be solved by clever design of refinement methods; and interfacing with alternative data sources (such as NMR) might be achievable.
 
 \subsection{Placements of hydrogens and dummy atoms needs additional trials}
-An EDM with explicity hydrogens will improve the overall models.
+An EDM with explicit hydrogens will improve the overall models.
 At the moment there is a mixing of explicit and implicity hydrogens depending on need for analysis and input.
@@
 
-\section{Summary of figures included in the supporting information}
+\section{Summary of additional figures }
 The supporting information consists of a series of figures in addition to the figures in the main paper.
-Figure~\ref{fig:moonshot_data} contains the 2D structures of all fragments used in the experiments as references.
-In Figure~\ref{fig:examples_curated} an example of two curated experiments is shown with different SILVR rates $r_{mathrm{s}}$.
-Next, Figure~\ref{fig:fragmentation}, shows on average how many molecules were not-fragmented in a set of 1000 samples for different SILVR rates.
-Figure~\ref{fig:SA_QED} includes synthetical accessibility(~\cite{ertl2009estimation}) and QED~\cite{bickerton2012quantifying, wildman1999prediction}.
+Figure S~\ref{fig:moonshot_data} contains the 2D structures of all fragments used in the experiments as references.
+In Figure S~\ref{fig:examples_curated} an example of two curated experiments is shown with different SILVR rates $r_{\mathrm{s}}$.
+Next, Figure S~\ref{fig:fragmentation}, shows on average how many molecules were not-fragmented in a set of 1000 samples for different SILVR rates.
+Figure S~\ref{fig:SA_QED} includes synthetical accessibility(~\cite{ertl2009estimation}) and QED~\cite{bickerton2012quantifying, wildman1999prediction}.
 An example of how far atoms, even involved in a reference structure are displaced in the denoising diffusion process is shown in an example in Figure~\ref{fig:atom_displacement} \textbf{A} for a whole molecule generated and in \textbf{B} for the heterocycle.
@@
     \centering
-    \includegraphics[width=0.8\textwidth]{paper/Figures/FigS4/si_fig_2d_measures.png}
-    \caption{A: The synthetic accessibility score (SA score) estimates the synthetic feasibility of molecules based on fragments contributions.~\cite{ertl2009estimation} This was calculated for all non-fragmented samples using an RDKit implementation of the scoring function SAScorer.
+    \includegraphics[width=0.8\textwidth]{paper/Figures/FigS4/si_fig_2d_measures_violins.png}
+    \caption{A: The synthetic accessibility score (SA score) estimates the synthetic feasibility of molecules based on fragments contributions.~\cite{ertl2009estimation} This was calculated for all non-fragmented samples using an RDKit implementation of the scoring function SAScorer~\cite{ertl2009estimation}.
 A lower score indicates an easier-to-synthesise molecule: most catalogue and bioactive molecules fall in the range 2-5, while a score greater than 7 represents the upper end of complexity for natural products.~\cite{ertl2009estimation} B: The Quantitative Estimate of Druglikeness (QED) score combines a selection of descriptors such as molecular weight and calculated Log(P) to estimate the drug-likeness of a molecule.~\cite{bickerton2012quantifying, wildman1999prediction} This was calculated using RDKit with default settings across all non-fragmented samples.

### 2023-04-18 paper/manuscript.tex
Draft version before fixing refs.

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\author{Nicholas Runcie}
+\author{Nicholas T. Runcie}
 \altaffiliation{EaSTCHEM School of Chemistry, University of Edinburgh, EH9 3FJ}
@@
 \begin{abstract}
-Computationally generating novel synthetically accessible compounds with high affinity and low toxicity is a great challenge in drug-design.
+Computationally generating novel synthetically accessible compounds with high affinity and low toxicity is a great challenge in drug design.
 Machine-learning models beyond conventional pharmacophoric methods have shown promise in generating novel small molecule compounds, but require significant tuning for a specific protein target.
-Here, we introduce a method called selective iterative latent variable refinement (SILVR) for conditioning an existing a diffusion-based equivariant generative model without retraining.
+Here, we introduce a method called selective iterative latent variable refinement (SILVR) for conditioning an existing diffusion-based equivariant generative model without retraining.
 The model allows the generation of new molecules that fit into a binding site of a protein based on fragment hits.
-We use the SARS-CoV-2 Main protease fragments from Diamond X-Chem that form part of the COVID Moonshot project as a reference dataset for conditioning the molecular generation.
+We use the SARS-CoV-2 Main protease fragments from Diamond X-Chem that form part of the COVID Moonshot project as a reference dataset for conditioning the molecule generation.
 The SILVR rate controls the extent of conditioning and we show that moderate SILVR rates make it possible to generate new molecules of similar shape to the original fragments, meaning that the new molecules fit the binding site without knowledge of the protein.
 We can also merge up to 3 fragments into a new molecule without affecting the quality of molecules generated by the underlying generative model.
-Our method is generalizable to any protein target with known fragments and any diffusion based model for molecule generation.
+Our method is generalizable to any protein target with known fragments and any diffusion-based model for molecule generation.
 \end{abstract}
@@
 \label{sec:intro}
-Sampling from a very large space of possible drug-like compounds to find suitable hits for a given target protein is a open challenge in drug-design.
+Sampling from a very large space of possible drug-like compounds to find suitable hits for a given target protein is an open challenge in drug-design.
 It is estimated that there are between $10^{23}$ to $10^{60}$ feasible compounds, while only around $10^8$ have been synthesised so far~\cite{polishchuk2013estimation, reymond2012exploring}.
@@
 Satorras \textit{et al.}~\cite{satorras2022equivariant} have proposed an E(n) Equivariant Graph Neural Network on which the EDM by Hoogeboom \textit{et al.} builds.
-The actual diffusion process of the Hoogeboom model relies taking place on a set of points $M$ $\{\{\mathbf{x}_i,\mathbf{h}_i\}\}_{i\ldots M}$ on a latent variable that combines atom coordinates $\mathbf{x}_i$ and node features of each atom $\mathbf{h}_i$, such that: $\mathbf{z}_t = [\mathbf{z}_t^{(x), \mathbf{z}_t^{(h)}}$.
-The diffusion process similar to that of equation~\ref{eqn:noising} can be written as:
+The actual diffusion process of the Hoogeboom model relies taking place on a set of points $M$ $\{\{\mathbf{x}_i,\mathbf{h}_i\}\}_{i\ldots M}$ on a latent variable that combines atom coordinates $\mathbf{x}_i$ and node features of each atom $\mathbf{h}_i$, such that: $\mathbf{z}_t = [\mathbf{z}_t^{(x)}, \mathbf{z}_t^{(h)}]$.
+The node features in practice are an array of values containing information such as atom type.
+These features are encoded using a onehot encoding.
+For more details on this see~\cite{hoogeboom2022equivariant}.
+Based on this latent variable $\mathbf{z}_t$, the diffusion process can be defined similarly to that of equation~\ref{eqn:noising} as:
 \begin{equation}
@@
 \end{equation}
-Similarly for the genreative denoising process the distribution can be written as:
+In the same way, for the genereative denoising process the distribution can be written as:
 \begin{equation}
@@
 \end{equation}
-This is the equivlanet of equation~\ref{eqn:denoising}, using $\mathbf{\hat{x}}, \mathbf{\hat{h}}$ as the data variables that are estimated by the neural network.
-The neural network $\phi(\mathbf{z}_t,t)$ outputs an auxillary variable $\hat\epsilon = [\hat{\epsilon}^{(x)},\hat{\epsilon}^{(h}]$, from which $\mathbf{\hat{x}}, \mathbf{\hat{h}}$ can be recovered as:
+This is the equivalent of equation~\ref{eqn:denoising}, using $\mathbf{\hat{x}}, \mathbf{\hat{h}}$ as the data variables that are estimated by the neural network.
+The neural network $\phi(\mathbf{z}_t,t)$ outputs an auxiliary variable $\hat\epsilon = [\hat{\epsilon}^{(x)},\hat{\epsilon}^{(h}]$, from which $\mathbf{\hat{x}}, \mathbf{\hat{h}}$ can be recovered as:
 \begin{equation}
@@
 \end{equation}
-We use the notation $\kappa$ to generate a sample [\mathbf{\hat{x}}, \mathbf{\hat{h}}] from the EDM. 
-For full details on the architecture and their code see~\cite{hoogeboom2022equivariant}. 
+We use the notation $\kappa$ to generate a sample $\kappa = [\mathbf{\hat{x}}, \mathbf{\hat{h}}]$ from the EDM. 
+For more details details on the architecture and Hoogeboom's code see~\cite{hoogeboom2022equivariant}. 
 %Something on latent variables Z.
@@
 The reference can be a single molecule or a series of superimposed fragments, and additional unguided dummy atoms can also be defined at the beginning of the denoising process.
-We consider latent space variables $\mathbf{z}=[\mathbf{z}^{(x)},\mathbf{z}^{(h)}]$ for the standard denoising process and  $\mathbf{\tilde{z}} = [\mathbf{z}^{(y)},\mathbf{z}^{(h)}]$, for the set of reference coordinates given by $\mathbf{y}$.
+We consider latent space variables $\mathbf{z}=[\mathbf{z}^{(x)},\mathbf{z}^{(h)}]$ for the standard denoising process and  $\mathbf{\tilde{z}} = [\mathbf{z}^{(y)},\mathbf{z}^{(h_y)}]$, for the set of reference coordinates given by $\mathbf{y}$.
 The vector $\mathbf{h}$ contains all the scalar node properties of the equivariant EDM for the reference.
-Similarly to Choi \texit{et al.}, we update the diffusion process at noise level $t$ in the latent space $\mathbf{z}$, with the reference $\mathbf{\tilde{z}}$ using a factor, or vector if used at variable rates for different atoms $r_{\mathrm{S}}$.
+Similarly to Choi \textit{et al.}, we update the diffusion process at noise level $t$ in the latent space $\mathbf{z}$, with the reference $\mathbf{\tilde{z}}$ using a factor, or vector if used at variable rates for different atoms $r_{\mathrm{S}}$.
 We call $r_{\mathrm{S}}$ the SILVR rate and this leads to an overall update or conditioning towards a reference at each step in the generative denoising process according to the SILVR equation~\ref{eqn:SILVR}.
@@
 \begin{algorithmic}[1]
-\State \textbf{Input:} Reference molecule $\mathbf{y,h}$, EDM $\kappa$ 
+\State \textbf{Input:} Reference molecule $\mathbf{y,h}_y$, EDM $\kappa$ 
 \State \textbf{Output:} Generated molecule $\mathbf{x,h}$
 
-\State Compute $\mathbf{\tilde{z}}_0$ from $\mathbf{y, h}$, such that $[\mathbf{z}^{(y)} , \mathbf{z}^{(h)} ] = f (\mathbf{y, h})$ is E(3). 
+\State Compute $\mathbf{\tilde{z}}_0$ from $\mathbf{y, h}_h$, such that $[\mathbf{z}^{(y)} , \mathbf{z}^{(h_y)} ] = f (\mathbf{y, h}_y)$ is E(3). 
 \State Subtract center of geometry (COG) from $\mathbf{\tilde{z}}_0$
@@
     \State Subtract COG from $\boldsymbol{\epsilon}^{(x)}$
-    \State $\mathbf{\hat{z}}_{t-1} = \alpha_{t-1} \mathbf{\tilde{z}}_{0} +\sigma_{t-1}\times\boldsymbol{\epsilon}$
+    \State $\mathbf{\tilde{z}}_{t-1} = \alpha_{t-1} \mathbf{\tilde{z}}_{0} +\sigma_{t-1}\times\boldsymbol{\epsilon}$
     \Comment{Noise reference to t=t-1}
@@
     \State Update $\mathbf{z}^{\prime}_{t-1} \gets \mathbf{z}_{t-1} - \alpha_{t-1}r_{\mathrm{S}} \mathbf{z}_{t-1} +r_{\mathrm{S}} \mathbf{\tilde{z}}_{t-1}$ \Comment{SILVR equation}
-    \State Subtract COG from updated $\mathbf{z}_{t-1}$
+    \State Subtract COG from updated $\mathbf{z}^{\prime}_{t-1}$
 
 \EndFor
-\State Add COG($\mathbf{\hat{z}}_0$) to $\mathbf{z}_0$
+\State Add COG($\mathbf{\tilde{z}}_0$) to $\mathbf{z}^{\prime}_0$
 \Comment{Move sample to original position of reference}
-\State Sample $\mathbf{x,h} \sim p(\mathbf{x,h|z_0})$ 
+\State Sample $\mathbf{x,h} \sim p_{\theta}(\mathbf{x,h}|\mathbf{z}^{\prime}_0)$ \Comment{See equation~\ref{eqn:denoising}}
 \end{algorithmic}
@@
 Using the SILVR equation~\ref{eqn:SILVR}, an existing denoising step (light blue) is brought closer to the reference (purple) in the latent space according to a scaled version of purple using $r_{\mathrm{S}}$ (green).
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\textwidth]{paper/Figures/Fig2/fig2.jpg}
+    \caption{Schematic illustrating the influence of applying the SILVR equation~\ref{eqn:SILVR} to molecules in the latent space.
+At each denoising step if $r_{\mathrm{S}}>0$ the latent space vector (light blue) gets shifted according to a scaled reference vector (purple) using the SILVR rate (light green).
+This results in an updated latent vector (dark blue).
+This is done from the first denoising timestep (top) until the last one (bottom).
+Repeatedly applying SILVR will result in a molecule that resembles the references for $r_{\mathrm{S}}>0$ (right) and does not for $r_{\mathrm{S}}=0$ (left)}
+    \label{fig:silvr_explanation}
+\end{figure}
 
@@
 
-\begin{figure}
-    \centering
-    \includegraphics[width=0.5\textwidth]{paper/Figures/Fig2/fig2.jpg}
-    \caption{Schematic illustrating the influence of applying the SILVR equation~\ref{eqn:SILVR} to molecules in the latent space.
-At each denoising step if $r_{\mathrm{S}}>0$ the latent space vector gets scaled by the SILVR rate.
-This is done from the first denoising timestep (top) until the last one (bottom).
-Repeatedly applying SILVR will result in a molecule that resembles the references for $r_{\mathrm{S}}>0$ (right) and does not for $r_{\mathrm{S}}=0$ (left)}
-    \label{fig:silvr_explanation}
-\end{figure}
-
 The EDM was adapted by introducing SILVR within the denoising process, algorithm~\ref{alg:silvr}, as outlined previously.
@@
 That is, the amount of structure remaining from the reference should match the amount of structure formed by the generative process.
-A small refinement is applied to add information from the reference coordinates to the latent variable of the denoising process.
+A small refinement is applied to add information from the reference coordinates to the latent variable of the denoising process (see line 11 of the algorithim and equation~\ref{eqn:SILVR}.
 This equation has the effect of shrinking the coordinates towards the origin and then expanding the coordinates out in the direction of the reference, see Figure~\ref{fig:silvr_explanation}.
 Importantly, the extent of this refinement is defined by the SILVR rate $r_{\mathrm{S}}$, with $r_{\mathrm{S}}=0$ providing no additional refinement and $r_{\mathrm{S}}=1$ resulting in a total replacement of atoms.
-The diffusion of the reference to $t=\tau$ requires the sampling of a Gaussian; while in principle this sampling could be performed only once prior to the denoising loop, in practice better results were obtained when the Gaussian was repeatedly sampled during the denoising process.
+The diffusion of the reference to $t$ requires the sampling of a Gaussian; at each step of the denoising loop the reference is repeatedly sampled.
 Once denoising is complete, sampled molecules can be translated back to the same coordinates as the reference by reverting the initial centre of geometry transformation; in the case of fragment data, this has the effect of returning samples to the binding site of the protein.
@@
 Reference molecules were selected from the original 23 non-covalent hits of the SARS-CoV-2 Main Protease (Mpro) identified by the XChem fragment screen~\cite{douangamath2020crystallographic} as part of the COVID Moonshot Project~\cite{consortium2023open, consortium2021open}.
-A more detailed picture of all fragments is presented in Figure~\ref{fig:moonshot_data} of the Supplementary information(SI).
+A more detailed picture of all fragments is presented in Figure S~\ref{fig:moonshot_data} of the Supplementary information(SI).
 Fragments were visualised using NGLview version 3.03 and combinations were arbitrarily selected as test cases for the following different experimental settings for trying to understand the similarity between the reference and the new sample.
@@
 \begin{enumerate}
-    \item Using three distinct fragments with substantial overlap as a reference to generate new samples.
+    \item Using three distinct fragments with substantial overlap as a reference to generate a new sample.
     \item Using two fragments with a slight overlap to generate a single new sample.
@@
 Fragments \texttt{x0072} and \texttt{x0354} were selected for benchmarking the effect of the SILVR rate $r_{\mathrm{S}}$ on sampling; \texttt{x1093}, \texttt{x0072}, \texttt{x2193} were randomly chosen to represent three significantly overlapping fragments; \texttt{x0434}, \texttt{x0305} and \texttt{x0072}, \texttt{x0354} were used as partially overlapping fragments; and \texttt{x0874}, \texttt{x0397} were used as two disconnected fragments, resembling a linker design type problem.
-The bonding information of selected fragments was deleted and cartesian coordinates were combined into a single \texttt{XYZ} file.
+The bonding information of selected fragments was deleted and Cartesian coordinates were combined into a single \texttt{XYZ} file.
 Values of $r_{\mathrm{S}}$ were selected and added to the \texttt{XYZ} file to create a reference file containing all experiment setup information.
@@
 \subsubsection{Geometric stability - Auto3D}
-To determine whether samples represent a true molecular geometry, an independent minimisation of molecular geometries was performed using Atoms In Molecules Neural Network Potential (AIMNet) with Auto3D.
+To determine whether samples represent a true molecular geometry, an independent minimisation of molecular geometries was performed using Atoms In Molecules Neural Network Potential (AIMNet) with Auto3D~\cite{liu2022auto3d}.
 All samples were read by RDKit and samples containing more than one molecule were removed from the test set.
@@
 
-\subsubsection{Shape complementarity of sample to protein}
+\subsubsection{Shape complementarity of the generated sample to the protein}
 The agreement in the shape of samples and the MPro binding site was determined using openeye-toolkit version 2022.2.2 Gaussian scoring function \textit{shapegauss}~\cite{openeyescientificsoftwareinc.oedocking,kelley2015posit}.
@@
 
+\begin{figure}[ht!]
+    \centering
+    \includegraphics[width=0.8\textwidth]{paper/Figures/Fig3/fig_3_metrics_violins.png}
+    \caption{Validation measures of SILVR model using fragments \texttt{x0072} and \texttt{x0354} as reference coordinates.
+A: Ratio of stable atoms - an atom is determined as stable if the valence matches the expected valence for the element B: RMSD from reference - the calculated RMSD between the reference and sample, using an absolute one-to-one mapping ignore atom identity with low RMSD meaning molecules are similar to the reference and high RMSD they are not.
+C: OpenEye measure \textit{shapegauss} - a Gaussian scoring function describing the shape fit between Mpro and samples, ignoring chemical interactions.
+A lower score means a better shape fit of the molecule D: Geometry stability - AIMNet geometry optimisation was completed with Auto3D using the SMILES string of each sample.
+RMSD was calculated between the predicted geometry and the sampled geometry using RDKit.
+Horizontal Lines indicated the sample median and circles the sample mean.}
+    \label{fig:measures}
+\end{figure}
 \subsection{Intermediate SILVR rates produce stable and varied molecules}
@@
 are not a single connected molecular graph, with respect to $r_{\mathrm{S}}$.
-This is presented in Figure S\ref{fig:fragmentation} in the SI.
+This is presented in Figure~\ref{fig:fragmentation} in the SI.
 At an intermediate $r_{\mathrm{S}} = 0.025$ just over 50\% of the generated samples are not fragmented meaning that, only one in two generated molecules can be analysed further.
@@
 
-\begin{figure}[ht!]
-    \centering
-    \includegraphics[width=0.8\textwidth]{paper/Figures/Fig3/fig_3_metrics_violins.png}
-    \caption{Validation measures of SILVR model using fragments \texttt{x0072} and \texttt{x0354} as reference coordinates.
-A: Ratio of stable atoms - an atom is determined as stable if the valence matches the expected valence for the element B: RMSD from reference - the calculated RMSD between the reference and sample, using an absolute one-to-one mapping ignore atom identity with low RMSD meaning molecules are similar to the reference and high RMSD they are not.
-C: OpenEye measure \textit{shapegauss} - a Gaussian scoring function describing the shape fit between Mpro and samples, ignoring chemical interactions.
-A lower score means a better shape fit of the molecule D: Geometry stability - AIMNet geometry optimisation was completed with Auto3D using the SMILES string of each sample.
-RMSD was calculated between the predicted geometry and the sampled geometry using RDKit.
-Lines indicated the sample median and circles the sample mean.}
-    \label{fig:measures}
-\end{figure}
 The similarity of samples to their reference set of coordinates was determined by RMSD, with a clear inverse correlation observed between $r_{\mathrm{S}}$ and RMSD, as seen in Figure~\ref{fig:measures} B.
@@
 Importantly, no difference is seen between the control set ($r_{\mathrm{S}}=0$) and the SILVR samples ($0 < r_{\mathrm{S}} < 0.01$) indicating the quality of generated molecules was not impeded by the SILVR protocol.
+ The synthetic accessibility (RDKit SA score~\cite{ertl2009estimation}) (Figure~\ref{fig:SA_QED} A) and the Quantitative Estimation of Druglikeness (QED) (Figure~\ref{fig:SA_QED} B) were also estimated as shown in Figure~\ref{fig:SA_QED} in the SI.
+SILVR does not affect the SA score, or massively change QED for $r_{\mathrm{S}}$ with the best outcomes.
 
@@
 The results of these experiments continue the same trends previously observed where the $r_{\mathrm{S}}=0.005$ produces samples of approximate geometric similarity, whereas $r_{\mathrm{S}}=0.01$ produces a more strict mapping, with a clearly preserved urea group, a slightly modified ring system, and an amide interpreted as a carboxylate.
+When varying the number of dummy atoms used for linker generation, the atoms stability measure is not impacted for $r_{\mathrm{S}}=0.005$, as seen in Figure~\ref{fig:dummy_atoms} A and for $r_{\mathrm{S}}=0.01$ (Figure~\ref{fig:dummy_atoms} B) it only marginally improves with more dummy atoms.
+Using a better EDM which resolves explicit and implicit hydrogens better will likely improve this more.
 
@@
 The supporting information consists of a series of figures in addition to the figures in the main paper.
-Figure S~\ref{fig:moonshot_data} contains the 2D structures of all fragments used in the experiments as references.
-In Figure S~\ref{fig:examples_curated} an example of two curated experiments is shown with different SILVR rates $r_{\mathrm{s}}$.
-Next, Figure S~\ref{fig:fragmentation}, shows on average how many molecules were not-fragmented in a set of 1000 samples for different SILVR rates.
-Figure S~\ref{fig:SA_QED} includes synthetical accessibility(~\cite{ertl2009estimation}) and QED~\cite{bickerton2012quantifying, wildman1999prediction}.
-An example of how far atoms, even involved in a reference structure are displaced in the denoising diffusion process is shown in an example in Figure~\ref{fig:atom_displacement} \textbf{A} for a whole molecule generated and in \textbf{B} for the heterocycle.
+Figure~\ref{fig:moonshot_data} contains the 2D structures of all fragments used in the experiments as references.
+In Figure~\ref{fig:examples_curated} an example of two curated experiments is shown with different SILVR rates $r_{\mathrm{s}}$.
+Next, Figure~\ref{fig:fragmentation}, shows on average how many molecules were not-fragmented in a set of 1000 samples for different SILVR rates.
+Figure~\ref{fig:SA_QED} includes synthetic accessibility(~\cite{ertl2009estimation}) and QED~\cite{bickerton2012quantifying, wildman1999prediction}.
+An example of how far atoms, even involved in a reference structure are displaced in the denoising-diffusion process is shown in an example in Figure~\ref{fig:atom_displacement} \textbf{A} for a whole molecule generated and in \textbf{B} for the heterocycle.
+A summary of experiments around the number of dummy atoms used in linker design experiments is seen in Figure~\ref{fig:dummy_atoms} for $r_{\mathrm{s}}=0.005$ in A and $r_{\mathrm{s}}=0.01$ in B.
 The final Figure~\ref{fig:uncurated_sample} shows a series of 2D samples from an uncurated list of samples, clearly showing examples of fragmented molecules.
@@
 It was hoped that in the case of disconnected fragments, these dummy atoms would be able to form a linker.
-In practice, linker formation was observed both with and without dummy atoms, however, the presence of dummy atoms seemed to produce better results (by visual inspection).
+In practice, linker formation was observed both with and without dummy atoms.
 It was also observed that dummy atoms often populate unfilled valences as hydrogen atoms.
 
-The total number of atoms used by the model was set as the sum reference (nr) and dummy atoms (nd).
-The SILVR vector was created such that the first nr indices held the value of rs defined in the experiment protocol, and the subsequent nd indices were set to zero so that SILVR would not be applied to the dummy atoms.
+The total number of atoms used by the model was set as the sum reference ($n_r$) and dummy atoms ($n_d$).
+The SILVR vector was created such that the first $n_r$ indices held the value of $r_{\mathrm{S}}$ defined in the experiment protocol, and the subsequent $n_d$ indices were set to zero so that SILVR would not be applied to the dummy atoms.
 

### 2023-04-21 paper/manuscript.tex
Final version for arxiv submission

@@
 \usepackage{amssymb}
-\usepackage[
+%\usepackage[
  %eprint=false,
- %style=plain
-]{biblatex}
+ %style=unsrt
+%]{biblatex}
 %\ExecuteBibliographyOptions[misc]{eprint=true}
+%\addbibresource{SILVR.bib}
+\usepackage[backend=biber,style=chem-acs, eprint=false]{biblatex}
+\ExecuteBibliographyOptions[misc]{eprint=true}
 \addbibresource{SILVR.bib}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%https://www.overleaf.com/project/63f3627b7f8554c152639004
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newcommand*\mycommand[1]{\texttt{\emph{#1}}}
+%\newcommand*\mycommand[1]{\texttt{\emph{#1}}}
 
@@
 \author{Nicholas T. Runcie}
-\altaffiliation{EaSTCHEM School of Chemistry, University of Edinburgh, EH9 3FJ}
+\affiliation{EaSTCHEM School of Chemistry, University of Edinburgh, EH9 3FJ}
+\email{n.runcie@sms.ed.ac.uk}
 \author{Antonia S.J.S. Mey}
-\altaffiliation{EaSTCHEM School of Chemistry, University of Edinburgh, EH9 3FJ}
+\affiliation{EaSTCHEM School of Chemistry, University of Edinburgh, EH9 3FJ}
 \email{antonia.mey@ed.ac.uk}
@@
 \label{sec:intro}
-Sampling from a very large space of possible drug-like compounds to find suitable hits for a given target protein is an open challenge in drug-design.
+Sampling from a very large space of possible drug-like compounds to find suitable hits for a given target protein is an open challenge in drug design.
 It is estimated that there are between $10^{23}$ to $10^{60}$ feasible compounds, while only around $10^8$ have been synthesised so far~\cite{polishchuk2013estimation, reymond2012exploring}.
@@
 Typically the objective is to generate a diverse set of molecules for a given target protein that are easily synthetically accessible, and ideally with high binding affinity and low predicted toxicity~\cite{prieto-martinez2019chapter}.
-A plethora of methods has been devised for the generation of molecules, but also assessment of binding affinity such as more traditional docking~\cite{fu2018predictive} and molecular simulation-based affinity prediction methods~\cite{mey2016blinded,mey2020best} and machine learning methods more recently.
-Machine learning has lead to an explosion of methods for molecular generation, from recurrent neural networks (RNN)~\cite{xu2021novo,zhang2022novo} and conditional variational autoencoders~\cite{ragoza2022generating} to GANs~\cite{skalic2019target}, various approaches have been used to generate molecules for a binding site where in each case the training is conditioned based on the target protein~\cite{xie2022advances}.
-Some of these models incorporate a ligand score directly~\cite{ragoza2022generating}, while others require methods based on machine learning (ML)~\cite{masters2023deep}, or more conventional methods.
+A plethora of methods have been devised for the generation of molecules, as well as assessing their suitability as drug candidates.
+For example, for binding affinity predictions traditional docking~\cite{fu2018predictive} and molecular simulation-based affinity prediction methods~\cite{mey2016blinded,mey2020best} have dominated the field until recently.
+Now ML methods are gaining momentum and various approaches have been used to generate molecules for a binding site where in each case the training is conditioned towards the target protein~\cite{xie2022advances}.
+Some of these models incorporate a ligand score directly~\cite{ragoza2022generating}, while others require methods based on machine learning (ML)~\cite{masters2023deep}, or more conventional affinity prediction methods downstream (e.g.
+docking or free energy calculations).
 Even with a variety of ways to assess the synthetic accessibility of generated compounds~\cite{coley2018scscore, thakkar2021retrosynthetic}, molecules generated with these methods are often not easily synthesizable and in the worst cases can be chemically infeasible.
@@
 These fragments are selected such that they present promiscuous binding, allowing exploration of many types of interactions a drug-like molecule could adopt within a given target.
-Individual fragments can not be drugs in and of themselves as they do not possess enough intermolecular interactions to achieve a sufficient binding affinity with a target, however, by considering an ensemble of known fragment hits, new high affinity binders can be constructed by merging and linking known fragments together and elaborating on singular fragments.
+Individual fragments can not be drugs in and of themselves as they do not possess enough intermolecular interactions to achieve a sufficient binding affinity with a target, however, by considering an ensemble of known fragment hits, new high-affinity binders can be constructed by merging and linking known fragments together and elaborating on singular fragments.
 An array of screening methods exist for determining if fragments bind a target, however here we focus on X-ray crystallography techniques.
-Protein drug targets can either be co-crystallized with fragments, or be crystalised unbound and subsequently soaked in a fragment solution.
+Protein drug targets can either be co-crystallized with fragments or be crystalised unbound and subsequently soaked in a fragment solution.
 These crystals can be resolved by X-ray crystallography, with the results showing a high-quality electron density map revealing the exact binding geometry and interactions a fragment obtains with the given target.
 % A couple of sentences on the covid Moonshot dataset here to set the scene. 
-An application of generative models is therefore in the interpretation of such fragment data for the automated design of high affinity binders.
-Goa \textit{et al.} have introduced a way to generate linkers for fragments using reinforcement learning strategies~\cite{guo2023linkinvent}, while Imrie \textit{et al.}~\cite{imrie2020deep} have used variational auto encoders on this task without the use of protein information.
-More recently Huang \textit{et al.}~\cite{huang20223dlinker} and Igashov \textit{et al.}~\cite{igashov2022equivariant} have tackled this challenge using equivariant variational auto-encoder and diffusion based models respectively.
+An application of generative models is therefore in the interpretation of such fragment data for the automated design of high-affinity binders.
+Goa \textit{et al.} have introduced a way to generate linkers for fragments using reinforcement learning strategies~\cite{guo2023linkinvent}, while Imrie \textit{et al.}~\cite{imrie2020deep} have used variational autoencoders on this task without the use of protein information.
+More recently Huang \textit{et al.}~\cite{huang20223dlinker} and Igashov \textit{et al.}~\cite{igashov2022equivariant} have tackled this challenge using equivariant variational auto-encoder and diffusion-based models respectively.
 Each of these models were explicitly trained for the specific purpose of linker generation.
@@
 In this paper, we present SILVR, a selective iterative latent variable refinement (SILVR) method for conditioning an existing pre-trained equivariant diffusion model (EDM) towards the task of fragment merging and linker generation, yielding compounds similar to existing hits without specific training towards this task.
-To achieve this we combine the EDM by Hoogeboom \textit{et al.}~\cite{hoogeboom2022equivariant} with the iterative latent variable refinement method proposed by Choi et al.~\cite{choi2021ilvr} for image generation networks.
+To achieve this we combine the EDM by Hoogeboom \textit{et al.}~\cite{hoogeboom2022equivariant} with the iterative latent variable refinement method proposed by Choi \textit{et al.}~\cite{choi2021ilvr} for image generation networks.
 This allows the generation of new molecules in the shape of a binding site using information from existing fragment hits, without specific training or knowledge of this task.
@@
 They are evolved over time $t$ in the diffusion process to resemble a Gaussian distribution at $t=T$, see Equation~\ref{eqn:noising}.
-The $\beta$ added represents the noise added at each step, and the dots show the omitted steps from time $t=3$ to $t=T$.
+The $\beta$ represents the noise added at each step, and the dots show the omitted steps from time $t=3$ to $t=T$.
 As atoms effectively 'diffuse' they can be perceived as changing position.
@@
 What we are actually interested in is learning the reverse process, i.e.
-the denoising and generating $\mathbf{x}$, however the reverse of the process $q(\mathbf{x}_{t-1}|\mathbf{x}_{t=T})$ is intractable.
+the denoising and generating a new sample $\mathbf{\hat{x}}_0$, however the reverse of the process $q(\mathbf{x}_{t-1}|\mathbf{x}_{t=T})$ is intractable.
 The DDPM learns this reverse transitions $p_{\theta}(\hat{\mathbf{x}}_{t-1}|\mathbf{x}_{t=T})$, which is also a Gaussian transition kernel.
@@
 \begin{equation}
- \hat{\mathbf{x}}_{t-1} = \frac{1}{\alpha_t}(\mathbf{x}_t-\frac{1-\alpha_t}{\sqrt{1-\langle\alpha_t\rangle}}\phi(\mathbf{x}_t,t)) +\sigma \rho,   
+ \hat{\mathbf{x}}_{t-1} = \frac{1}{\alpha_t}(\mathbf{x}_t-\frac{1-\alpha_t}{\sqrt{1-\langle\alpha_t\rangle}}\phi(\mathbf{\hat{x}}_t,t)) +\sigma \rho,   
 \end{equation}
@@
 
-The process is iterated until $t=1$ and as such a new sample $\mathbf{x}_0$ of the denoising process---which is intended to represent a proposed molecule design---is generated.
+The process is iterated until $t=1$ and as such a new sample $\mathbf{\hat{x}}_0$ of the denoising process---which is intended to represent a proposed molecule design---is generated.
 
@@
 \subsection*{Equivariant diffusion model (EDM)}
-In the previous section introduced the neural network $\phi(\mathbf{x}_t,t)$.
-In practice, it makes sense to use an euqivariant graph neural network, as it is a data efficient way to learn about molecules.
-If a model has rotational and translational equivarience, it means a neural network does not need to learn orientations and transations of molecules from scratch.
+In the previous section we introduced the neural network $\phi(\mathbf{x}_t,t)$.
+In practice, it makes sense to use an equivariant graph neural network, as it is a data-efficient way to learn about molecules.
+If a model has rotational and translational equivariance, it means a neural network does not need to learn orientations and translations of molecules from scratch.
 We chose the EDM by Hoogeboom \textit{et al.}~\cite{hoogeboom2022equivariant} as our baseline generative model, as it provides a generative model for new molecules and has equivariance already built into it.
-Furthermore, it has all code and pretrained weights available online~\cite{githubhoogeboom}.
-The basic concept behind equivarience is that the model is invariant to rotations and translation, in this case the E(3) group.
+Furthermore, it has all code and pre-trained weights available online at:~\url{https://github.com/ehoogeboom/e3_diffusion_for_molecules}.
+The basic concept behind equivariance is that the model is invariant to rotations and translation, in this case the E(3) group.
 This means that scalar (features such as atom types) and vector node properties (such as the positions) will be invariant to group transformations.
@@
 The input to the model can be rotated and diffusion/denoising  applied to get a structure, or the diffusion/denoising process can be applied first followed by the same rotation to get the same output.
-Mathematically this means that if we have a set of point $\mathbf{x} = (\mathbf{x}_1,\ldots,\mathbf{x}_N) \in \mathbb{R}^{N\times 3}$ and each of these points has an associated set of scalar feature vectors $\mathbf{h}\in \mathbb{R}^{N\times k}$ which are invariant to group transformations.
-The position translations and rotations are defined according to the orthogonal matrix: $\mathbf{Rx + t} = (\mathbf{R_{x_1}+t},\ldots, \mathbf{R_{x_N}+t})$.
+Mathematically this means that if we have a set of point $\mathbf{x} = (\mathbf{x}_1,\ldots,\mathbf{x}_N) \in \mathbb{R}^{N\times 3}$ and each of these points has an associated set of scalar feature vectors $\mathbf{h}\in \mathbb{R}^{N\times k}$ which are invariant to group transformations, the position translations and rotations are defined according to the orthogonal matrix: $\mathbf{Rx + t} = (\mathbf{R_{x_1}+t},\ldots, \mathbf{R_{x_N}+t})$.
 Satorras \textit{et al.}~\cite{satorras2022equivariant} have proposed an E(n) Equivariant Graph Neural Network on which the EDM by Hoogeboom \textit{et al.} builds.
-The actual diffusion process of the Hoogeboom model relies taking place on a set of points $M$ $\{\{\mathbf{x}_i,\mathbf{h}_i\}\}_{i\ldots M}$ on a latent variable that combines atom coordinates $\mathbf{x}_i$ and node features of each atom $\mathbf{h}_i$, such that: $\mathbf{z}_t = [\mathbf{z}_t^{(x)}, \mathbf{z}_t^{(h)}]$.
+The actual diffusion process of the Hoogeboom model relies on a set of points $M$ $\{\{\mathbf{x}_i,\mathbf{h}_i\}\}_{i\ldots M}$ for a latent variable that combines atom coordinates $\mathbf{x}_i$ and node features of each atom $\mathbf{h}_i$, such that: $\mathbf{z}_t = [\mathbf{z}_t^{(x)}, \mathbf{z}_t^{(h)}]$.
 The node features in practice are an array of values containing information such as atom type.
-These features are encoded using a onehot encoding.
+These features are encoded using a one-hot encoding.
 For more details on this see~\cite{hoogeboom2022equivariant}.
@@
 \begin{equation}
-    q(\mathbf{z}_t| \mathbf{x},\mathbf{h}) = \mathcal{N}_{xh}(\mathbf{z}_t|\alpha_t[\mathbf{x,h}],\sigma^2_t\mathbf{I}].
+    q(\mathbf{z}_t| \mathbf{x},\mathbf{h}) = \mathcal{N}_{xh}(\mathbf{z}_t|\alpha_t[\mathbf{x,h}],\sigma^2_t\mathbf{I}).
 \end{equation}
-In the same way, for the genereative denoising process the distribution can be written as:
+In the same way, for the generative denoising process the distribution can be written as:
 \begin{equation}
@@
 We use the notation $\kappa$ to generate a sample $\kappa = [\mathbf{\hat{x}}, \mathbf{\hat{h}}]$ from the EDM. 
-For more details details on the architecture and Hoogeboom's code see~\cite{hoogeboom2022equivariant}. 
+For more details on the architecture and Hoogeboom's code see~\cite{hoogeboom2022equivariant}. 
 %Something on latent variables Z.
@@
 \subsection{Iterative Latent Variable Refinement as a conditioning tool}
-Conditioning diffusion models is often desirable to for example generate similar images to that of an original input images, or in the example of Hoogeboom generating molecules in the presence of an external electric field resulting in more polarizable molecules.
+Conditioning diffusion models is often desirable to, for example, generate similar images to that of original input images, or in the example of Hoogeboom, generate molecules in the presence of an external electric field resulting in more polarizable molecules.
 However, this conditioning requires retraining the network to accommodate for this condition.
@@
 \begin{equation}
-    p_{\theta}[x_{\{0:T\}}|c]=p(x_T)\prod_{t=1}^{T}p_{\theta}(x_{t-1}|x_t,c)
+    p_{\theta}[x_{\{0:T\}}|c]=p(x_T)\prod_{t=1}^{T}p_{\theta}(x_{t-1}|x_t,c).
 \end{equation}
-Their trick is to use a reference image $\mathbf{y}$ and place it in the same downsampled fillter $\psi(\mathbf{y})$ as the generated image $\psi(\mathbf{\hat{x}}_0)$.
-This means the original image and the reference are  in the same latent space.
-So in each denoising step the proposal distribution is matched with that of the reference $y$ noised up to the appropriate timestep.
+Their trick is to use a reference image $\mathbf{y}$ and place it in the same downsampled filter $\psi(\mathbf{y})$ as the generated image $\psi(\mathbf{\hat{x}}_0)$.
+This means the original image and the reference are  in the same latent space, so in each denoising step the proposal distribution is matched with that of the reference $\mathbf{y}_t$ noised  to the appropriate timestep.
 An unconditional distribution at $t$ is generated first:
@@
 We consider latent space variables $\mathbf{z}=[\mathbf{z}^{(x)},\mathbf{z}^{(h)}]$ for the standard denoising process and  $\mathbf{\tilde{z}} = [\mathbf{z}^{(y)},\mathbf{z}^{(h_y)}]$, for the set of reference coordinates given by $\mathbf{y}$.
-The vector $\mathbf{h}$ contains all the scalar node properties of the equivariant EDM for the reference.
+The vector $\mathbf{h}_y$ contains all the scalar node properties of the equivariant EDM for the reference.
 Similarly to Choi \textit{et al.}, we update the diffusion process at noise level $t$ in the latent space $\mathbf{z}$, with the reference $\mathbf{\tilde{z}}$ using a factor, or vector if used at variable rates for different atoms $r_{\mathrm{S}}$.
@@
 This is done from the first denoising timestep (top) until the last one (bottom).
-Repeatedly applying SILVR will result in a molecule that resembles the references for $r_{\mathrm{S}}>0$ (right) and does not for $r_{\mathrm{S}}=0$ (left)}
+Repeatedly applying SILVR will result in a molecule that resembles the references for $r_{\mathrm{S}}>0$ (right) and does not for $r_{\mathrm{S}}=0$ (left).}
     \label{fig:silvr_explanation}
@@
 The EDM by Hoodgeboom \textit{et al.} was trained on the 30 lowest energy conformations of 430,000 molecules from the GEOM dataset with explicit hydrogens~\cite{axelrod2022geom}.
-One appealing feature of GEOM is that it contains drug-like molecules including 6000 compounds for SARS-CoV-2 targets making this an appealing dataset for generating new molecules that combine or expand fragments for a SARS-CoV-2 target.
+One desirable feature of GEOM is that it contains drug-like molecules including 6000 compounds for SARS-CoV-2 targets making this an appealing dataset for generating new molecules that combine or expand fragments for a SARS-CoV-2 target.
 For more details on the model theory and training refer to Hoogeboom \textit{et al.}~\cite{hoogeboom2022equivariant}.
@@
 That is, the amount of structure remaining from the reference should match the amount of structure formed by the generative process.
-A small refinement is applied to add information from the reference coordinates to the latent variable of the denoising process (see line 11 of the algorithim and equation~\ref{eqn:SILVR}.
+A small refinement is applied to add information from the reference coordinates to the latent variable of the denoising process (see line 11 of the algorithm and equation~\ref{eqn:SILVR}).
 This equation has the effect of shrinking the coordinates towards the origin and then expanding the coordinates out in the direction of the reference, see Figure~\ref{fig:silvr_explanation}.
@@
 Reference molecules were selected from the original 23 non-covalent hits of the SARS-CoV-2 Main Protease (Mpro) identified by the XChem fragment screen~\cite{douangamath2020crystallographic} as part of the COVID Moonshot Project~\cite{consortium2023open, consortium2021open}.
-A more detailed picture of all fragments is presented in Figure S~\ref{fig:moonshot_data} of the Supplementary information(SI).
+A more detailed picture of all fragments is presented in Figure~\ref{fig:moonshot_data} of the Supplementary information (SI).
 Fragments were visualised using NGLview version 3.03 and combinations were arbitrarily selected as test cases for the following different experimental settings for trying to understand the similarity between the reference and the new sample.
@@
 Qualitatively, the generated molecular samples from the conditioned EDM  show a clear resemblance to their reference structures, with similarity increasing with $r_{\mathrm{S}}$.
-Figure S~\ref{fig:examplese_curated}, in the SI, shows two example samples started from fragments \texttt{x0072} and \texttt{x0354} over a range of SILVR rates between $r_{\mathrm{S}}=0$ to $r_{\mathrm{S}}=0.02$.
+Figure~\ref{fig:examples_curated}, in the SI, shows two example samples started from fragments \texttt{x0072} and \texttt{x0354} over a range of SILVR rates between $r_{\mathrm{S}}=0$ to $r_{\mathrm{S}}=0.02$.
 As expected at no conditioning random samples are generated that do not resemble the reference fragments.
@@
 Of particular interest is the formation of the fused ring system.
-At first glance, it might be assumed that reference atoms map to the sample atom closest in space, however in actuality they travel up to 1.7 \AA~to arrive at their final position (Figure S\ref{fig:atom_displacement}) In this case, the nitrogen atoms observed in the fused ring are directly obtained from the nitrogen atoms in the reference, however, their final position is one bond’s length from their reference.
+At first glance, it might be assumed that reference atoms map to the sample atom closest in space, however in actuality they travel up to 1.7 \AA~to arrive at their final position (Figure~\ref{fig:atom_displacement} in the SI) In this case, the nitrogen atoms observed in the fused ring are directly obtained from the nitrogen atoms in the reference, however, their final position is one bond’s length from their reference.
 This shows the flexibility of each sample atom to explore within a radius (defined by $r_{\mathrm{S}}$) of the reference atom.
@@
 The results of these experiments continue the same trends previously observed where the $r_{\mathrm{S}}=0.005$ produces samples of approximate geometric similarity, whereas $r_{\mathrm{S}}=0.01$ produces a more strict mapping, with a clearly preserved urea group, a slightly modified ring system, and an amide interpreted as a carboxylate.
-When varying the number of dummy atoms used for linker generation, the atoms stability measure is not impacted for $r_{\mathrm{S}}=0.005$, as seen in Figure~\ref{fig:dummy_atoms} A and for $r_{\mathrm{S}}=0.01$ (Figure~\ref{fig:dummy_atoms} B) it only marginally improves with more dummy atoms.
+When varying the number of dummy atoms used for linker generation, the atoms stability measure is not impacted for $r_{\mathrm{S}}=0.005$, as seen in Figure~\ref{fig:dummy_atoms} A and for $r_{\mathrm{S}}=0.01$ (Figure~\ref{fig:dummy_atoms} B) it only marginally improves with more dummy atoms found in the SI.
 Using a better EDM which resolves explicit and implicit hydrogens better will likely improve this more.
@@
 When applied in a drug-design context, the conservation of key functional groups in exact spacial positions is crucial to maintain desirable protein-ligand interactions.
-The series of molecules in Figure S2 shows a loss of the sulfonyl chloride group present in the reference, which may be undesirable.
+The series of molecules in Figure~\ref{fig:examples_curated} of the SI shows a loss of the sulfonyl chloride group present in the reference, which may be undesirable.
 This issue can be solved by changing $r_{\mathrm{S}}$ from a scalar to a vector ($\mathbf{r}_{\mathrm{S}}$), and by assigning particularly high $\mathbf{r}_{\mathrm{S}}$ values to selected atoms of the reference.
@@
 An EDM with explicit hydrogens will improve the overall models.
-At the moment there is a mixing of explicit and implicity hydrogens depending on need for analysis and input.
-An optimal model can account for hydrogens both explicityly and implicityly allowing for scoring of either.
+At the moment there is a mixing of explicit and implicit hydrogens depending on the need for analysis and input.
+An optimal model can account for hydrogens both explicitly and implicitly allowing for scoring of either.
 In addition using dummy atoms strategically for growing certain parts of a fragment is something to be explored further in the future.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\bibliographystyle{hunsrt}
+%\bibliographystyle{unsrt}
+%\bibliography{SILVR}
 \printbibliography
@@
 Figure~\ref{fig:SA_QED} includes synthetic accessibility(~\cite{ertl2009estimation}) and QED~\cite{bickerton2012quantifying, wildman1999prediction}.
-An example of how far atoms, even involved in a reference structure are displaced in the denoising-diffusion process is shown in an example in Figure~\ref{fig:atom_displacement} \textbf{A} for a whole molecule generated and in \textbf{B} for the heterocycle.
+An example of how far atoms in the reference structure are displaced in the denoising-diffusion process is shown in an example in Figure~\ref{fig:atom_displacement} \textbf{A} for a whole molecule generated and in \textbf{B} for the heterocycle.
 A summary of experiments around the number of dummy atoms used in linker design experiments is seen in Figure~\ref{fig:dummy_atoms} for $r_{\mathrm{s}}=0.005$ in A and $r_{\mathrm{s}}=0.01$ in B.
@@
 
-The total number of atoms used by the model was set as the sum reference ($n_r$) and dummy atoms ($n_d$).
+The total number of atoms used by the model was set as the sum of reference ($n_r$) and dummy atoms ($n_d$).
 The SILVR vector was created such that the first $n_r$ indices held the value of $r_{\mathrm{S}}$ defined in the experiment protocol, and the subsequent $n_d$ indices were set to zero so that SILVR would not be applied to the dummy atoms.
